### 2019-10-21 Paper/paper.tex
paper text + iso2c from WID

%!TEX program = xelatex

\documentclass[a4paper,11pt]{article}
\usepackage{graphicx}
\usepackage{amsmath, amsthm, amssymb} %better typesetting of mathematical expressions, theorems, additional symbols
\usepackage{booktabs} %fixltx2e
\usepackage[flushleft]{threeparttable}
\usepackage{tabularx}
% \usepackage[capposition=top]{floatrow}
\usepackage{caption}
\captionsetup[measuredfigure]{labelformat=empty}
% \usepackage{bm}
% \usepackage{subcaption}
%\usepackage{epstopdf} % not supported by xelatex
% \usepackage{fullpage}
% \usepackage{enumerate}
\usepackage{authblk}
\usepackage{fontspec}
\usepackage{lscape}
\usepackage[round]{natbib} %standard package for bibliography
\usepackage{multirow}
% \usepackage[figuresleft]{rotating}
\usepackage[dvipsnames]{xcolor}
\usepackage{colortbl} % color table
\usepackage{geometry}
% \usepackage{hyperref} %hyperlink support
% \hypersetup{pdfstartview={XYZ null null 1.00},bookmarksnumbered, hypertexnames=false, colorlinks=true, linkcolor=BrickRed, citecolor=MidnightBlue, urlcolor=MidnightBlue} % zoom: default screen; numbers for bookmarks

%
%
%\usepackage{parskip}
\usepackage[nolist]{acronym}
\usepackage{longtable}
\usepackage{setspace}

\acrodef{bma}[BMA]{Bayesian Model Averaging}
\acrodef{bic}[BIC]{Bayesian Information Criterion}
\acrodef{cbf}[CBF]{Conditional Bayes Factor}
\acrodefplural{cbf}[CBFs]{Conditional Bayes Factors}
\acrodef{cswd}[CSWD]{Credit Suisse Wealth Databook}
\acrodef{gfdd}[GFDD]{Global Financial Development Database}
\acrodef{pip}[PIP]{Posterior Inclusion Probability}
\acrodefplural{pip}[PIPs]{Posterior Inclusion Probabilities}
\acrodef{fdi}[FDI]{Foreign Direct Investment}
\acrodefplural{fdi}[FDIs]{Foreign Direct Investments}
\acrodef{gdp}[GDP]{Gross Domestic Product}
\acrodef{uip}[UIP]{Unit Information Prior}
\acrodef{wb}[WB]{World Bank}
\acrodef{oecd}[OECD]{Organisation for Co-operation and Development}
\acrodef{ivbma}[IVBMA]{Instrumental Variable Bayesian Model Averaging}
\acrodef{wid}[WID]{World Inequality Database}
\acrodef{hbs}[HBS]{Household Balance Sheet}
\acrodef{uk}[UK]{United Kingdom}
\acrodef{us}[US]{United States}
\acrodef{efw}[EFW]{Economic Freedom of the World}
\acrodef{pmp}[PMP]{Posterior Model Probability}

%
%

\setlength\parindent{34pt}
\onehalfspacing
\title{Finance and Wealth Inequality\thanks{We thank two anonymous referees, ...}}
\author[a]{Roman Horvath}
\author[a]{Jan Mares\footnote{\footnotesize Corresponding author's address: IES FSV UK, Opletalova 26, 110 00, Praha 1; \textbf{e-mail:janxmares@gmail.com}}}
\affil[a]{Charles University, Prague}

%\renewcommand\Authands{ and }
\date{\today}

\begin{document}

\clubpenalty 9999 % no orphants (typographic properties)
\widowpenalty 9999 % no widows (typographic properties)
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi} % shortcut for Stata tables

\maketitle

\thispagestyle{empty}
\begin{abstract}
Using a global sample, this paper investigates the determinants of wealth inequality capturing various economic, financial, political, institutional, and geographical indicators. Using instrumental variable Bayesian model averaging, it reveals that only a handful of indicators robustly matters and finance plays a key role. It reports that while financial depth increases wealth inequality, efficiency and access to finance reduce inequality. In addition, redistribution and education are associated with lower inequality whereas wars and openness to international trade contribute to greater wealth inequality.
\end{abstract}

\bigskip

\begin{tabular}{p{0.25\hsize}p{0.6\hsize}} %0.15
\textbf{Keywords:} & Income inequality, finance, Bayesian model averaging
\end{tabular}

\bigskip

\begin{tabular}{p{0.25\hsize}p{0.6\hsize}}
\textbf{JEL Codes:} & D31, E21\\
\end{tabular}

\clearpage
\setcounter{page}{1}

\section{Introduction}
% \label{sec:intro}

\section{Related literature}
% \label{sec:literature}
\citet{pikettyandzucman2014}
\section{Data}
% \label{sec:data}

\section{Methodology}
% \label{sec:meth}

\section{Results}
% \label{sec:results}

\section{Concluding Remarks}
% \label{sec:conclusion}

\clearpage
%
\bibliographystyle{chicagoa}
\bibliography{literature}
%
\clearpage
%
\appendix
\section{Appendix}
% \label{sec:app}

\renewcommand{\thesection}{A\arabic{section}}%
\renewcommand{\thetable}{A\arabic{table}}%
\renewcommand{\thefigure}{A\arabic{figure}}%
\renewcommand{\theequation}{A\arabic{eq}}

\end{document}

### 2019-10-21 181203 Union density from ILO.R

@@ -21,11 +21,13 @@ data_ilo_ud <- data.table(get_ilostat("ILR_TUMT_NOC_RT_A"))
 data_ilo_minwage <- data.table(get_ilostat("EAR_4MNP_NOC_NB_A"))
 data_ilo_meanwage <- data.table(get_ilostat("EAR_4MTH_SEX_ECO_CUR_NB_A"))
 
-head(data.ilo.meanwage)
+
+head(data_ilo_meanwage)
 
 data_ilo_ud[ref_area=="CZE",]
 
-unique(data.unions$ref_area)
+
+unique(data_ilo_ud$ref_area)
 
 
-View(data.ilo.meanwage)
+View(data_ilo_meanwage)

### 2019-10-21 191021 Read the WID iso2 codes.R

@@ -1,0 +1,36 @@
+# iso2 codes from WID world
+# Jan Mares
+# 21/10/2019
+
+# Libraries
+library(data.table)
+library(countrycode)
+library(stringr)
+library(countrycode)
+
+# Set the working directory
+wd <- c("c:/Users/JM/Dropbox/Research/Finance and income inequality/")
+setwd(wd)
+
+# read data
+iso2 <- data.table(read.csv("Auxiliary data/WID iso2codes.csv", header = F, stringsAsFactors=F))
+
+# correct the first csv line
+iso2[1, V1:="AD Andorra"]
+
+# set column name
+setnames(iso2, c("country"))
+
+# split by the first space and bind into data table
+l1 <- str_replace(iso2$country, "[:space:]{1}", "_")
+l1 <- str_split(l1, "_")
+l1_dt <- data.table(do.call(rbind, l1))
+
+# set the data table column names
+setnames(l1_dt, c("iso2c","country"))
+
+# check
+# View(l1_dt)
+
+# write into csv
+write.csv(l1_dt, file="Auxiliary data/WID iso2codes.csv", row.names = F)

### 2019-10-21 191021 WB data.R

@@ -1,0 +1,13 @@
+# iso2 codes from WID world
+# Jan Mares
+# 21/10/2019
+
+# Libraries
+library(data.table)
+library(countrycode)
+library(stringr)
+library(countrycode)
+
+# Set the working directory
+wd <- c("c:/Users/JM/Dropbox/Research/Finance and income inequality/")
+setwd(wd)

### 2019-10-22 181203 Union density from ILO.R
workgin through ILO mean wage data

@@ -1,7 +1,6 @@
 # Union density from ILO
 # Jan Mares
 # 3/12/2018
-
 # Analysis of WID data on inequality
 # Jan Mares, 181108
 
@@ -21,13 +20,102 @@ data_ilo_ud <- data.table(get_ilostat("ILR_TUMT_NOC_RT_A"))
 data_ilo_minwage <- data.table(get_ilostat("EAR_4MNP_NOC_NB_A"))
 data_ilo_meanwage <- data.table(get_ilostat("EAR_4MTH_SEX_ECO_CUR_NB_A"))
 
+#
+#
 
-head(data_ilo_meanwage)
+## minimum wage
 
-data_ilo_ud[ref_area=="CZE",]
+# filter required columns
+data_minwage <- data_ilo_minwage[, .(ref_area, indicator, time, obs_value)]
+setnames(data_minwage, c("iso3c","indicator","year","value"))
+
+# data_minwage[, country := countrycode(iso3c, "iso3c", "country.name")] # test iso3 codes
+
+# replace the indicator value by simpler one
+data_minwage[, indicator := "minwage"]
+head(data_minwage)
+
+#
+#
+
+## mean wage
+
+# filter values in PPP2011 dollars, totals (for both genders)
+type <- c("ECO_AGGREGATE_TOTAL", "ECO_AGGREGATE_AGR","ECO_AGGREGATE_MAN",
+	 	  "ECO_AGGREGATE_CON","ECO_AGGREGATE_MEL","ECO_AGGREGATE_MKT",
+	 	  "ECO_AGGREGATE_PUB")
+ 
+data_meanwage <- data_ilo_meanwage[sex == "SEX_T" & classif1 %in% type & classif2 == "CUR_TYPE_PPP2011",]
+
+# required columns
+data_meanwage <- data_meanwage[classif1 == "ECO_AGGREGATE_TOTAL", .(ref_area, source, indicator, type = classif1, year = time,
+							       value = obs_value)]
+
+#
+#
+
+# selecting the sources for series where we have multiple values
+# this is also to remove the duplicates, although not completely as we leave some to
+# link the series
+
+# armenia
+data_meanwage <- data_meanwage[!(ref_area == "ARM" & !(source %in% c("BA:867","DA:177"))),]
+
+# austria
+data_meanwage <- data_meanwage[!(ref_area == "AUT" & !(source %in% c("FA:883","FX:272"))),]
+
+# bulgaria 
[…]

### 2019-10-23 181203 Union density from ILO.R
finalizing code to adjust ILO mean wage

@@ -11,7 +11,7 @@ library(Rilostat)
 
 
 # Set the working directory
-wd <- c("c:/Users/JM/Dropbox/Research/Finance and income inequality/")
+wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
 setwd(wd)
 
 a <- get_ilostat_toc(search="union")
@@ -82,13 +82,116 @@ data_meanwage <- data_meanwage[!(ref_area == "CAN" & !(source %in% c("BA:147")))
 # chile
 data_meanwage <- data_meanwage[!(ref_area == "CHL" & !(source %in% c("DA:134","BX:138"))),]
 
+# costa rica
+data_meanwage <- data_meanwage[!(ref_area == "CRI" & !(source %in% c("BX:142"))),]
 
+# cyprus
+data_meanwage <- data_meanwage[!(ref_area == "CYP" & !(source %in% c("FA:38"))),]
 
+# czech republic
+data_meanwage <- data_meanwage[!(ref_area == "CZE" & !(source %in% c("DA:1050","DA:293"))),]
 
+# germany
+data_meanwage <- data_meanwage[!(ref_area == "DEU" & !(source %in% c("BB:304"))),]
 
+# egypt
+data_meanwage <- data_meanwage[!(ref_area == "EGY" & !(source %in% c("BA:582","CA:1162"))),]
 
+# spain
+data_meanwage <- data_meanwage[!(ref_area == "ESP" & !(source %in% c("DA:248","FA:1131"))),]
 
-View(data_meanwage)
+# greece
+data_meanwage <- data_meanwage[!(ref_area == "GRC" & !(source %in% c("BA:317"))),]
+
+# guatemala
+data_meanwage <- data_meanwage[!(ref_area == "GTM" & !(source %in% c("FA:1203","BA:660"))),]
+
+# hungary
+data_meanwage <- data_meanwage[!(ref_area == "HUN" & !(source %in% c("CA:1237","DA:3089"))),]
+
+# ireland
+data_meanwage <- data_meanwage[!(ref_area == "IRL" & !(source %in% c("DA:410"))),]
+
+# cambodia
+data_meanwage <- data_meanwage[!(ref_area == "KHM" & !(source %in% c("BB:3064","BA:1321"))),]
+
+# lithuania
+data_meanwage <- data_meanwage[!(ref_area == "LTU" & !(source %in% c("DA:12956","EA:6396"))),]
+
+# luxembourg
[…]

### 2019-10-24 180511 Analysis of WID income data.R
191024 update

@@ -6,7 +6,7 @@ library(data.table)
 library(countrycode)
 
 # Set the working directory
-wd <- c("c:/Users/JM/Dropbox/Research/Finance and Income Inequality/")
+wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
 setwd(wd)
 
 data <- data.table(read.csv(file="WID data top income shares.csv", header=T, stringsAsFactors=F))

### 2019-10-24 181203 Data WGI.R

@@ -12,7 +12,7 @@ library(devtools)
 library(WDI)
 
 # Set the working directory
-wd <- c("c:/Users/JM/Dropbox/Research/Finance and income inequality/")
+wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
 setwd(wd)
 
 # Define governance indicators 

### 2019-10-24 181203 Union density from ILO.R

@@ -8,217 +8,36 @@
 library(data.table)
 library(countrycode)
 library(Rilostat)
+library(ggplot2)
 
 
 # Set the working directory
 wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
 setwd(wd)
 
-a <- get_ilostat_toc(search="union")
+srch <- get_ilostat_toc(search="union")
 
+# download the series froim ILOSTAT
 data_ilo_ud <- data.table(get_ilostat("ILR_TUMT_NOC_RT_A"))
-data_ilo_minwage <- data.table(get_ilostat("EAR_4MNP_NOC_NB_A"))
-data_ilo_meanwage <- data.table(get_ilostat("EAR_4MTH_SEX_ECO_CUR_NB_A"))
 
 #
 #
 
-## minimum wage
+# filter required columns + rename
+data_ilo_ud <- data_ilo_ud[,.(iso3c = ref_area, year = time, ud = obs_value)]
 
-# filter required columns
-data_minwage <- data_ilo_minwage[, .(ref_area, indicator, time, obs_value)]
-setnames(data_minwage, c("iso3c","indicator","year","value"))
-
-# data_minwage[, country := countrycode(iso3c, "iso3c", "country.name")] # test iso3 codes
-
-# replace the indicator value by simpler one
-data_minwage[, indicator := "minwage"]
-head(data_minwage)
-
-#
-#
-
-## mean wage
-
-# filter values in PPP2011 dollars, totals (for both genders)
-type <- c("ECO_AGGREGATE_TOTAL", "ECO_AGGREGATE_AGR","ECO_AGGREGATE_MAN",
-	 	  "ECO_AGGREGATE_CON","ECO_AGGREGATE_MEL","ECO_AGGREGATE_MKT",
-	 	  "ECO_AGGREGATE_PUB")
- 
-data_meanwage <- data_ilo_meanwage[sex == "SEX_T" & classif1 %in% type & classif2 == "CUR_TYPE_PPP2011",]
-
-# required columns
-data_meanwage <- data_meanwage[classif1 == "ECO_AGGREGATE_TOTAL", .(ref_area, source, indicator, type = classif1, year = time,
-							       value = obs_value)]
-
-#
-#
-
-# selecting the sources for series where we have multiple values
-# this is also to remove the duplicates, although not completely as we leave some to
-# link the series
-
-# armenia
[…]

### 2019-10-24 191021 Read the WID iso2 codes.R

@@ -9,7 +9,7 @@ library(stringr)
 library(countrycode)
 
 # Set the working directory
-wd <- c("c:/Users/JM/Dropbox/Research/Finance and income inequality/")
+wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
 setwd(wd)
 
 # read data

### 2019-10-24 191021 WB data.R

@@ -9,5 +9,5 @@ library(stringr)
 library(countrycode)
 
 # Set the working directory
-wd <- c("c:/Users/JM/Dropbox/Research/Finance and income inequality/")
+wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
 setwd(wd)

### 2019-10-24 191024 minimum and mean wage ILO.R

@@ -1,0 +1,204 @@
+# Minimum and mean wage from ILO
+# Jan Mares
+# 24/10/2019
+
+# Libraries
+library(data.table)
+library(countrycode)
+library(Rilostat)
+
+
+# Set the working directory
+wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
+setwd(wd)
+
+# both indicators in PPP2011 USD
+
+data_ilo_minwage <- data.table(get_ilostat("EAR_4MNP_NOC_NB_A"))
+data_ilo_meanwage <- data.table(get_ilostat("EAR_4MTH_SEX_ECO_CUR_NB_A"))
+
+#
+#
+
+## minimum wage
+
+# filter required columns
+data_minwage <- data_ilo_minwage[, .(ref_area, indicator, time, obs_value)]
+setnames(data_minwage, c("iso3c","indicator","year","value"))
+
+# data_minwage[, country := countrycode(iso3c, "iso3c", "country.name")] # test iso3 codes
+View(data_ilo_minwage)
+# replace the indicator value by simpler one
+data_minwage[, indicator := "minwage"]
+head(data_minwage)
+
+#
+#
+
+## mean wage
+
+# filter values in PPP2011 dollars, totals (for both genders)
+type <- c("ECO_AGGREGATE_TOTAL", "ECO_AGGREGATE_AGR","ECO_AGGREGATE_MAN",
+	 	  "ECO_AGGREGATE_CON","ECO_AGGREGATE_MEL","ECO_AGGREGATE_MKT",
+	 	  "ECO_AGGREGATE_PUB")
+ 
+data_meanwage <- data_ilo_meanwage[sex == "SEX_T" & classif1 %in% type & classif2 == "CUR_TYPE_PPP2011",]
+
+# required columns
+data_meanwage <- data_meanwage[classif1 == "ECO_AGGREGATE_TOTAL", .(ref_area, source, indicator, type = classif1, year = time,
+							       value = obs_value)]
+
+#
+#
+
+# selecting the sources for series where we have multiple values
+# this is also to remove the duplicates, although not completely as we leave some to
+# link the series
+
+# armenia
+data_meanwage <- data_meanwage[!(ref_area == "ARM" & !(source %in% c("BA:867","DA:177"))),]
+
+# austria
[…]

### 2019-10-24 191024 minimum and median wage OECD.R

@@ -1,0 +1,35 @@
+# minimum/median wages OECD
+# Jan Mares
+# 24/10/2019
+
+# Libraries
+library(data.table)
+library(countrycode)
+library(stringr)
+library(countrycode)
+library(rsdmx)
+
+# Set the working directory
+wd <- c("c:/Users/JM/Dropbox/Research/Finance and income inequality/")
+setwd(wd)
+
+# read the OECD data on ratio btetween minimum wage and median wage
+# define the query for min/median ratio
+q <- "https://stats.oecd.org/restsdmx/sdmx.ashx/GetData/MIN2AVE/AUS+BEL+CAN+CHL+CZE+EST+FRA+DEU+GRC+HUN+IRL+ISR+JPN+KOR+LVA+LTU+LUX+MEX+NLD+NZL+POL+PRT+SVK+SVN+ESP+TUR+GBR+USA+CRI+ROU+COL.MEAN+MEDIAN/all?startTime=1960&endTime=2018"
+oecd_wage <- readSDMX(q, isURL = T)
+
+# transform this to data table
+oecd_wage_dt <- as.data.table(oecd_wage)
+
+#
+#
+
+# define the query for statutory minimum wage
+q <- "https://stats.oecd.org/restsdmx/sdmx.ashx/GetData/MW_CURP/AUS+BEL+CAN+CHL+CZE+EST+FRA+DEU+GRC+HUN+IRL+ISR+JPN+KOR+LVA+LTU+LUX+MEX+NLD+NZL+POL+PRT+SVK+SVN+ESP+TUR+GBR+USA+CRI+BRA+MLT+ROU+RUS+COL.H+D+W+M+A/all?startTime=1960&endTime=2018"
+oecd_minwage <- readSDMX(q, isURL = T)
+
+# transform in data table
+oecd_minwage_dt <- as.data.table(oecd_minwage)
+
+str(oecd_minwage_dt)
+head(oecd_minwage_dt)

### 2019-11-27 Scripts/191101 Merge all data.R
191127 update

@@ -33,6 +33,8 @@ leftwing <- data.table(read.csv(file = "Auxiliary data/to merge/IDB leftwing.csv
 swiid <- data.table(read.csv(file = "Auxiliary data/to merge/swiid.csv", header = T, stringsAsFactors = F))
 icrg <- data.table(read.csv(file = "Auxiliary data/to merge/icrg.csv", header = T, stringsAsFactors = F))
 glob <- data.table(read.csv(file = "Auxiliary data/to merge/kof.csv", header = T, stringsAsFactors = F))
+medage <- data.table(read.csv(file = "Auxiliary data/to merge/median age.csv", header = T, stringsAsFactors = F))
+
 
 #
 #
@@ -84,6 +86,7 @@ data <- merge(data, leftwing, by = c("iso3c","year"), all = T)
 data <- merge(data, swiid, by = c("iso3c","year"), all = T)
 data <- merge(data, icrg, by = c("iso3c","year"), all = T)
 data <- merge(data, glob, by = c("iso3c","year"), all = T)
+data <- merge(data, medage, by = c("iso3c","year"), all = T)
 
 # corrected for the missing waryears, revcoups, leftwign obs.
 data[is.na(WarYears), WarYears := 0]

### 2019-11-27 Scripts/191121 median age.R

@@ -1,0 +1,41 @@
+# Libraries
+library('data.table')
+library('countrycode')
+library('here')
+library('XLConnect')
+library('stringr')
+
+#
+#
+
+# Load the workbook with median age data
+wbage <- loadWorkbook("Auxiliary data/WPP2019_POP_F05_MEDIAN_AGE.xlsx")
+
+# Load sheet with SubSaharan countries
+medage <- data.table(readWorksheet(wbage, sheet = "ESTIMATES", startRow = 18, startCol = 2, header = F))
+
+# drop redundant columns
+medage[, c("Col1","Col3","Col5","Col6") := NULL]
+
+# set the proper names
+setnames(medage, c('country', 'ccode', 'x1950', 'x1955', 'x1960', 'x1965', 'x1970',
+                   'x1975','x1980','x1985','x1990','x1995','x2000','x2005',
+                   'x2010','x2015','x2020'))
+
+# assign proper iso3 codes
+medage[, iso3c := countrycode(ccode, "un", "iso3c")]
+
+# keep only the observations where iso3 codes are available (drop regions etc/)
+medage <- medage[!is.na(iso3c), ]
+medage[, c('ccode') := NULL]
+
+# transform into long format
+medage <- melt(medage, id.vars = c('country','iso3c'))
+medage[, year := as.numeric(str_replace(variable, "x", ""))]
+medage[, c('country','variable') := NULL]
+
+# only keep the years 1980+
+medage <- medage[year >= 1980, ]
+
+# write csv
+write.csv(medage, file = "Auxiliary data/to merge/median age.csv", row.names = F)

### 2019-11-28 Paper/paper.tex
Updates from Overleaf
191128 update

@@
 %!TEX program = xelatex
+%!BIB program = bibtex
 
@@
 \onehalfspacing
-\title{Finance and Wealth Inequality\thanks{We thank two anonymous referees, ...}}
+\title{Finance and Income Inequality - panel BMA approach\thanks{We thank two anonymous referees, ...}}
 \author[a]{Roman Horvath}
@@
 \citet{pikettyandzucman2014}
+\citet{van2018inequality} provide evidence that income inequality in the \ac{us} has different implications for the future income growth of the rich and the poor.
+High inequality seems to hurt the prospects of the poor while the top for the distribution is unaffected.
+The rich thus disproportianately benefit from higher inequality as their subsequent income exhibit faster growth.
+The authors attribute this effect to the political channel the rich use to lobby for implementation for the policies which back their economic interests.
+Preferences of the rich  are ultimately more likely to determine public policy than the preferences of the majority \citep{gilens_page_2014}.
+High inequality together with a credit constraint and rich driving the political process results in low government spending and lasting high inequality.
+
+\citet{marrero2013inequality} introduce two types of inequality - inequality of oppportunity and inequality of effort/luck.
+Applying this to the EU countries and US states, they show that inequality of opportunity (driven by race, parental education, citizenship, ...) is negatively related to growth while the residual "good" inequality is growth inducing.
+
 \section{Data}
@@
 
-\section{Concluding Remarks}
 % \label{sec:conclusion}

### 2019-11-28 Scripts/191101 Merge all data.R

@@ -35,7 +35,6 @@ icrg <- data.table(read.csv(file = "Auxiliary data/to merge/icrg.csv", header = 
 glob <- data.table(read.csv(file = "Auxiliary data/to merge/kof.csv", header = T, stringsAsFactors = F))
 medage <- data.table(read.csv(file = "Auxiliary data/to merge/median age.csv", header = T, stringsAsFactors = F))
 
-
 #
 #
 
@@ -96,6 +95,10 @@ data[is.na(LeftWing), LeftWing := 0]
 # checks
 # View(data)
 # View(data[duplicated(data[, j = .(iso3c,year)]),]) # should be NULL table
+
+# Add square of GDPpc and Inflation
+data[, GDPpc2 := GDPpc2^2]
+data[, Infl2 := Infl^2]
 
 # write final data
 write.csv(data, file = "data_hm.csv", row.names = F)

### 2019-11-28 Scripts/191108 analysis.R

@@ -146,7 +146,7 @@ data_try <- data_try[complete.cases(data_try), ]
 # demean the data
 data_dm <- data_try[, lapply(.SD[,2:ncol(.SD)], demean, na.rm = T), by = c("iso3c","country")]
 
-View(data_dm)
+# View(data_dm)
 
 # add back the year column
 data_dm[, period := data_try$period]

### 2019-12-03 Paper/paper.tex
update 191203

@@
 \usepackage{caption}
-\captionsetup[measuredfigure]{labelformat=empty}
+% \captionsetup[measuredfigure]{labelformat=empty}
 % \usepackage{bm}
@@
 \citet{marrero2013inequality} introduce two types of inequality - inequality of oppportunity and inequality of effort/luck.
-Applying this to the EU countries and US states, they show that inequality of opportunity (driven by race, parental education, citizenship, ...) is negatively related to growth while the residual "good" inequality is growth inducing.
+Applying this to the EU countries and US states, they show that inequality of opportunity (driven by race and parental education ...) is negatively related to growth while the residual "good" inequality is growth inducing.
 
+Income inequality and growth may intersect through varying channels.
+Accumulation of savings, unobservable effort, and investment project size favour the prediction on growth inducing inequality.
+Negative impact of inequality on human capital accumulation, entrepreneurial activity, and hence growth provides argument for the opposing view.
 \section{Data}

### 2019-12-03 Scripts/191101 Merge all data.R

@@ -7,10 +7,11 @@ library(countrycode)
 library(zoo)
 library(BMS)
 library(imputeTS)
+library(here)
 
 # Set the working directory
-wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
-setwd(wd)
+# wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
+# setwd(wd)
 
 # read all the underlying data
 wid <- data.table(read.csv(file = "Auxiliary data/to merge/wid data.csv", header = T, stringsAsFactors = F))
@@ -95,10 +96,6 @@ data[is.na(LeftWing), LeftWing := 0]
 # checks
 # View(data)
 # View(data[duplicated(data[, j = .(iso3c,year)]),]) # should be NULL table
-
-# Add square of GDPpc and Inflation
-data[, GDPpc2 := GDPpc2^2]
-data[, Infl2 := Infl^2]
 
 # write final data
 write.csv(data, file = "data_hm.csv", row.names = F)

### 2019-12-03 Scripts/191108 analysis.R

@@ -5,7 +5,7 @@
 library(data.table)
 library(countrycode)
 # library(zoo)
-library(BMS)
+library(dilutBMS2)
 library(imputeTS)
 library(here)
 library(dummies)
@@ -23,6 +23,9 @@ data <- data[!(iso3c %in% c("COM","STP","SYC","PSE")),]
 
 # adjust the variables
 data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
+data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
 
 #
 #
@@ -51,7 +54,8 @@ drops <- c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # 
 		   "MetalOre","ExpFuel", # correlated with NatRes
 		   "corr","burreau","demAcc","lawOrder", # ICRG data
 		   "GFCF", # duplicit with NonequipI + EquipI 
-		   "WarYears","RevCoups") # invariant in the sample
+		   "WarYears","RevCoups", # invariant in the sample
+		   "Age") # only available in 5-year intervals
 
 
 data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
@@ -66,11 +70,10 @@ data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","E
 		   # "WarYears","RevCoups", # invariant in the sample
 		   "GiniMarket","RedistRel", # alternative inequality indicators
 		   "GovSize","LegalSystem","SoundMoney","TradeFreedom","Regulation", # subcomponents of EFW
-		   "PrEdu","SecEdu","TerEdu") := NULL] 
+		   "PrEdu","SecEdu","TerEdu","Age") := NULL] 
 
 # how many observations are NAN for individual variables?
-sapply(data_try, function(x){sum(is.na(x))})
-
+# sapply(data_try, function(x){sum(is.na(x))})
 
 data_try <- data_try[complete.cases(data_try), ]
 
@@ -87,14 +90,14 @@ setcolorder(data_dm, c("GiniNet", names(data_dm[,names(data_dm)!="GiniNet"])))
 
 dum <- dummy(data_dm$year, sep = "_")
 data_dm <- cbind(data_dm, dum)
[…]

### 2019-12-03 Scripts/191121 median age.R

@@ -30,7 +30,7 @@ medage <- medage[!is.na(iso3c), ]
 medage[, c('ccode') := NULL]
 
 # transform into long format
-medage <- melt(medage, id.vars = c('country','iso3c'))
+medage <- melt(medage, id.vars = c('country','iso3c'), value.name = 'Age')
 medage[, year := as.numeric(str_replace(variable, "x", ""))]
 medage[, c('country','variable') := NULL]
 

### 2019-12-03 Scripts/191108 analysis.R
191203 update

@@ -1,4 +1,4 @@
-# Preliminary analysis
+# Preliminary analysis, market gini coefficient
 # Jan Mares, 191108
 
 # Libraries
@@ -39,7 +39,7 @@ data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, 
 data_f <- data[year %in% c(2000:2014),]
 
 # drop observations for which we do not have the depenedent variale
-data_f <- data_f[!is.na(Top10share), ]
+data_f <- data_f[!is.na(GiniNet), ]
 
 # interpolate rule of law (missing for 2001)
 data_f[, RuleofLaw := na.interpolation(RuleofLaw, option = "linear"), by = c("iso3c")]
@@ -96,14 +96,15 @@ data_dm <- data_dm[complete.cases(data_dm), ]
 dummies <- colnames(dum)[2:(ncol(dum))]
 
 # run BMA
-bma <- bms(data_dm, iter=500000, burn=100000, mprior = "uniform", g = "hyper",
+bma_1y_gini_baseline <- bms(data_dm, iter=5000000, burn=1000000, mprior = "uniform", g = "hyper",
                     nmodel=5000, mcmc="bd.int",
                     fixed.reg = dummies, user.int = F)
 
-coef(bma, exact = T)
-summary(bma)
+coef(bma_1y_gini_baseline, exact = T)
+summary(bma_1y_gini_baseline)
 
-View(data_dm)
+# save the results into file
+save(bma_1y_gini_baseline, file = here("Results/bma_1y_gini_baseline.Rdata"))
 
 #
 #
@@ -131,7 +132,7 @@ data_f[year %in% c(2012:2014), period := 5]
 data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
 data_f[, c('year') := NULL]
 
-data_f <- data_f[!is.nan(Top10share),]
+data_f <- data_f[!is.nan(GiniNet),]
 
 data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
 		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
@@ -169,12 +170,16 @@ data_dm[, c("period","period_1","WarYears","RevCoups","Top1share","Top10share") 
[…]

### 2019-12-03 Scripts/191203 analysis, top shares.R

@@ -1,0 +1,472 @@
+# Preliminary analysis, top10 and top1 shares
+# Jan Mares, 191203
+
+# Libraries
+library(data.table)
+library(countrycode)
+# library(zoo)
+library(dilutBMS2)
+library(imputeTS)
+library(here)
+library(dummies)
+
+# # Set the working directory
+# wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
+# setwd(wd)
+
+# read data file
+data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
+# View(data)
+
+# drop countries with unavailable data
+data <- data[!(iso3c %in% c("COM","STP","SYC","PSE")),]
+
+# adjust the variables
+data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
+data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
+
+#
+#
+
+# analysis
+
+#
+#
+
+# filter only data 2000+
+data_f <- data[year %in% c(2000:2014),]
+
+# drop observations for which we do not have the depenedent variale
+data_f <- data_f[!is.na(Top10share), ]
+
+# interpolate rule of law (missing for 2001)
+data_f[, RuleofLaw := na.interpolation(RuleofLaw, option = "linear"), by = c("iso3c")]
+
+# columns to drop, most missing values
+drops <- c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
+		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
+		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
+		   "FMA","FME", # unavailable for a lot of years
+		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
+		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
+		   "MetalOre","ExpFuel", # correlated with NatRes
+		   "corr","burreau","demAcc","lawOrder", # ICRG data
+		   "GFCF", # duplicit with NonequipI + EquipI 
+		   "WarYears","RevCoups", # invariant in the sample
[…]

### 2019-12-03 Scripts/Output tables to LaTeX.R

@@ -1,0 +1,52 @@
+# Output tables to LaTeX
+# Jan Mares, 191203
+
+# Libraries
+library(data.table)
+library(here)
+library(xtable)
+
+# Load the results
+load(file = here("Results/bma_3y_baseline.Rdata"))
+
+#
+#
+
+# Creating ouput tables
+
+# Variable names setup
+explvarshort <- c("NatRes","GovExp","GDSavings","PopGrowth","PopTot","Infl","EducExp",
+                  "VAI","VAA","NetFDIin","RuleofLaw","Mortality","GDPpc","TradeOpen",
+                  "LifeExp","Unemployment","EquipI","NonequipI","kaopen","FID","FIA",
+                  "FIE","FMD","EFW","EducIndex","CLandPR","LeftWing","RedistAbs","globFin",
+                  "globRestrict","globSoc","globPol","GDPpc#GDPpc","Infl#Infl",
+                  "EducIndex#EducIndex","period_2","period_3","period_4","period_5")
+
+explvarnames <- c("Natural resources rents","Government expenditures","Gross domestic savings",
+                  "Population growth","Total population","Inflation","Education expenditures",
+                  "Value added in industry","Value added in agriculture","Net FDI (% GDP)",
+                  "Rule of law","Mortality","GDP per capita", "Trade openness","Life expectancy",
+                  "Unemployment","Equipment investment","Non-equipment investment","Chinn-Ito index",
+                  "Financial institutions depth","Access to financial institutions","Financial institutions efficiency",
+                  "Financial markets depth","Economic freedom","Education index (UN)","Civil liberties & political rights",
+                  "Left-wing orientation","Redistribution","Financial globalization","Restrictions on globalization",
+                  "Social globalization","Political globalization","GDP per capita^2", "Inflation^2",
+                  "Education index^2","Period 2","Period 3","Period 4","Period 5")
+
+# Bind short and long variables names together
+varnames <- cbind(explvarshort, explvarnames)  
+
+# Results hyper prior
+results_all <- coef(bma_3y_baseline, exact=T)
+
+# List of variables sorted by PIP 
+varlist <- data.frame(explvarshort=row.names(results_all))
+
+# Creating concordance table
+concordance <- merge(varlist, varnames, by="explvarshort", sort=F)
+
+# Inputt variable names
+row.names(results_all) <- concordance$explvarnames
+
+# LaTeX output with proper varnames
+xtable(results_all[,1:3], digits=c(0,2,5,5))

### 2019-12-21 Paper/paper.tex
figures + 3y analysis

@@
 \begin{abstract}
-Using a global sample, this paper investigates the determinants of wealth inequality capturing various economic, financial, political, institutional, and geographical indicators.
-Using instrumental variable Bayesian model averaging, it reveals that only a handful of indicators robustly matters and finance plays a key role.
-It reports that while financial depth increases wealth inequality, efficiency and access to finance reduce inequality.
-In addition, redistribution and education are associated with lower inequality whereas wars and openness to international trade contribute to greater wealth inequality.
+    We investigate the impact of financial development on income inequality differentiating between depth, efficiency and access to financial markets and institutions.
+We apply panel Bayesian model averaging framework to address model uncertainty to reveal that financial development has complex influence on the income distribution within countries.
+The access to and efficiency of banking decrease income inequality.
+The size of the markets has no influence on overall income inequality, but contributes to the increasing top income shares.
+Moreover, unemployment along with investment into non-tangible assets increase income inequality while higher redistribution and physical capital investment imply lower levels of inequality.
 \end{abstract}
@@
 Negative impact of inequality on human capital accumulation, entrepreneurial activity, and hence growth provides argument for the opposing view.
+
+The average income inequality rose across \ac{oecd} by 1.4 percentage points \citep{oecd2013crisis}.
+
+\citet{LawSingh2014} argue that financial development decrease income inequality by allowing the poor to invest in human and physical capital.
+
+\subsection{theory}
+\citet{galormoav2004} provide a theoretical framework of thinking about the link between income inequality and financial development.
+In their model, the income inequality falls due to the decrease of financial friction which allows for physical and human capital accumulation of the poor.
+That is the extensive margin of financial development at work.
+
+\citet{banerjeenewman1990} also offer a theoretical setup in favour of the capital accumulation hypothesis due to finance (liberalization as claimed by \citet{de2017finance}) and inequality decreasing effect.
+
+\citet{GreenwoodJovanovic1990} on the contrary present a model where the intensive margin of the financial development is the key force and the benefits of more finance are accrued by the incumbents - primarilly the rich - which implies an inequality increasing effect.
+
+An exemption from focus solely on the size of financial sector is the study by \citet{naceurzhang2016} which applies a similar approach to ours, taking into consideration the access, efficiency, and stability of the financial sector.
+However, there are severe limitation to their study.
+First, their coverarage does not correspond to the availability of the data.
+Second, they do not account for the different dimensions of finance simultaneously and always use a single indicator of development at a time.
+Third, they do not explicitly differentiate between the banking sector and financial markets.
+We provide a more detailed as well as more robust picture of the financial development effect on finance in these dimensions.
+
+\citet{gwartney2017} provide an economic freedom index of the Fraser Institute.
+
+\citet{de2017finance} examine different dimensions of finance on income inequality.
+Their results suggest that financial development, financial liberalization, and banking crises all increase market income inequality within countries.
+Additionally, they show that the effect of financial liberalization is conditional on democratic accountability.
+The higher the accountability, the less severe is the negative impact of liberalization on inequality.
+On the contrary, the financial development, proxied by the credit to GDP ratio, has inequality increasing effect irrespective of the institutional background.
+
+\citet{claessens2007finance} explore the connectedness of growth, finance, and inequality, stressing the importance of institutional background in the effects.
+
+\citet{perottivolpin2004} present a theoretical framework how established interests lobby for lower investor protection to prevent entrance of the new competitors.
+Greater accountability of the politicians is reflected in higher bribe required from teh lobbyist therefore higher accountability implies higher market entry on more competition.
+After confirming their prediction empirically, they show that the most important factor of accountability is not the formal measure of democratic institutions, but \emph{newspaper readership} which they interpret as broad awareness of policy choices and their outcomes.
+
 \section{Data}
 % \label{sec:data}
+\begin{figure}
+    \caption{Gini Coefficient and Efficiency of Financial Institutions}
+    \label{fig:ginifie}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/FIEGiniNet}
+\end{figure}
 
+\begin{figure}
+    \caption{Gini Coefficient and Access to Financial Institutions}
+    \label{fig:ginifia}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/FIAGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Depth of Financial Institutions}
+    \label{fig:ginifid}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/FIDGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Depth of Financial Markets}
+    \label{fig:ginifmd}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/FMDGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Government Expenditures}
+    \label{fig:ginigovexp}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/GovExpGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Mortality}
+    \label{fig:ginimort}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/MortGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and \ac{gdp} per capita}
+    \label{fig:ginigdp}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/GDPGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Rule of Law}
+    \label{fig:ginirule}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/RuleLawGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Inflation}
+    \label{fig:giniinfl}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/InflGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Equipment Investment}
+    \label{fig:giniequipi}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/EquipIGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Trade Openness}
+    \label{fig:ginitrade}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/TradeOpenGiniNet}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Financial Development Indicators}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini}
+\end{figure}
+
+%
+%
+
+\begin{figure}
+    \caption{Gini Coefficient and Efficiency of Financial Institutions}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/FIEGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Access to Financial Institutions}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/FIAGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Depth of Financial Institutions}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/FIDGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Depth of Financial Markets}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/FMDGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Government Expenditures}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/GovExpGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Mortality}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/MortGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and \ac{gdp} per capita}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/GDPGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Rule of Law}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/RuleLawGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Inflation}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/InflGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Equipment Investment}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/EquipIGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Trade Openness}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/TradeOpenGiniNet_dm}
+\end{figure}
+
+\begin{figure}
+    \caption{Gini Coefficient and Financial Development Indicators}
+\includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini_dm}
+\end{figure}
+
+%
+%
 \section{Methodology}

### 2019-12-21 Scripts/191108 analysis.R

@@ -180,6 +180,18 @@ summary(bma_3y_gini_baseline)
 # save the results into file
 save(bma_3y_gini_baseline, file = here("Results/bma_3y_gini_baseline.Rdata"))
 
+#
+#
+
+bma_3y_gini_dilut <- bms(data_dm, iter=500000, burn=100000, mprior = "dilut",
+						  g = "hyper", nmodel=5000, mcmc="bd.int",
+                  		  fixed.reg = dummies, user.int = F)
+
+coef(bma_3y_gini_dilut, exact = T)
+summary(bma_3y_gini_dilut)
+
+# save the results into file
+save(bma_3y_gini_dilut, file = here("Results/bma_3y_gini_dilut.Rdata"))
 
 #
 #
@@ -214,7 +226,7 @@ data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","E
 		   # "WarYears","RevCoups", # invariant in the sample
 		   "GiniMarket","RedistRel", # alternative inequality indicators
 		   "GovSize","LegalSystem","SoundMoney","TradeFreedom","Regulation", # subcomponents of EFW
-		   "PrEdu","SecEdu","TerEdu","Age") := NULL]
+		   "PrEdu","SecEdu","TerEdu") := NULL]
 
 data_try <- data_try[complete.cases(data_try), ]
 
@@ -249,3 +261,16 @@ summary(bma_5y_gini_baseline)
 
 # save the results into file
 save(bma_5y_gini_baseline, file = here("Results/bma_5y_gini_baseline.Rdata"))
+
+#
+#
+
+bma_5y_gini_random <- bms(data_dm, iter=500000, burn=100000, mprior = "random", mprior.size = 5,
+				          g = "hyper=2", nmodel=5000, mcmc="bd.int",
+                    fixed.reg = dummies, user.int = F)
+
+coef(bma_5y_gini_random, exact = T, std.coefs = T)
+summary(bma_5y_gini_random)
+
+# save the results into file
+save(bma_5y_gini_random, file = here("Results/bma_5y_gini_random.Rdata"))

### 2019-12-21 Scripts/191203 analysis, top shares.R

@@ -169,7 +169,7 @@ data_dm[, c("period","period_1","WarYears","RevCoups","Top1share","GiniNet") := 
 
 dummies <- colnames(dum)[2:(ncol(dum))]
 
-# run BMA
+# run BMA, baseline
 bma_3y_top10_baseline <- bms(data_dm, iter=5000000, burn=1000000, mprior = "uniform", g = "hyper",
                     nmodel=5000, mcmc="bd.int",
                     fixed.reg = dummies, user.int = F)
@@ -180,6 +180,17 @@ summary(bma_3y_top10_baseline)
 # save the results into file
 save(bma_3y_top10_baseline, file = here("Results/bma_3y_top10_baseline.Rdata"))
 
+# run BMA, robustness
+bma_3y_top10_random <- bms(data_dm, iter=5000000, burn=1000000, mprior = "random", mprior.size = 7,
+						   g = "hyper", nmodel=5000, mcmc="bd.int",
+                    	   fixed.reg = dummies, user.int = F)
+
+coef(bma_3y_top10_random, exact = T)
+summary(bma_3y_top10_random)
+
+# save the results into file
+save(bma_3y_top10_random, file = here("Results/bma_3y_top10_random.Rdata"))
+plotModelsize(bma_3y_top10_random)
 
 #
 #
@@ -435,7 +446,7 @@ data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","E
 		   # "WarYears","RevCoups", # invariant in the sample
 		   "GiniMarket","RedistRel", # alternative inequality indicators
 		   "GovSize","LegalSystem","SoundMoney","TradeFreedom","Regulation", # subcomponents of EFW
-		   "PrEdu","SecEdu","TerEdu","Age") := NULL] 
+		   "PrEdu","SecEdu","TerEdu") := NULL] 
 
 data_try <- data_try[complete.cases(data_try), ]
 

### 2019-12-21 Scripts/191218 Figures.R

@@ -1,0 +1,580 @@
+# Figures for the paaper
+# Jan Mares, 191218
+
+# Libraries
+library(data.table)
+library(imputeTS)
+library(here)
+library(ggplot2)
+library(gridExtra)
+
+# read data file
+data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
+# View(data)
+
+# drop countries with unavailable data
+data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+
+# adjust the variables
+data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
+data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
+
+#
+#
+#   Averages - 3 year
+#
+#
+
+# Averages over years
+data_f <- data[year %in% c(2000:2014), ]
+data_f[year %in% c(2000:2002), period := 1]
+data_f[year %in% c(2003:2005), period := 2]
+data_f[year %in% c(2006:2008), period := 3] 
+data_f[year %in% c(2009:2011), period := 4]
+data_f[year %in% c(2012:2014), period := 5]
+
+# average observations by period and country
+data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
+data_f[, c('year') := NULL]
+
+data_f <- data_f[!is.nan(GiniNet),]
+
+data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
+		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
+		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
+		   "FMA","FME", # unavailable for a lot of years
+		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
+		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
+		   "ExpMetalOre","ExpFuel", # correlated with NatRes
[…]

### 2019-12-21 Scripts/191219 analysis baseline 3-year.R

@@ -1,0 +1,311 @@
+# Financial Development and Income Inequality
+# Baseline analysis, 3-y averages
+# Jan Mares, 1912019
+
+# Libraries
+library(data.table)
+library(countrycode)
+library(dilutBMS2)
+library(imputeTS)
+library(here)
+library(dummies)
+
+# read data file
+data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
+# View(data)
+
+# drop countries with unavailable data
+data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+
+# adjust the variables
+data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
+data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
+
+#
+#
+# Averages - 3 year, Gini index
+#
+#
+
+# Averages over years
+data_f <- data[year %in% c(2000:2014), ]
+data_f[year %in% c(2000:2002), period := 1]
+data_f[year %in% c(2003:2005), period := 2]
+data_f[year %in% c(2006:2008), period := 3] 
+data_f[year %in% c(2009:2011), period := 4]
+data_f[year %in% c(2012:2014), period := 5]
+
+# average observations by period and country
+data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
+data_f[, c('year') := NULL]
+
+data_f <- data_f[!is.nan(GiniNet),]
+
+data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
+		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
+		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
+		   "FMA","FME", # unavailable for a lot of years
+		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
+		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
+		   "ExpMetalOre","ExpFuel", # correlated with NatRes
[…]

### 2019-12-21 Scripts/Output tables to LaTeX.R

@@ -7,7 +7,16 @@ library(here)
 library(xtable)
 
 # Load the results
-load(file = here("Results/bma_3y_baseline.Rdata"))
+# load(file = here("Results/bma_1y_gini_baseline.Rdata"))
+# load(file = here("Results/bma_3y_gini_baseline.Rdata"))
+# load(file = here("Results/bma_5y_gini_baseline.Rdata"))
+# load(file = here("Results/bma_1y_top10_baseline.Rdata"))
+# load(file = here("Results/bma_3y_top10_baseline.Rdata"))
+# load(file = here("Results/bma_5y_top10_baseline.Rdata"))
+# load(file = here("Results/bma_1y_top1_baseline.Rdata"))
+# load(file = here("Results/bma_3y_top1_baseline.Rdata"))
+load(file = here("Results/bma_5y_top1_baseline.Rdata"))
+summary(bma_5y_top1_baseline)
 
 #
 #
@@ -20,7 +29,10 @@ explvarshort <- c("NatRes","GovExp","GDSavings","PopGrowth","PopTot","Infl","Edu
                   "LifeExp","Unemployment","EquipI","NonequipI","kaopen","FID","FIA",
                   "FIE","FMD","EFW","EducIndex","CLandPR","LeftWing","RedistAbs","globFin",
                   "globRestrict","globSoc","globPol","GDPpc#GDPpc","Infl#Infl",
-                  "EducIndex#EducIndex","period_2","period_3","period_4","period_5")
+                  "EducIndex#EducIndex","period_2","period_3","period_4","period_5",
+                  "year_2001","year_2002","year_2003","year_2004","year_2005","year_2006",
+                  "year_2007","year_2008","year_2009","year_2010","year_2011","year_2012",
+                  "year_2013","year_2014","Age")
 
 explvarnames <- c("Natural resources rents","Government expenditures","Gross domestic savings",
                   "Population growth","Total population","Inflation","Education expenditures",
@@ -31,13 +43,23 @@ explvarnames <- c("Natural resources rents","Government expenditures","Gross dom
                   "Financial markets depth","Economic freedom","Education index (UN)","Civil liberties & political rights",
                   "Left-wing orientation","Redistribution","Financial globalization","Restrictions on globalization",
                   "Social globalization","Political globalization","GDP per capita^2", "Inflation^2",
[…]

### 2019-12-21 Scripts/191030 WB data.R
update

@@ -37,6 +37,7 @@ setwd(wd)
 # SL.EMP.TOTL.SP.NE.ZS # 	Employment to population ratio, 15+, total (%) (national estimate)
 # SE.XPD.TOTL.GD.ZS # 		Public education expenditures as a share of GDP
 # NY.GDP.PCAP.PP.KD #		GDP per capita, PPP (constant 2011 international $)
+# NY.GDP.MKTP.KD.ZG			GDP, annual growth (%)
 # NE.TRD.GNFS.ZS #          Trade (% of GDP)
 # SP.DYN.LE00.IN # 			Life expectancy at birth
 # TX.VAL.MMTL.ZS.UN #		Ores and metals exports (% of merchandise exports)
@@ -67,10 +68,11 @@ setwd(wd)
 wbvars <- c("NY.GDP.TOTL.RT.ZS","NE.CON.GOVT.ZS","NY.GDS.TOTL.ZS","SP.POP.GROW","SP.POP.TOTL",
 			"FP.CPI.TOTL.ZG","NY.ADJ.AEDU.GN.ZS","NV.IND.TOTL.ZS","NV.AGR.TOTL.ZS",
 			"BX.KLT.DINV.WD.GD.ZS","BM.KLT.DINV.WD.GD.ZS","GC.TAX.TOTL.GD.ZS",
-			"NY.ADJ.NNAT.GN.ZS","RL.EST","SP.DYN.IMRT.IN","NE.GDI.FTOT.ZS",
+			"NY.ADJ.NNAT.GN.ZS","RL.EST","NE.GDI.FTOT.ZS",
 			"SL.EMP.TOTL.SP.NE.ZS","SE.XPD.TOTL.GD.ZS","NY.GDP.PCAP.PP.KD",
-			"NE.TRD.GNFS.ZS","SP.DYN.LE00.IN","TX.VAL.MMTL.ZS.UN","TX.VAL.FUEL.ZS.UN",
-			"SE.PRM.ENRR","SE.SEC.ENRR","SE.TER.ENRR","SL.UEM.TOTL.ZS")
+			"NY.GDP.MKTP.KD.ZG","NE.TRD.GNFS.ZS","SP.DYN.LE00.IN",
+			"TX.VAL.MMTL.ZS.UN","TX.VAL.FUEL.ZS.UN","SE.PRM.ENRR",
+			"SE.SEC.ENRR","SE.TER.ENRR","SL.UEM.TOTL.ZS")
 
[…]

### 2019-12-21 Scripts/191218 Figures Top10share.R

@@ -1,0 +1,580 @@
+# Figures for the paaper, top 10 share
+# Jan Mares, 191218
+
+# Libraries
+library(data.table)
+library(imputeTS)
+library(here)
+library(ggplot2)
+library(gridExtra)
+
+# read data file
+data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
+# View(data)
+
+# drop countries with unavailable data
+data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+
+# adjust the variables
+data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
+data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
+
+#
+#
+#   Averages - 3 year
+#
+#
+
+# Averages over years
+data_f <- data[year %in% c(2000:2014), ]
+data_f[year %in% c(2000:2002), period := 1]
+data_f[year %in% c(2003:2005), period := 2]
+data_f[year %in% c(2006:2008), period := 3] 
+data_f[year %in% c(2009:2011), period := 4]
+data_f[year %in% c(2012:2014), period := 5]
+
+# average observations by period and country
+data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
+data_f[, c('year') := NULL]
+
+data_f <- data_f[!is.nan(Top10shareshare),]
+
+data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
+		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
+		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
+		   "FMA","FME", # unavailable for a lot of years
+		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
+		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
+		   "ExpMetalOre","ExpFuel", # correlated with NatRes
[…]

### 2019-12-21 Scripts/191218 Figures Top1share.R

@@ -1,0 +1,580 @@
+# Figures for the paaper, top 10 share
+# Jan Mares, 191218
+
+# Libraries
+library(data.table)
+library(imputeTS)
+library(here)
+library(ggplot2)
+library(gridExtra)
+
+# read data file
+data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
+# View(data)
+
+# drop countries with unavailable data
+data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+
+# adjust the variables
+data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
+data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
+
+#
+#
+#   Averages - 3 year
+#
+#
+
+# Averages over years
+data_f <- data[year %in% c(2000:2014), ]
+data_f[year %in% c(2000:2002), period := 1]
+data_f[year %in% c(2003:2005), period := 2]
+data_f[year %in% c(2006:2008), period := 3] 
+data_f[year %in% c(2009:2011), period := 4]
+data_f[year %in% c(2012:2014), period := 5]
+
+# average observations by period and country
+data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
+data_f[, c('year') := NULL]
+
+data_f <- data_f[!is.nan(Top1share0share),]
+
+data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
+		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
+		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
+		   "FMA","FME", # unavailable for a lot of years
+		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
+		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
+		   "ExpMetalOre","ExpFuel", # correlated with NatRes
[…]

### 2019-12-21 Scripts/191218 Figures.R

@@ -1,580 +1,0 @@
-# Figures for the paaper
-# Jan Mares, 191218
-
-# Libraries
-library(data.table)
-library(imputeTS)
-library(here)
-library(ggplot2)
-library(gridExtra)
-
-# read data file
-data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
-# View(data)
-
-# drop countries with unavailable data
-data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
-
-# adjust the variables
-data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
-
-# Add square of GDPpc and Inflation
-data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
-
-#
-#
-#   Averages - 3 year
-#
-#
-
-# Averages over years
-data_f <- data[year %in% c(2000:2014), ]
-data_f[year %in% c(2000:2002), period := 1]
-data_f[year %in% c(2003:2005), period := 2]
-data_f[year %in% c(2006:2008), period := 3] 
-data_f[year %in% c(2009:2011), period := 4]
-data_f[year %in% c(2012:2014), period := 5]
-
-# average observations by period and country
-data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
-data_f[, c('year') := NULL]
-
-data_f <- data_f[!is.nan(GiniNet),]
-
-data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
-		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
-		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
-		   "FMA","FME", # unavailable for a lot of years
-		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
-		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
-		   "ExpMetalOre","ExpFuel", # correlated with NatRes
[…]

### 2019-12-21 Scripts/191219 analysis baseline 3-year.R

@@ -108,7 +108,7 @@ save(bma_3y_gini_dilut, file = here("Results/bma_3y_gini_dilut.Rdata"))
 #
 
 # random model prior (accounting for a number of regressors in prior model probability)
-bma_3y_gini_random <- bms(data_dm, iter=5000000, burn=1000000, mprior = "random",
+bma_3y_gini_random <- bms(data_dm, iter=5000000, burn=1000000, mprior = "random", mprior.size = 5,
 						  g = "hyper", nmodel=5000, mcmc="bd.int",
                   		  fixed.reg = dummies, user.int = F)
 

### 2019-12-21 Scripts/191221 Figures robustness.R

@@ -1,0 +1,53 @@
+#  Figures for the paper, top 10 share
+# Jan Mares, 191221
+
+# Libraries
+library(data.table)
+library(dilutBMS2)
+library(here)
+library(ggplot2)
+
+# Gini 
+
+# load the files with results
+load(here("Results/bma_3y_gini_dilut.Rdata"))
+load(here("Results/bma_3y_gini_baseline.Rdata"))
+load(here("Results/bma_3y_gini_random.Rdata"))
+
+pl_comp_gini <- plotComp(bma_3y_gini_baseline, bma_3y_gini_dilut, bma_3y_gini_random,
+                         lwd=1, varNr = NULL, comp = "PIP", exact = T, include.legend = TRUE,
+                         add.grid = F, do.par = T, cex.axis=0.8, cex.xaxis=0.65,
+                         col=c(rgb(0.03, 0.15, 0.4),rgb(0.753,0,0),rgb(0, 0, 0)))
+
+cairo_ps(file = paste0(folderfig,"model_priors_comparison_gini.eps"), width=9, height=6, family="helvetica")
+pl_comp_gini
+dev.off()
+
+# Top 10 share
+load(here("Results/bma_3y_top10_dilut.Rdata"))
+load(here("Results/bma_3y_top10_baseline.Rdata"))
+load(here("Results/bma_3y_top10_random.Rdata"))
+
+pl_comp_top10 <- plotComp(bma_3y_top10_baseline, bma_3y_top10_dilut, bma_3y_top10_random,
+                          lwd=1, varNr = NULL, comp = "PIP", exact = T, include.legend = TRUE,
+                          add.grid = F, do.par = T, cex.axis=0.8, cex.xaxis=0.65,
+                          col=c(rgb(0.03, 0.15, 0.4),rgb(0.753,0,0),rgb(0, 0, 0)))
+
+cairo_ps(file = paste0(folderfig,"model_priors_comparison_top10.eps"), width=9, height=6, family="helvetica")
+pl_comp_top10
+dev.off()
+
+
+# Top 1 share
+load(here("Results/bma_3y_top1_dilut.Rdata"))
+load(here("Results/bma_3y_top1_baseline.Rdata"))
+load(here("Results/bma_3y_top1_random.Rdata"))
+
+pl_comp_top1 <- plotComp(bma_3y_top1_baseline, bma_3y_top1_dilut, bma_3y_top1_random,
[…]

### 2019-12-30 Paper/paper.tex
update literature

@@
 \acrodef{pmp}[PMP]{Posterior Model Probability}
+\acrodef{wals}[WALS]{Weighted-average Least Squares}
 
@@
 
+\citet{nolan2019drivers} bring a survey of the literature on determinants of inequality, summarizing the complexity of the inequality dynamics.
+They stress that many of the determinants are interlinked which implies difficulty in assigning precise effects to individual drivers of inequality.
+Additionally, they encourage complementary individual country case studies to support the finding of the general cross-country estimates.
+
 \subsection{theory}
@@
 
-\citet{perottivolpin2004} present a theoretical framework how established interests lobby for lower investor protection to prevent entrance of the new competitors.
+\citet{perotti2007investor} present a theoretical framework how established interests lobby for lower investor protection to prevent entrance of the new competitors.
 Greater accountability of the politicians is reflected in higher bribe required from teh lobbyist therefore higher accountability implies higher market entry on more competition.
 After confirming their prediction empirically, they show that the most important factor of accountability is not the formal measure of democratic institutions, but \emph{newspaper readership} which they interpret as broad awareness of policy choices and their outcomes.
+
+Summary of the theoretical preditions in \citet{demirgucc2009finance} on pages 20-23 for the review.
+
+\citet{furceri2019robust} apply \ac{wals} to identify robust determinants of income inequality.
+They are the closest paper to ours since they account for model uncertainty in the estimation.
+Their focus if more general rathar than on finance specifically.
+Out work offers more detailed contribution in terms of the role of finance in shaping income inequality.
+On the top of that, we differ from their analysis by examining multiple measures of inequality and specifically identifying the determinants of top income shares along with the determinants of the overall income distribution represented by income Gini index.
+We provide further evidence in all these dimensions.
 

### 2019-12-30 Scripts/191219 analysis baseline 3-year.R

@@ -5,6 +5,7 @@
 # Libraries
 library(data.table)
 library(countrycode)
+library(BMS)
 library(dilutBMS2)
 library(imputeTS)
 library(here)
@@ -94,7 +95,7 @@ save(bma_3y_gini_baseline, file = here("Results/bma_3y_gini_baseline.Rdata"))
 #
 
 # dilution model prior (accounting for corrlation between regressors)
-bma_3y_gini_dilut <- bms(data_dm, iter=5000000, burn=1000000, mprior = "dilut",
+bma_3y_gini_dilut <- dilutBMS2::bms(data_dm, iter=5000000, burn=1000000, mprior = "dilut",
 						  g = "hyper", nmodel=5000, mcmc="bd.int",
                   		  fixed.reg = dummies, user.int = F)
 
@@ -190,7 +191,7 @@ save(bma_3y_top10_baseline, file = here("Results/bma_3y_top10_baseline.Rdata"))
 #
 
 # dilution model prior (accounting for corrlation between regressors)
-bma_3y_top10_dilut <- bms(data_dm, iter=5000000, burn=1000000, mprior = "dilut",
+bma_3y_top10_dilut <- dilutBMS2::bms(data_dm, iter=5000000, burn=1000000, mprior = "dilut",
 						  g = "hyper", nmodel=5000, mcmc="bd.int",
                   		  fixed.reg = dummies, user.int = F)
 
@@ -286,7 +287,7 @@ save(bma_3y_top1_baseline, file = here("Results/bma_3y_top1_baseline.Rdata"))
 #
 
 # dilution model prior (accounting for corrlation between regressors)
-bma_3y_top1_dilut <- bms(data_dm, iter=5000000, burn=1000000, mprior = "dilut",
+bma_3y_top1_dilut <- dilutBMS2::bms(data_dm, iter=5000000, burn=1000000, mprior = "dilut",
 						  g = "hyper", nmodel=5000, mcmc="bd.int",
                   		  fixed.reg = dummies, user.int = F)
 

### 2020-01-03 Paper/paper.tex
update

@@
 % \label{sec:literature}
+
 \citet{pikettyandzucman2014}
 \citet{van2018inequality} provide evidence that income inequality in the \ac{us} has different implications for the future income growth of the rich and the poor.
-High inequality seems to hurt the prospects of the poor while the top for the distribution is unaffected.
+High inequality seems to hurt the prospects of the poor while the top of the distribution is unaffected.
 The rich thus disproportianately benefit from higher inequality as their subsequent income exhibit faster growth.
-The authors attribute this effect to the political channel the rich use to lobby for implementation for the policies which back their economic interests.
+The authors attribute this effect to the political channel the rich use to lobby in favour of the policies which support their economic interests.
 Preferences of the rich are ultimately more likely to determine public policy than the preferences of the majority \citep{gilens_page_2014}.
-High inequality together with a credit constraint and rich driving the political process results in low government spending and lasting high inequality.
+High inequality together with a credit constraint and rich driving the political process results in low government spending and lasting inequality.
 
-\citet{marrero2013inequality} introduce two types of inequality - inequality of oppportunity and inequality of effort/luck.
+\citet{marrero2013inequality} introduce two types of inequality - inequality of opportunity and inequality of effort/luck.
 Applying this to the EU countries and US states, they show that inequality of opportunity (driven by race and parental education ...) is negatively related to growth while the residual "good" inequality is growth inducing.

### 2020-01-03 Scripts/191218 Figures GiniNet.R

@@ -13,7 +13,7 @@ data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFac
 # View(data)
 
 # drop countries with unavailable data
-data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+data <- data[!(iso3c %in% c("BEN","COM","GIN","LBR","OMN","PSE","STP","SYC","SWZ")),]
 
 # adjust the variables
 data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
@@ -51,7 +51,8 @@ data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","E
 		   "corr","burreau","demAcc","lawOrder", # ICRG data
 		   "GFCF", # duplicit with NonequipI + EquipI ) := NULL]
 		   # "WarYears","RevCoups", # invariant in the sample
-		   "GiniMarket","RedistRel", # alternative inequality indicators
+		   "RedistRel", # alternative inequality indicators
+		#  "GiniMarket","RedistRel", # alternative inequality indicators
 		   "GovSize","LegalSystem","SoundMoney","TradeFreedom","Regulation", # subcomponents of EFW
 		   "PrEdu","SecEdu","TerEdu","Age") := NULL] 
 
@@ -80,6 +81,37 @@ data_try <- data_try[complete.cases(data_try), ]
 # Set up color scheme
 colset <- matrix(c(192,0,0,
                    55,96,146), ncol=3, byrow=T)
+
+# Top10 and top1 share to GiniNet
+pl1 <- ggplot(data_try, aes(Top10share, GiniNet)) +
+		      ylab("Gini index - income (after redistribution)") +
+		      xlab('Top 10% income share')
+		      
+pl1g <- pl1 + geom_point(size=2) +
+		theme_bw() +
+		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
+			  panel.border = element_rect(fill = NA, colour = 'black'),
+    		  legend.position="none") +
+        stat_smooth(method='lm', se=F) +
+        scale_colour_manual(values=rgb(colset/255)) + 
+        scale_fill_manual(values=rgb(colset/255))
+
+pl1g
+
+pl2 <- ggplot(data_try, aes(Top1share, GiniNet)) +
+		      ylab("Gini index - income (after redistribution)") +
[…]

### 2020-01-03 Scripts/191219 analysis baseline 3-year.R

@@ -10,6 +10,22 @@ library(dilutBMS2)
 library(imputeTS)
 library(here)
 library(dummies)
+library(ineq)
+
+
+
+lines(p,Lc.lognorm(p, parameter=0.5),col="red")
+lines(p,Lc.lognorm(p, parameter=1.0),col="blue")
+lines(p,Lc.lognorm(p, parameter=1.5),col="green")
+
+gini = 1/((2*aplha+0.45)-1)
+
+aplha
+s = sqrt(2)*qnorm((0.56+1)/2)
+alpha = 1/(2*0.56)+1/2
+p = seq(0,1,0.01)
+plot(p,Lc.lognorm(p, parameter=s), type="l", col="brown")
+lines(p,Lc.pareto(p, parameter=alpha), col="blue")
 
 # read data file
 data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
@@ -53,8 +69,9 @@ data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","E
 		   "ExpMetalOre","ExpFuel", # correlated with NatRes
 		   "corr","burreau","demAcc","lawOrder", # ICRG data
 		   "GFCF", # duplicit with NonequipI + EquipI ) := NULL]
-		   # "WarYears","RevCoups", # invariant in the sample
+		#  "WarYears","RevCoups", # invariant in the sample
 		   "GiniMarket","RedistRel", # alternative inequality indicators
+		#    "GiniNet","RedistRel", # alternative inequality indicators
 		   "GovSize","LegalSystem","SoundMoney","TradeFreedom","Regulation", # subcomponents of EFW
 		   "PrEdu","SecEdu","TerEdu","Age") := NULL] 
 
@@ -71,6 +88,7 @@ data_dm[, period := data_try$period]
 data_dm[, c("iso3c","country") := NULL]
 
 # order variables, the dependent must be first
+# setcolorder(data_dm, c("GiniNet", names(data_dm[,names(data_dm)!="GiniNet"])))
 setcolorder(data_dm, c("GiniNet", names(data_dm[,names(data_dm)!="GiniNet"])))
 
 
@@ -81,7 +99,7 @@ data_dm[, c("period","period_1","WarYears","RevCoups","Top1share","Top10share") 
 dummies <- colnames(dum)[2:(ncol(dum))]
 
 # run BMA
[…]

### 2020-01-03 Scripts/200103 Theoretical Gini indices.r

@@ -1,0 +1,25 @@
+# Theoretical Gini indices
+# Jan Mares, 200103
+
+# Libraries
+library(data.table)
+library(ineq)
+
+# define the inputs
+p = seq(0,1,0.01)
+gini = 0.56
+
+# lines(p,Lc.lognorm(p, parameter=0.5),col="red")
+# lines(p,Lc.lognorm(p, parameter=1.0),col="blue")
+# lines(p,Lc.lognorm(p, parameter=1.5),col="green")
+
+# define the variables
+s = sqrt(2)*qnorm((gini+1)/2) # for log-normal distribution
+alpha = 1/(2*gini)+1/2 # for pareto distribution
+
+# plot
+plot(p,Lc.lognorm(p, parameter=s), type="l", col="brown")
+lines(p,Lc.pareto(p, parameter=alpha), col="blue")
+
+# compute the gini with specific alpha coefficient in Pareto distribution
+gini = 1/((2*aplha+0.45)-1)

### 2020-01-05 Paper/paper.tex
update

@@
 % \label{sec:intro}
+Finance captures the capacity of financial intermediaries and markets to screen investment opportunities, monitor the debtors who were provided funding, as well as pooling and management of risk.
+With inequality, we focus in the paper on the inequality in the distribution of income.
+Arguably, the literature discusses other concepts of inequality, e.g.
+intergenerational persistence of relative income differences or equality of opportunity \citep{demirgucc2009finance}.
 
@@
 % \label{sec:literature}
+The research in the area of financial development and income inequality is well established.
+\citet{demirgucc2009finance}, \citet{claessens2007finance}, and more recently \cite{de2017finance} provide extensive reviews of the topic.
+The similar theme emerges in all these papers, with the conclusions from both theoretical and empirical contributions provide conflicting prediction about the relationship and well as not unambiguously conclusive empirical results.
+Tentative conclusions point towards finance tightening the distribution of income.
+Among other channels, accessing credit allows human capital accumulation and new firm formation and growth equalize economic opportunities.
+Furthermore, the indirect effect of finance on labour markets through faster economic growth and increased competition diproportionately benefits the low-income individuals.
+
+Conflicting prediction of finance on inequality - intensive vs.
+extensive margin.
+Intergenerational persistence might fall with increased economic oportunities while widening the income distribution in every generation.
+
+DIRECT vs.
+INDIRECT mechanism - access to finance, expanding opportunities of the previsously disadvantaged groups or intensifying the use of financial services by the incumbent individuals and firms.
+The financial system having second-order effects on inequality by influencing aggregate production and allocation of credit and consequantly the demand for high- vs.
+low- skilled workers.
+\citet{braunetal2019} link growth, inequality, and finance in their theoretical model (access to markets inhibits growth).
+
+\citet{delis2014} bank regulations and inequality - nice intro
+
+% Financial development alters how much are economic opportunities depend on the individual skills, family endowments, social status or political connections.
+Individual depend on financial system to provide loans to start new business, attain education
+
+Both theoretical and empirical studies leave out the issue of importing the financial services from abroad. We include financial globalization from KOF among our control variables.
 
@@
 \subsection{theory}
+\textsc{human capital investment} - with imperfect markets restrict the capital accumulation, differences between human (optimal is equal distribution) and physical capital (diminishin returns to physical capital do not materialize at the level of individual ownership of capital because it is not embodies in individuals \cite{demirgucc2009finance}.
+This implies that equal distribution is not necessarily optimal).
+\textsc{political consequences} - higher inequality will increase the proportion of population favouring greater redistribution.
+\textsc{wage discrimination by specific characteristics, e.
+g.
+race}
+%
+\textsc{investment opportunities, savings}.
+Minimum investment requirements of fixed costs associated with high-return investments, including entrepreneurship.
+=> higher return to wealthier individuals (e.g.
+\citet{AtjeJovanovich1993,demel2008qje}).
+\textsc{savings behaviour} - if inheritance is a function on income, then rich parents leave larger bequests than the poor and preserve relative differences in incomes.
+Research indicates that financial development does not alter aggregate savings rates.
+Is this also valid for more detailed look on poor and rich households?
+If the effect differs then the distribution of saving is altered by financial development influence the relative income levels across households while keeping the aggregate savings rate unchanged.
+
+\textsc{Summary}
+
 \citet{galormoav2004} provide a theoretical framework of thinking about the link between income inequality and financial development.
@@
 
-\citet{GreenwoodJovanovic1990} on the contrary present a model where the intensive margin of the financial development is the key force and the benefits of more finance are accrued by the incumbents - primarilly the rich - which implies an inequality increasing effect.
+\citet{GreenwoodJovanovic1990} on the contrary present a model where the intensive margin of the financial development is the key force and the benefits of more finance are accrued by the incumbents - primarily the rich - which implies an inequality increasing effect in the early stages of the development.
+With tim more agents meet the fixed costs of joining the financial intermediaries and they enjoy higher returns.
+The efficiency of resource allocation increases, which enhances growth and reduces inequality.
 
@@
 \citet{perotti2007investor} present a theoretical framework how established interests lobby for lower investor protection to prevent entrance of the new competitors.
-Greater accountability of the politicians is reflected in higher bribe required from teh lobbyist therefore higher accountability implies higher market entry on more competition.
-After confirming their prediction empirically, they show that the most important factor of accountability is not the formal measure of democratic institutions, but \emph{newspaper readership} which they interpret as broad awareness of policy choices and their outcomes.
+Greater accountability of the politicians leads to a higher bribe required from the lobbyists'.
+Investor protection strengthens and increases market entry and competition.
+They examine their prediction in a cross-section and show that better investor protection correlates with larger entry rates and higher firm density in more financially intensive sectors.
+Furthermore, they show that the most important factor of accountability is not the formal measure of democratic institutions, but \emph{newspaper readership} which they interpret as broad awareness of policy choices and their outcomes.
+\textsc{size does not matter for entry when investor protection is accounted for}.
 
@@
 
+\subsection{empirics}
+\citet{beck2007finance} show that the development of financial intermediaries is accompanied by faster reductions in income inquality and poverty rates.
+\citet{beck2010big} document how bank deregulation in the \ac{us} tightened income distribution by benefiting primarily the bottom 50\% of earners.
+They attribute the effect to the changes in labour market conditions and relatively higher wages and working hours of the low-skilled workers following the reforms.
+
+\citet{claessens2007finance} argue that althugh the deeper financial systems generally provide better opportunities of access to finance, the relationship is not universal.
+
+The benefits of capital markets liberalization seem to be concetrated to the top of the income distribution.
+Top quintile of the distribution accrues nearly all of the income growth following the liberalization while the share of middle three quantiles decreases and the bottom remains unaffected \citep{das2003income}.
+
+Diwan (1999) show how labour shares fall follwing a crises only partially recovering in the subsequent years.
+
+\citet{kroszneretal2007} show that financial crises have relative more severe impact on the sectors which depend more on external financing.
+The consequences of crises on firms relate to institutional environment and materialize through lower production capacity and competition.
+
+\citet{glaeseretal2003} model intitutional subversion by the rich classes and their ability to prevent institutional reforms.
+Such environment often leads to massive redistribution from the rich to the poor, which is assumed to be less efficient than carefully implemented institutional reforms.
+
 \section{Data}
+\cite{roineetal2009}, \citet{nolan2019drivers}, \citet{furceri2019robust} for the determinants of income inequality.
+
 % \label{sec:data}

### 2020-01-08 Paper/paper.tex
update

--- 2020-01-03 Paper/paper.tex
@@
-%!TEX program = xelatex
+%!TEX program = pdflatex
 %!BIB program = bibtex
@@
 % \usepackage{subcaption}
-%\usepackage{epstopdf} % not supported by xelatex
+\usepackage{epstopdf} % not supported by xelatex
 % \usepackage{fullpage}
@@
 \usepackage{authblk}
-\usepackage{fontspec}
+\usepackage[T1]{fontenc}
+\usepackage{inputenc}
+\usepackage{lmodern}
+% \usepackage{fontspec}
 \usepackage{lscape}
@@
 % \label{sec:intro}
+% Financial development alters how much are economic opportunities depend on the individual skills, family endowments, social status or political connections.
+Individual depend on financial system to provide loans to start new business, attain education
+
+Finance captures the capacity of financial intermediaries and markets to screen investment opportunities, monitor the debtors who were provided funding, as well as pooling and management of risk.
+With inequality, we focus in the paper on the inequality in the distribution of income.
+Arguably, the literature discusses other concepts of inequality, e.g.
+intergenerational persistence of relative income differences or equality of opportunity \citep{demirgucc2009finance}.
+
+\citet{claessens2007finance} argue that although deeper financial systems generally provide better opportunities of access to finance, the relationship is not universal.
+
+The average income inequality rose across \ac{oecd} by 1.4 percentage points \citep{oecd2013crisis}.
 
@@
 % \label{sec:literature}
+The research in the area of financial development and income inequality is well established.
+\citet{demirgucc2009finance}, \citet{claessens2007finance}, and more recently \cite{de2017finance} provide extensive reviews of the topic.
+A similar theme emerges in all three papers.
+The implications from theoretical contributions provide conflicting predictions about the relationship and empirical results bring evidence for both positive and negative effect.
+Although majority of the papers point towards finance tightening the distribution of income this results is not universal with some papers suggesting the opposite while other stress potential non-linearities.
 
-\citet{pikettyandzucman2014}
-\citet{van2018inequality} provide evidence that income inequality in the \ac{us} has different implications for the future income growth of the rich and the poor.
+A key divide appears between the effect of financial develpoment on extenstive and intensive margin.
+The extensive margin captures the extend to which individuals, who had not been using financial services before, gain access.
+On the other hand, the intensive margin describes growing use of finance by the agents who had already been using it before \citep{demirgucc2009finance}.
+Financial development on the extensive margin might lead to more equal opportunities and outcomes.
+Access to credit by previously disadvantaged groups allows human capital accumulation \citep{galorzeira1993income, galormoav2004, braunetal2019}, formation and growth of new firms \citep{evans1989estimated,banerjeenewman1990}, with more evenly distributed economic opportunities as a result\footnote{Having similar economic opportunities might decrease the cross-generational inequality, by diminishing the effect of e.g.
+parental wealth.
+Depending on the innate abilities and talents of the individuals, however, it may increase the inequality of income within every generation at the same time.}.
+
+On the contrary, intensive margin of financial development might disproportionately benefit the rich who may leverage financial services for their further benefit or to protect their existing rents.
+\citet{GreenwoodJovanovic1990} present a model where the finance is the key driver of inequality and the welfare gains accrued by the incumbents - primarily the rich - in the initial development stage.
+With time, more agents meet the fixed costs of joining the financial intermediaries and they enjoy higher returns.
+Consequently, the efficiency of resource allocation also increases, which enhances growth and reduces inequality.
+\citet{perotti2007investor} present a framework based on politcal economy.
+They argument depends on a lobby for lower investor protection to prevent entrance of the new competitors.
+The politicians require higher bribe from the lobbyist the greater is their accountability for policy decisions.
+Thus, with increasing accountability, investor protection strengthens and spurs market entry and competition.
+The authors examine their prediction in a cross-section and show that better investor protection correlates with larger entry rates and higher firm density in more financially intensive sectors\footnote{In addition, they show that the most important factor of accountability is not the formal measure of democratic institutions, but \emph{newspaper readership} which they interpret as broad awareness of policy choices and their outcomes.}.
+
+Financial development may also have indirect effect on income inequality through economic growth.
+\citet{townsendeueda2006} model how finance interacts with production and allocation of credit.
+If increased use of finance increases the demand for low- relatively to the high-skilled workers, then it may have equalizing consequences for income distribution.
+Empirical evidence by \citet{beck2010big} show that bank deregulation and increased competition in loan provision in the \ac{us} primarily benefited the workers with income below the median.
+Similarly, \citet{delis2014} provide evidence of bank deregulation and liberalization tightening the income distribution, although this effect is only present in countries with high-quality institutions.
+They attribute the effect to the changes in labour market conditions and relatively higher wages and working hours of the low-skilled workers following the reforms.
+
+A set of distinct papers explores the relationship between inequality and growth while stressing financial markest imperfections driving the outcomes.
+Income inequality and growth may intersect through varying channels.
+Accumulation of savings, unobservable effort, and investment project size favour the prediction of growth inducing inequality.
+Negative impact of inequality on human capital accumulation, entrepreneurial activity provide argument for the opposing view.
+\citet{van2018inequality} report how income inequality in the \ac{us} has different implications for the future income growth of the rich and the poor.
 High inequality seems to hurt the prospects of the poor while the top of the distribution is unaffected.
@@
 
-\citet{marrero2013inequality} introduce two types of inequality - inequality of opportunity and inequality of effort/luck.
-Applying this to the EU countries and US states, they show that inequality of opportunity (driven by race and parental education ...) is negatively related to growth while the residual "good" inequality is growth inducing.
+The literature does not converge on the conclusions even in the empirical cross-country and panel data studies.
+The papers link higher levels of financial development with lower levels of inequality \citep{beck2007finance, hamori2012, gimet2011closer, kunieda2014finance}\footnote{For an extended list, we refer to \citet{de2017finance}.}.
+On the other hand, several other estimate a growth inducing effect of finance \citep{Jaumotte2013, jauch2016financial, de2017finance}.
+Finally, some authors claim there the relationship might be non-linear, conditional on a threshold value of financial development \citep{kim2011nonlinearity,tan2012nonlinear} or institutional quality \citep{LawSingh2014, delis2014}.
 
-Income inequality and growth may intersect through varying channels.
-Accumulation of savings, unobservable effort, and investment project size favour the prediction on growth inducing inequality.
-Negative impact of inequality on human capital accumulation, entrepreneurial activity, and hence growth provides argument for the opposing view.
+Three papers are the closest to ours, each in a different respect.
+First, \citet{de2017finance} examine different dimensions of finance on income inequality.
+Their results suggest that financial development, financial liberalization, and banking crises all increase pre-tax income inequality within countries.
+Additionally, they show that the effect of financial liberalization is conditional on democratic accountability.
+Higher accountability mitigates the impact of liberalization on inequality.
+On the contrary, the financial development, proxied by the credit to GDP ratio, has inequality increasing effect irrespective of the institutional background.
+Second, \citet{naceurzhang2016} take similar approach in considering multiple dimensions --- the access, efficiency, and stability of the financial sector, although not examining the indicators simultaneously.
+Third, \citet{furceri2019robust} apply \ac{wals} to identify robust determinants of income inequality.
+Their approach mirrors ours in accounting for model uncertainty in the estimation.
+Their focus is more general rather than focused primarily on finance.
+We provide synthesis and extension to these papers in providing more detailed view on the link between finance in shaping income inequality and examining multiple measures of inequality while specifically identifying the determinants of top income shares along with the determinants of the overall income distribution\footnote{Captured by income Gini index}.
 
-The average income inequality rose across \ac{oecd} by 1.4 percentage points \citep{oecd2013crisis}.
-
-\citet{LawSingh2014} argue that financial development decrease income inequality by allowing the poor to invest in human and physical capital.
-
-\citet{nolan2019drivers} bring a survey of the literature on determinants of inequality, summarizing the complexity of the inequality dynamics.
-They stress that many of the determinants are interlinked which implies difficulty in assigning precise effects to individual drivers of inequality.
-Additionally, they encourage complementary individual country case studies to support the finding of the general cross-country estimates.
-
-\subsection{theory}
-\citet{galormoav2004} provide a theoretical framework of thinking about the link between income inequality and financial development.
-In their model, the income inequality falls due to the decrease of financial friction which allows for physical and human capital accumulation of the poor.
-That is the extensive margin of financial development at work.
-
-\citet{banerjeenewman1990} also offer a theoretical setup in favour of the capital accumulation hypothesis due to finance (liberalization as claimed by \citet{de2017finance}) and inequality decreasing effect.
-
-\citet{GreenwoodJovanovic1990} on the contrary present a model where the intensive margin of the financial development is the key force and the benefits of more finance are accrued by the incumbents - primarilly the rich - which implies an inequality increasing effect.
-
-An exemption from focus solely on the size of financial sector is the study by \citet{naceurzhang2016} which applies a similar approach to ours, taking into consideration the access, efficiency, and stability of the financial sector.
+% An exemption from focus solely on the size of financial sector is the study by \citet{naceurzhang2016} which applies a similar approach to ours, taking into consideration the access, efficiency, and stability of the financial sector.
 However, there are severe limitation to their study.
@@
 
-\citet{gwartney2017} provide an economic freedom index of the Fraser Institute.
-
-\citet{de2017finance} examine different dimensions of finance on income inequality.
+% \citet{de2017finance} examine different dimensions of finance on income inequality.
 Their results suggest that financial development, financial liberalization, and banking crises all increase market income inequality within countries.
@@
 
-\citet{claessens2007finance} explore the connectedness of growth, finance, and inequality, stressing the importance of institutional background in the effects.
-
-\citet{perotti2007investor} present a theoretical framework how established interests lobby for lower investor protection to prevent entrance of the new competitors.
-Greater accountability of the politicians is reflected in higher bribe required from teh lobbyist therefore higher accountability implies higher market entry on more competition.
-After confirming their prediction empirically, they show that the most important factor of accountability is not the formal measure of democratic institutions, but \emph{newspaper readership} which they interpret as broad awareness of policy choices and their outcomes.
-
-Summary of the theoretical preditions in \citet{demirgucc2009finance} on pages 20-23 for the review.
-
-\citet{furceri2019robust} apply \ac{wals} to identify robust determinants of income inequality.
+% \citet{furceri2019robust} apply WALS to identify robust determinants of income inequality.
 They are the closest paper to ours since they account for model uncertainty in the estimation.
@@
 
-\section{Data}
-% \label{sec:data}
-\begin{figure}
-    \caption{Gini Coefficient and Efficiency of Financial Institutions}
-    \label{fig:ginifie}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/FIEGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Access to Financial Institutions}
-    \label{fig:ginifia}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/FIAGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Depth of Financial Institutions}
-    \label{fig:ginifid}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/FIDGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Depth of Financial Markets}
-    \label{fig:ginifmd}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/FMDGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Government Expenditures}
-    \label{fig:ginigovexp}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/GovExpGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Mortality}
-    \label{fig:ginimort}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/MortGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and \ac{gdp} per capita}
-    \label{fig:ginigdp}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/GDPGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Rule of Law}
-    \label{fig:ginirule}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/RuleLawGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Inflation}
-    \label{fig:giniinfl}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/InflGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Equipment Investment}
-    \label{fig:giniequipi}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/EquipIGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Trade Openness}
-    \label{fig:ginitrade}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/TradeOpenGiniNet}
-\end{figure}
-
-\begin{figure}
-    \caption{Gini Coefficient and Financial Development Indicators}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini}
-\end{figure}
-
+%
+%
+%
 %
@@
 
-\begin{figure}
-    \caption{Gini Coefficient and Efficiency of Financial Institutions}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/FIEGiniNet_dm}
-\end{figure}
+\section{Data}
+\cite{roineetal2009}, \citet{furceri2019robust} for the determinants of income inequality.
 
-\begin{figure}
-    \caption{Gini Coefficient and Access to Financial Institutions}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/FIAGiniNet_dm}
-\end{figure}
+\citet{nolan2019drivers} bring a survey of the literature on determinants of inequality, summarizing the complexity of the inequality dynamics.
+They stress that many of the determinants are interlinked which implies difficulty in assigning precise effects to individual drivers of inequality.
+Additionally, they encourage complementary individual country case studies to support the finding of the general cross-country estimates.
 
-\begin{figure}
-    \caption{Gini Coefficient and Depth of Financial Institutions}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/FIDGiniNet_dm}
-\end{figure}
+\citet{gwartney2017} provide an economic freedom index of the Fraser Institute.
 
-\begin{figure}
-    \caption{Gini Coefficient and Depth of Financial Markets}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/FMDGiniNet_dm}
-\end{figure}
+Both theoretical and empirical studies leave out the issue of importing the financial services from abroad. We include financial globalization from KOF among our control variables.
 
-\begin{figure}
-    \caption{Gini Coefficient and Government Expenditures}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/GovExpGiniNet_dm}
-\end{figure}
+% % \label{sec:data}
+% \begin{figure}
+%     \caption{Gini Coefficient and Efficiency of Financial Institutions}
+%     \label{fig:ginifie}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/FIEGiniNet}
+% \end{figure}
 
-\begin{figure}
-    \caption{Gini Coefficient and Mortality}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/MortGiniNet_dm}
-\end{figure}
+% \begin{figure}
+%     \caption{Gini Coefficient and Access to Financial Institutions}
+%     \label{fig:ginifia}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/FIAGiniNet}
+% \end{figure}
 
-\begin{figure}
-    \caption{Gini Coefficient and \ac{gdp} per capita}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/GDPGiniNet_dm}
-\end{figure}
+% \begin{figure}
+%     \caption{Gini Coefficient and Depth of Financial Institutions}
+%     \label{fig:ginifid}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/FIDGiniNet}
+% \end{figure}
 
-\begin{figure}
-    \caption{Gini Coefficient and Rule of Law}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/RuleLawGiniNet_dm}
-\end{figure}
+% \begin{figure}
+%     \caption{Gini Coefficient and Depth of Financial Markets}
+%     \label{fig:ginifmd}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/FMDGiniNet}
+% \end{figure}
 
-\begin{figure}
-    \caption{Gini Coefficient and Inflation}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/InflGiniNet_dm}
-\end{figure}
+% \begin{figure}
+%     \caption{Gini Coefficient and Government Expenditures}
+%     \label{fig:ginigovexp}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/GovExpGiniNet}
+% \end{figure}
 
-\begin{figure}
-    \caption{Gini Coefficient and Equipment Investment}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/EquipIGiniNet_dm}
-\end{figure}
+% \begin{figure}
+%     \caption{Gini Coefficient and Mortality}
+%     \label{fig:ginimort}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/MortGiniNet}
+% \end{figure}
 
-\begin{figure}
-    \caption{Gini Coefficient and Trade Openness}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/TradeOpenGiniNet_dm}
-\end{figure}
+% \begin{figure}
+%     \caption{Gini Coefficient and \ac{gdp} per capita}
+%     \label{fig:ginigdp}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/GDPGiniNet}
+% \end{figure}
 
-\begin{figure}
-    \caption{Gini Coefficient and Financial Development Indicators}
-\includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini_dm}
-\end{figure}
+% \begin{figure}
+%     \caption{Gini Coefficient and Rule of Law}
+%     \label{fig:ginirule}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/RuleLawGiniNet}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Inflation}
+%     \label{fig:giniinfl}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/InflGiniNet}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Equipment Investment}
+%     \label{fig:giniequipi}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/EquipIGiniNet}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Trade Openness}
+%     \label{fig:ginitrade}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/TradeOpenGiniNet}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Financial Development Indicators}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini}
+% \end{figure}
+
+% %
+% %
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Efficiency of Financial Institutions}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/FIEGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Access to Financial Institutions}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/FIAGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Depth of Financial Institutions}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/FIDGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Depth of Financial Markets}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/FMDGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Government Expenditures}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/GovExpGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Mortality}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/MortGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and \ac{gdp} per capita}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/GDPGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Rule of Law}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/RuleLawGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Inflation}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/InflGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Equipment Investment}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/EquipIGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Trade Openness}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/TradeOpenGiniNet_dm}
+% \end{figure}
+
+% \begin{figure}
+%     \caption{Gini Coefficient and Financial Development Indicators}
+% \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini_dm}
+% \end{figure}
 
@@
 % \label{sec:results}
+The benefits of capital markets liberalization seem to be concetrated to the top of the income distribution.
+Top quintile of the distribution accrues nearly all of the income growth following the liberalization while the share of middle three quantiles decreases and the bottom remains unaffected \citep{das2003income}.
 
+\citet{kroszneretal2007} show that financial crises have relative more severe impact on the sectors which depend more on external financing.
+The consequences of crises on firms relate to institutional environment and materialize through lower production capacity and competition.
 % \label{sec:conclusion}

### 2020-01-08 Paper/paper.tex
update

@@
 
+\acrodef{bic}[BIC]{Bayesian Information Criterion}
 \acrodef{bma}[BMA]{Bayesian Model Averaging}
-\acrodef{bic}[BIC]{Bayesian Information Criterion}
 \acrodef{cbf}[CBF]{Conditional Bayes Factor}
@@
 \acrodef{cswd}[CSWD]{Credit Suisse Wealth Databook}
-\acrodef{gfdd}[GFDD]{Global Financial Development Database}
-\acrodef{pip}[PIP]{Posterior Inclusion Probability}
-\acrodefplural{pip}[PIPs]{Posterior Inclusion Probabilities}
+\acrodef{dina}[DINA]{Distributional National Accounts}
+\acrodef{efw}[EFW]{Economic Freedom of the World}
 \acrodef{fdi}[FDI]{Foreign Direct Investment}
@@
 \acrodef{gdp}[GDP]{Gross Domestic Product}
+\acrodef{gfdd}[GFDD]{Global Financial Development Database}
+\acrodef{hbs}[HBS]{Household Balance Sheet}
+\acrodef{ivbma}[IVBMA]{Instrumental Variable Bayesian Model Averaging}
+\acrodef{lis}[LIS]{Luxembourg Income Study}
+\acrodef{oecd}[OECD]{Organisation for Co-operation and Development}
+\acrodef{pip}[PIP]{Posterior Inclusion Probability}
+\acrodefplural{pip}[PIPs]{Posterior Inclusion Probabilities}
+\acrodef{pmp}[PMP]{Posterior Model Probability}
+\acrodef{swiid}[SWIID]{Standardized World Income Inequality Database}
 \acrodef{uip}[UIP]{Unit Information Prior}
-\acrodef{wb}[WB]{World Bank}
-\acrodef{oecd}[OECD]{Organisation for Co-operation and Development}
-\acrodef{ivbma}[IVBMA]{Instrumental Variable Bayesian Model Averaging}
-\acrodef{wid}[WID]{World Inequality Database}
-\acrodef{hbs}[HBS]{Household Balance Sheet}
 \acrodef{uk}[UK]{United Kingdom}
 \acrodef{us}[US]{United States}
-\acrodef{efw}[EFW]{Economic Freedom of the World}
-\acrodef{pmp}[PMP]{Posterior Model Probability}
 \acrodef{wals}[WALS]{Weighted-average Least Squares}
+\acrodef{wb}[WB]{World Bank}
+\acrodef{wid}[WID]{World Inequality Database}
+\acrodef{wiid}[WIID]{World Income Inequality Database}
 
@@
 \onehalfspacing
-\title{Finance and Income Inequality - panel BMA approach\thanks{We thank two anonymous referees, ...}}
-\author[a]{Roman Horvath}
+\title{Finance and Income Inequality - panel BMA approach\thanks{The author acknowledges support from Charles University Research Centre program No. UNCE/HUM/035}}
+% \author[a]{Roman Horvath}
 \author[a]{Jan Mares\footnote{\footnotesize Corresponding author's address: IES FSV UK, Opletalova 26, 110 00, Praha 1; \textbf{e-mail:janxmares@gmail.com}}}
@@
 
-\clubpenalty 9999 % no orphants (typographic properties)
+\clubpenalty 9999 % no orphans (typographic properties)
 \widowpenalty 9999 % no widows (typographic properties)
@@
 
-A key divide appears between the effect of financial develpoment on extenstive and intensive margin.
+A key divide appears between the effect of financial development on extensive and intensive margin.
 The extensive margin captures the extend to which individuals, who had not been using financial services before, gain access.
@@
 Consequently, the efficiency of resource allocation also increases, which enhances growth and reduces inequality.
-\citet{perotti2007investor} present a framework based on politcal economy.
+\citet{perotti2007investor} present a framework based on political economy.
 They argument depends on a lobby for lower investor protection to prevent entrance of the new competitors.
@@
 Thus, with increasing accountability, investor protection strengthens and spurs market entry and competition.
-The authors examine their prediction in a cross-section and show that better investor protection correlates with larger entry rates and higher firm density in more financially intensive sectors\footnote{In addition, they show that the most important factor of accountability is not the formal measure of democratic institutions, but \emph{newspaper readership} which they interpret as broad awareness of policy choices and their outcomes.}.
+The authors examine their prediction in a cross-section and show that better investor protection correlates with larger entry rates and higher firm density in more financially intensive sectors\footnote{In addition, they show that the most important factor of accountability is not the formal measure of democratic institutions, but newspaper readership which they interpret as broad awareness of policy choices and their outcomes.}.
 
@@
 
-A set of distinct papers explores the relationship between inequality and growth while stressing financial markest imperfections driving the outcomes.
+A set of distinct papers explores the relationship between inequality and growth while stressing financial markets imperfections driving the outcomes.
 Income inequality and growth may intersect through varying channels.
-Accumulation of savings, unobservable effort, and investment project size favour the prediction of growth inducing inequality.
+Accumulation of savings, unobservable effort, and investment project size favor the prediction of growth inducing inequality.
 Negative impact of inequality on human capital accumulation, entrepreneurial activity provide argument for the opposing view.
@@
 High inequality seems to hurt the prospects of the poor while the top of the distribution is unaffected.
-The rich thus disproportianately benefit from higher inequality as their subsequent income exhibit faster growth.
-The authors attribute this effect to the political channel the rich use to lobby in favour of the policies which support their economic interests.
+The rich thus disproportionately benefit from higher inequality as their subsequent income exhibit faster growth.
+The authors attribute this effect to the political channel the rich use to lobby in favor of the policies which support their economic interests.
 Preferences of the rich are ultimately more likely to determine public policy than the preferences of the majority \citep{gilens_page_2014}.
@@
 However, there are severe limitation to their study.
-First, their coverarage does not correspond to the availability of the data.
+First, their coverage does not correspond to the availability of the data.
 Second, they do not account for the different dimensions of finance simultaneously and always use a single indicator of development at a time.
@@
 They are the closest paper to ours since they account for model uncertainty in the estimation.
-Their focus if more general rathar than on finance specifically.
+Their focus if more general rather than on finance specifically.
 Out work offers more detailed contribution in terms of the role of finance in shaping income inequality.
@@
 \section{Data}
+The key variable in the paper is the measure of income inequality.
+We want to examine how financial development affects income inequality and whether the effect might by different at the top quantiles of income distribution.
+As the overall measure of income inequality, we rely the after-tax Gini coefficient from \ac{swiid} by \citet{Solt2019}, which is a standard resource in the literature\footnote{There are alternative sources of for Gini coefficient, e.g.
+\ac{wiid} or \ac{lis}, but each of them brings limitations in terms of comparability or coverage.}.
+Its critical advantage lies in the widespread coverage across countries and time and a unified methodology which provides a reasonable level of comparability.
+It typically takes values in the interval between 0 and 100 where the former suggests perfect equality (everyone in the economy enjoys the same income) and the latter perfect inequality (all the income goes to only a single unit).
+We depart from existing papers slightly in considering the after-tax rather than the before-tax income distribution as a dependent variable.
+
+To explore the relationship in the top part of the distribution, we choose top income share from \ac{wid}\footnote{The methodology and guidelines to database are provided by \citet{alvaredo2016distributional}.}.
+The surveys suffer from well-known issues of underrepresentation of the top income earners and the distortions resulting from self-reported character of the data.
+This can influence not only the top income shares resulting from survey data, but also distortions in the overall measures of inequality.
+The data in \ac{wid} make use of income tax records in individual countries and the derived shares obtained using consistent methodology of \ac{dina} are arguably more reliable relative to the survey-based measures which are the primary source of majority estimates of income distributions.
+
+The data spans from 2000 to 2014.
+We follow the literature \citep{dabla2015causes,de2017finance} and average both the inequality measure (dependent variable) and the potential determinants (independent variables) across three year intervals.
+There are important reasons for looking at the averages than observation in individual years.
+Annual macroeconomic data are subject to fluctuations and the data on income inequality is noisy \citet{delis2014}.
+Averaging should diminish the level of noise.
+On the top of that, the variables at the center of our analysis, e.g.
+stock market capitalization or credit to \ac{gdp}, are likely to be affected by the business cycles and volatile on the yearly basis.
+Similar argument holds for top income shares, as they depend, among other things, on the bonuses paid out each year and capital income.
+We want to explore the long-term rather than the short-term relationship and that guides the choice of averaged data.
+Faced against the trade-off between length of the averaging periods and available observations in the time dimension, we take a compromise of three years in contrast to the literature, where generally the 5-year intervals apply.
+The availability of financial development indicators limits the analysis to a period from 2000 onward and we prefer to keep at least 5 unique time periods to just three under the case of 5-year average\footnote{Nevertheless, we run the estimation with 5-year averages of data as a robustness check and find no critical qualitative differences compared to the baseline.}.
+
+Table \ref{tab:ineq} report the summary statistics and correlations between the income inequality variables.
+
+\begin{table}[ht!]
+    \small
+    \caption{Summary statistics of inequality variables}
+    \label{tab:ineq}
+    \centering
+       \begin{tabular}{lrrrr}
+        \toprule
+        Variable & Mean & St. Dev. & Min & Max \\
+        After-tax Gini (SWIID) & & & & \\
+        Top 10\% income share & & & & \\
+        Top 1\% income share & & & & \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
 \cite{roineetal2009}, \citet{furceri2019robust} for the determinants of income inequality.
@@
 Additionally, they encourage complementary individual country case studies to support the finding of the general cross-country estimates.
-
-\citet{gwartney2017} provide an economic freedom index of the Fraser Institute.
 
@@
 % \label{sec:results}
-The benefits of capital markets liberalization seem to be concetrated to the top of the income distribution.
+The benefits of capital markets liberalization seem to be concentrated to the top of the income distribution.
 Top quintile of the distribution accrues nearly all of the income growth following the liberalization while the share of middle three quantiles decreases and the bottom remains unaffected \citep{das2003income}.
@@
 The consequences of crises on firms relate to institutional environment and materialize through lower production capacity and competition.
-% \label{sec:conclusion}
+
+\section{Conclusion}
+\label{sec:conclusion}
 
@@
 %
-\clearpage
+% \clearpage
 %
-\appendix
-\section{Appendix}
+% \appendix
+% \section{Appendix}
 % \label{sec:app}

### 2020-01-08 Scripts/191219 analysis alternative with pre-tax Gini index.R

@@ -1,0 +1,94 @@
+# Financial Development and Income Inequality
+# Alternative with pre-tax Gini index, 3-y averages
+# Jan Mares, 1912019
+
+# Libraries
+library(data.table)
+library(countrycode)
+library(BMS)
+library(dilutBMS2)
+library(imputeTS)
+library(here)
+library(dummies)
+
+# read data file
+data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
+# View(data)
+
+# drop countries with unavailable data
+data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+
+# adjust the variables
+data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
+data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
+
+#
+#
+# Averages - 3 year, Gini index
+#
+#
+
+# Averages over years
+data_f <- data[year %in% c(2000:2014), ]
+data_f[year %in% c(2000:2002), period := 1]
+data_f[year %in% c(2003:2005), period := 2]
+data_f[year %in% c(2006:2008), period := 3] 
+data_f[year %in% c(2009:2011), period := 4]
+data_f[year %in% c(2012:2014), period := 5]
+
+# average observations by period and country
+data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
+data_f[, c('year') := NULL]
+
+data_f <- data_f[!is.nan(GiniNet),]
+
+data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
+		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
+		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
+		   "FMA","FME", # unavailable for a lot of years
+		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
+		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
[…]

### 2020-01-08 Scripts/191219 analysis baseline 3-year.R

@@ -10,22 +10,6 @@ library(dilutBMS2)
 library(imputeTS)
 library(here)
 library(dummies)
-library(ineq)
-
-
-
-lines(p,Lc.lognorm(p, parameter=0.5),col="red")
-lines(p,Lc.lognorm(p, parameter=1.0),col="blue")
-lines(p,Lc.lognorm(p, parameter=1.5),col="green")
-
-gini = 1/((2*aplha+0.45)-1)
-
-aplha
-s = sqrt(2)*qnorm((0.56+1)/2)
-alpha = 1/(2*0.56)+1/2
-p = seq(0,1,0.01)
-plot(p,Lc.lognorm(p, parameter=s), type="l", col="brown")
-lines(p,Lc.pareto(p, parameter=alpha), col="blue")
 
 # read data file
 data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
@@ -99,7 +83,7 @@ data_dm[, c("period","period_1","WarYears","RevCoups","Top1share","Top10share") 
 dummies <- colnames(dum)[2:(ncol(dum))]
 
 # run BMA
-bma_3y_gini_baseline <- bms(data_dm, iter=500000, burn=100000, mprior = "uniform", g = "hyper",
+bma_3y_gini_baseline <- bms(data_dm, iter=5000000, burn=1000000, mprior = "uniform", g = "hyper",
                     nmodel=5000, mcmc="bd.int",
                     fixed.reg = dummies, user.int = F)
 

### 2020-01-08 Paper/paper.tex
update

@@
 % \author[a]{Roman Horvath}
-\author[a]{Jan Mares\footnote{\footnotesize Corresponding author's address: IES FSV UK, Opletalova 26, 110 00, Praha 1; \textbf{e-mail:janxmares@gmail.com}}}
-\affil[a]{Charles University, Prague}
+\author{Jan Mares\footnote{\footnotesize Corresponding author's address: IES FSV UK, Opletalova 26, 110 00, Praha 1; \textbf{e-mail:janxmares@gmail.com}}}
+\affil{Charles University, Prague}
 
@@
         Variable & Mean & St. Dev. & Min & Max \\
-        After-tax Gini (SWIID) & & & & \\
+        \midrule
+        After-tax Gini index  & & & & \\
         Top 10\% income share & & & & \\
@@
 
-\cite{roineetal2009}, \citet{furceri2019robust} for the determinants of income inequality.
+We obtain the financial development indicators from \ac{gfdd}.
+The database offers detailed indicators along four dimensions of financial systems and allows to estimate the affect of changes in access, size, efficiency, and stability of financial markets.
+Furthermore, we can distinguish between banking sector and financial markets in all these dimensions.
+The data in for access and stability of financial markets remains sparse in the concerned period we must leave them out of analysis.
+We use the version of financial indicators from \cite{svirydzenka2016introducing}.
+The authors make use of principal component analysis in order to construct aggregate indicators in each characteristic of financial sector.
+In summary, we have indicators of financial institutions depth (FID), financial markets depth (FMD), access to financial institutions (FIA), efficiency of financial markets (FME), and institutions (FIE)\footnote{\citet{svirydzenka2016introducing} extrapolate the indicators from top to bottom if the original variables are unavailable, we make sure that that at least one variable is available for the construction of the index and no artificial correlation introduced to the data.}.
+We report the composition of each indicator in Table \ref{tab:finind}.
 
-\citet{nolan2019drivers} bring a survey of the literature on determinants of inequality, summarizing the complexity of the inequality dynamics.
+\begin{figure}
+    \label{fig:gini_findev}
+        \caption{Gini Coefficient and Financial Development Indicators}
+    \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini}
+    \end{figure}
+
+Figure \ref{fig:gini_findev} shows the scatter plots between Gini coefficient and selected financial indicators.
+
+We build the choice of other explanatory variables on a reviews of income inequality drivers \citep{roineetal2009,nolan2019drivers}, related study of finance-inequality nexus \citep{de2017finance}, and a more general inquiry into the robust determinants of income inequality \citep{furceri2019robust}.
+Table \ref{app:exog} reports all the control variables and their sources.
+
+Specify the main categories of controls in the previous paragraph (based on Nolan + Furceri).
+
+% \citet{nolan2019drivers} bring a survey of the literature on determinants of inequality, summarizing the complexity of the inequality dynamics.
 They stress that many of the determinants are interlinked which implies difficulty in assigning precise effects to individual drivers of inequality.
@@
 
-Both theoretical and empirical studies leave out the issue of importing the financial services from abroad. We include financial globalization from KOF among our control variables.
+% Both theoretical and empirical studies leave out the issue of importing the financial services from abroad. We include financial globalization from KOF among our control variables.
+
+\begin{figure}
+    \label{fig:gini_findev_dm}
+        \caption{Gini Coefficient and Financial Development Indicators}
+    \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini_dm}
+    \end{figure}
 
@@
 
-% \begin{figure}
-%     \caption{Gini Coefficient and Financial Development Indicators}
-% \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini_dm}
-% \end{figure}
-
 %
@@
 %
-% \clearpage
+\clearpage
 %
 % \appendix
-% \section{Appendix}
-% \label{sec:app}
+\section{Appendix}
+\label{sec:app}
+
+\begin{table}[ht!]
+    \small
+    \caption{Underlying Components of Financial Development Indicators}
+    \label{tab:finind}
+    \centering
+    \begin{tabular}{ll}
+      \toprule
+      \textsc{Indicator} & \textsc{Measure} \\
+      \midrule
+      \multicolumn{2}{l}{\textbf{Financial institutions}} \\
+      \midrule
+      \multirow{2}{*}{Access} 	& Bank branches per 100,000 adults \\
+                                  & ATMs per 100,000 adults \\
+      \midrule
+      \multirow{6}{*}{Efficiency}		& Net interest margin \\
+                                  & Lending-deposits spread \\ 
+                                  & Noninterest income to total income \\
+                                  & Overhead costs to total assets \\
+                                  & Return on assets \\
+                                  & Return on equity \\
+
+      \midrule
+      \multirow{4}{*}{Depth}	& Domestic private credit to the real sector to the GDP \\
+                                  & Pension fund assets/GDP \\
+                                  & Mutual fund assets/GDP \\
+                                  & Insurance premiums life and nonlife/GDP \\
+      \midrule
+      \multicolumn{2}{l}{\textbf{Financial markets}} \\
+      \midrule
+      \multirow{6}{*}{Depth} 	& Stock market capitalization/GDP \\
+                                  & Stocks traded/GDP \\
+                                  & International debt securities of government/GDP \\
+                                  & Total debt securities of financial corporations/GDP \\
+                                  & Total debt securities of nonfinancial corporations/GDP \\
+      \midrule
+      Efficiency                  & Stock market turnover ratio (stocks traded/capitalization) \\
+      \bottomrule
+\end{tabular}
+\end{table}
 

### 2020-01-12 Paper/paper.tex
update

@@
 \usepackage{geometry}
-% \usepackage{hyperref} %hyperlink support
-% \hypersetup{pdfstartview={XYZ null null 1.00},bookmarksnumbered, hypertexnames=false, colorlinks=true, linkcolor=BrickRed, citecolor=MidnightBlue, urlcolor=MidnightBlue} % zoom: default screen; numbers for bookmarks
+\usepackage{hyperref} %hyperlink support  
+\hypersetup{colorlinks=false, breaklinks=false, unicode, linkcolor=black, urlcolor=black, pdfborder={0 0 0}, anchorcolor=black, citecolor=black, pdfstartview={}}
 
@@
 %\usepackage{parskip}
-\usepackage[nolist]{acronym}
+\usepackage[printonlyused]{acronym}
 \usepackage{longtable}
@@
 \acrodefplural{fdi}[FDIs]{Foreign Direct Investments}
+\acrodef{FIA}{Access to Financial Institutions}
+\acrodef{FID}{Financial Institutions Depth}
+\acrodef{FIE}{Financial Institutions Efficiency}
+\acrodef{FMD}{Financial Markets Depth}
+\acrodef{FME}{Financial Markets Efficiency}
 \acrodef{gdp}[GDP]{Gross Domestic Product}
@@
 \acrodef{pmp}[PMP]{Posterior Model Probability}
+\acrodef{PWT}{Penn World Table}
 \acrodef{swiid}[SWIID]{Standardized World Income Inequality Database}
@@
 \acrodef{wals}[WALS]{Weighted-average Least Squares}
-\acrodef{wb}[WB]{World Bank}
+\acrodef{WB}[WB]{World Bank}
 \acrodef{wid}[WID]{World Inequality Database}
@@
 Furthermore, we can distinguish between banking sector and financial markets in all these dimensions.
-The data in for access and stability of financial markets remains sparse in the concerned period we must leave them out of analysis.
+The data in for access and stability of stock markets remains sparse in the concerned period we must leave them out of analysis.
 We use the version of financial indicators from \cite{svirydzenka2016introducing}.
@@
 \begin{figure}
+    \caption{Gini Coefficient and Financial Development Indicators}
     \label{fig:gini_findev}
-        \caption{Gini Coefficient and Financial Development Indicators}
     \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini}
@@
 
-Figure \ref{fig:gini_findev} shows the scatter plots between Gini coefficient and selected financial indicators.
+\begin{figure}
+    \caption{Gini Coefficient and Financial Development Indicators}
+    \label{fig:gini_findev_dm}
+    \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini_dm}
+\end{figure}
+
+Figure \ref{fig:gini_findev} shows the scatter plots of all the available observations after averaging between Gini coefficient and selected financial indicators.
+Although the data is noisy, it suggest negative correlation between the variables.
+Figure \ref{fig:gini_findev_dm} then shows the corresponding scatter plots of the data which was demeaned using the within country averages over time.
+The relationship between Gini and financial indicators appears considerably weaker following the transformation, with the exception of access to financial institutions where visual inspection allows pointing to a negative correlation.
 
 We build the choice of other explanatory variables on a reviews of income inequality drivers \citep{roineetal2009,nolan2019drivers}, related study of finance-inequality nexus \citep{de2017finance}, and a more general inquiry into the robust determinants of income inequality \citep{furceri2019robust}.
+The potential regressors could be categorized in several groups.
+They control for economic and financial development, demographics, globalization, and institutional background.
 Table \ref{app:exog} reports all the control variables and their sources.
-
-Specify the main categories of controls in the previous paragraph (based on Nolan + Furceri).
 
@@
 % Both theoretical and empirical studies leave out the issue of importing the financial services from abroad. We include financial globalization from KOF among our control variables.
-
-\begin{figure}
-    \label{fig:gini_findev_dm}
-        \caption{Gini Coefficient and Financial Development Indicators}
-    \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini_dm}
-    \end{figure}
 
@@
 \section{Appendix}
-\label{sec:app}
+\label{sec:appch2}
+\renewcommand{\thesection}{A\arabic{section}}%
+\renewcommand{\thetable}{A\arabic{table}}%
+\renewcommand{\thefigure}{A\arabic{figure}}%
+\renewcommand{\theequation}{A\arabic{eq}}
 
+\subsection*{The composition of financial indicators}
+\label{subsec:finind_comp}
 \begin{table}[ht!]
@@
 
-\renewcommand{\thesection}{A\arabic{section}}%
-\renewcommand{\thetable}{A\arabic{table}}%
-\renewcommand{\thefigure}{A\arabic{figure}}%
-\renewcommand{\theequation}{A\arabic{eq}}
+\clearpage
+\subsection*{Dataset description}
+\begin{center}
+\footnotesize
+\begin{longtable}{l p{0.50\linewidth} p{0.3\linewidth}}
+    \caption{List of variables}
+    \label{app:exog} \\
+    \toprule 
+  Variable & Definition (+ optional comments) & Source \\
+    \midrule 
+  GiniNet & After\-tax Gini index based on distribution of income (The Standardized World Income Inequality Database). & \href{http://fsolt.org/swiid/}{Solt (2019)} \\
+  GiniMarket & Before-tax Gini index based on distribution of income  (The Standardized World Income Inequality Database). & \href{http://fsolt.org/swiid/}{Solt (2019)} \\	
+  Top10share & Share of income going top decile of the distribution. & \href{https://wid.world/}{WID} \\	
+  Top1share & Share of income going top percentile of the distribution. & \href{https://wid.world/}{WID} \\
+  FIA & Access to financial institutions & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\citet{svirydzenka2016introducing}} \\
+  FID & Financial institutions depth & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\citet{svirydzenka2016introducing}} \\
+  FIE & Financial institutions efficiency & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\citet{svirydzenka2016introducing}} \\
+  FMD & Financial markets depth & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\citet{svirydzenka2016introducing}} \\
+  FME & Financial markets efficiency & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\citet{svirydzenka2016introducing}} \\
+  GDPpc & Level of GDP per capita & \href{http://data.worldbank.org/indicator/NY.GDP.PCAP.PP.KD}{WB} \\
+  NatRes & Total natural resource rents are the sum of oil rents, natural gas rents, coal rents (hard and soft), mineral rents, and forest rents.
+& \href{http://data.worldbank.org/indicator/NY.GDP.TOTL.RT.ZS}{WB} \\
+  PopGrowth & Annual population growth 1980-2009 & \href{http://data.worldbank.org/indicator/SP.POP.GROW}{WB} \\
+  GovExp & General government final consumption expenditure (formerly general government consumption). & \href{http://data.worldbank.org/indicator/NE.CON.GOVT.ZS}{WB}\\
+  NNSavings & Net national savings (gross national savings less the value of consumption of fixed capital, \% GNI). & \href{http://data.worldbank.org/indicator/NY.ADJ.NNAT.GN.ZS}{WB} \\
+  EducExp & Education expenditure refers to the current operating expenditures in education, including wages and salaries and excluding capital investments in buildings and equipment..
+& \href{http://data.worldbank.org/indicator/NY.ADJ.AEDU.GN.ZS}{WB} \\
+  Infl & Inflation as measured by the consumer price index. & \href{http://data.worldbank.org/indicator/FP.CPI.TOTL.ZG}{WB} \\
+  VAA & Agriculture, forestry, and fishing value added (\% GDP). & \href{http://data.worldbank.org/indicator/NV.AGR.TOTL.ZS}{WB} \\
+  VAI & Industry value added (\% GDP). & \href{http://data.worldbank.org/indicator/NV.IND.TOTL.ZS}{WB} \\
+  GFCF & Gross fixed capital formation (\% of GDP). & \href{http://data.worldbank.org/indicator/NE.GDI.FTOT.ZS}{WB} \\
+  NetFDI &  Foreign direct investment, net inflows (\% of GDP). & \href{http://data.worldbank.org/indicator/BX.KLT.DINV.WD.GD.ZS}{WB} \\
+  GDPgrowth & Annual growth of GDP. & \href{http://data.worldbank.org/indicator/NY.GDP.MKTP.KD.ZG}{WB} \\
+  LifeExp & Life expectancy at birth.  & \href{http://data.worldbank.org/indicator/SP.DYN.LE00.IN}{WB}\\
+  LabForce & Total labor force comprises people ages 15 and older who meet the International Labor Organization definition of the economically active population: all people who supply labor for the production of goods and services during a specified period.
+Labor force total.
+& \href{http://data.worldbank.org/indicator/SL.TLF.CACT.ZS}{WB} \\
+%   PopDens & Population density (people per sq. km of land area). & \href{http://data.worldbank.org/indicator/EN.POP.DNST}{WB} \\
+  RuleOfLaw & Rule of law estimate & \href{http://data.worldbank.org/indicator/RL.EST}{WB} \\
+  CLandPR & Average of index for civil liberties and political rights & \href{https://freedomhouse.org/report/freedom-world-2016/methodology}{Freedom House} \\
+%   OutwardO & Trade (\% of GDP)  & \href{http://data.worldbank.org/indicator/NE.TRD.GNFS.ZS}{WB} \\
+  ChinnIto & Chinn-Ito index of financial openness. & \href{http://web.pdx.edu/~ito/Chinn-Ito_website.htm}{Chinn-Ito} \\ 
+  LeftWing & Dummy equal to 1 when left oriented party lead the country. & \href{http://www.nsd.uib.no/macrodataguide/set.html?id=11&sub=1}{DPI} \\
+  ActivRestrict & Activity restrictions.
+Regulatory restrictions on bank activities and the mixing of banking and commerce.
+& \href{http://faculty.haas.berkeley.edu/ross_levine/regulation.htm}{Barth et al.
+(2013)} \\
+  CapitalReg & Capital Regulatory index. & \href{http://faculty.haas.berkeley.edu/ross_levine/regulation.htm}{Barth et al. (2013)} \\
+  DiversIndex & Whether there are explicit, verifiable, quantifiable guidelines for asset diversification and banks are allowed to make loans abroad.
+& \href{http://faculty.haas.berkeley.edu/ross_levine/regulation.htm}{Barth et al.
+(2013)} \\
+  EducIndex & Calculated using mean years of schooling and expected years of schooling & \href{http://hdr.undp.org/en/content/education-index}{UN} \\
+  NetInterestMargin & Accounting value of banks' net interest revenue as a share of average interest-bearing assets; a measure of the efficiency of the banking sector.
+& \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
+  BankZScore & return on banks' assets plus the ratio of banks' equity and assets, divided by the standard deviation of the return on assets (ROA+equity/assets)/sd(ROA); a measure of stability of the banking sector & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
+  Privatecredit & Domestic private credit to the real sector to GDP; a measure of the depth of the banking sector & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
+  MarketCap & Value of listed shares to GDP; a measure of the depth of stock markets.& \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
+  MarketTurn & Stock market value traded to total market capitalization; a measure of the efficiency of stock markets.
+& \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
+  BankBranches & Number of bank branches per 100,000 adults & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
+  Loan2Deposits & Loan-to-deposit ratio. & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
+  Redist & Difference between market (pre-tax) and net (after-tax0 Gini index based on distribution of income  (The Standardized World Income Inequality Database).
+& \href{http://fsolt.org/swiid/}{Solt (2016)} \\
+  FinLib & Averaged components of Economic Freedom of the World index 3D (freedom to own foreign currency accounts), 4C (black-market exchange rates), 4D (controls of the movement of capital and people), and 5A (credit market regulations).
+& \href{https://www.fraserinstitute.org/economic-freedom/dataset}{\citet{gwartney2017}} \\
+  \bottomrule
+\end{longtable}
+\end{center}
 

### 2020-01-12 Scripts/191030 WB data.R

@@ -72,7 +72,7 @@ wbvars <- c("NY.GDP.TOTL.RT.ZS","NE.CON.GOVT.ZS","NY.GDS.TOTL.ZS","SP.POP.GROW",
 			"SL.EMP.TOTL.SP.NE.ZS","SE.XPD.TOTL.GD.ZS","NY.GDP.PCAP.PP.KD",
 			"NY.GDP.MKTP.KD.ZG","NE.TRD.GNFS.ZS","SP.DYN.LE00.IN",
 			"TX.VAL.MMTL.ZS.UN","TX.VAL.FUEL.ZS.UN","SE.PRM.ENRR",
-			"SE.SEC.ENRR","SE.TER.ENRR","SL.UEM.TOTL.ZS")
+			"SE.SEC.ENRR","SE.TER.ENRR","SL.UEM.TOTL.ZS","EN.POP.DNST")
 
 # Read first dataset to obtain selected country codes
 wbdata <- data.table(WDI(indicator = wbvars, start = 1980, end = 2017))
@@ -103,11 +103,11 @@ setnames(wbdatasub, c("NY.GDP.TOTL.RT.ZS","NE.CON.GOVT.ZS","NY.GDS.TOTL.ZS",
 				      "SE.XPD.TOTL.GD.ZS","NY.GDP.PCAP.PP.KD","NY.GDP.MKTP.KD.ZG",
 					  "NE.TRD.GNFS.ZS","SP.DYN.LE00.IN","TX.VAL.MMTL.ZS.UN",
 					  "TX.VAL.FUEL.ZS.UN","SE.PRM.ENRR","SE.SEC.ENRR","SE.TER.ENRR",
-					  "SL.UEM.TOTL.ZS"),
+					  "SL.UEM.TOTL.ZS","EN.POP.DNST"),
 					c("NatRes","GovExp","GDSavings","PopGrowth","PopTot","Infl","EducExp","VAI",
 					  "VAA","NetFDIin","NetFDIout","TaxR","NNSavings","RuleofLaw","GFCF",
 					  "EmplRate","PubEducExp","GDPpc","GDPgrowth","TradeOpen","LifeExp","ExpMetalOre",
-					  "ExpFuel","EnrolPri","EnrolSec","EnrolTer","Unemployment"))
+					  "ExpFuel","EnrolPri","EnrolSec","EnrolTer","Unemployment","PopDens"))
 
 # Check
 # names(wbdatasub)

### 2020-01-14 Scripts/191218 Figures GiniNet averages across time (BE estimate).R
update

@@ -1,0 +1,342 @@
+# Figures for the paper
+# Between estimates, figures
+# Jan Mares, 191218
+
+# Libraries
+library(data.table)
+library(imputeTS)
+library(here)
+library(ggplot2)
+library(gridExtra)
+
+# read data file
+data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
+# View(data)
+
+# drop countries with unavailable data
+# data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+data <- data[!(iso3c %in% c("BEN","COM","GIN","LBR","OMN","PSE","STP","SYC","SWZ")),] # excluding also 1 unit observations
+
+# adjust the variables
+data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
+data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
+
+#
+#
+#  Gini index
+#
+#
+
+# Averages over years
+data_f <- data[year %in% c(2000:2014), ]
+
+# average observations by country (across time)
+data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country")]
+data_f[, c('year') := NULL]
+
+data_f <- data_f[!is.nan(GiniNet),]
+
+data_try <- data_f[, c("country","iso3c","NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
+		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
+		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
+		   "FMA","FME", # unavailable for a lot of years
+		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
+		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
+		   "ExpMetalOre","ExpFuel", # correlated with NatRes
+		   "corr","burreau","demAcc","lawOrder", # ICRG data
+		   "GFCF", # duplicit with NonequipI + EquipI ) := NULL]
[…]

### 2020-01-14 Scripts/191218 Figures GiniNet.R

@@ -13,7 +13,8 @@ data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFac
 # View(data)
 
 # drop countries with unavailable data
-data <- data[!(iso3c %in% c("BEN","COM","GIN","LBR","OMN","PSE","STP","SYC","SWZ")),]
+# data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+data <- data[!(iso3c %in% c("BEN","COM","GIN","LBR","OMN","PSE","STP","SYC","SWZ")),] # excluding also 1 unit observations
 
 # adjust the variables
 data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
@@ -92,7 +93,7 @@ pl1g <- pl1 + geom_point(size=2) +
 		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
 			  panel.border = element_rect(fill = NA, colour = 'black'),
     		  legend.position="none") +
-        stat_smooth(method='lm', se=F) +
+        # stat_smooth(method='lm', se=F) +
         scale_colour_manual(values=rgb(colset/255)) + 
         scale_fill_manual(values=rgb(colset/255))
 
@@ -107,11 +108,19 @@ pl2g <- pl2 + geom_point(size=2) +
 		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
 			  panel.border = element_rect(fill = NA, colour = 'black'),
     		  legend.position="none") +
-        stat_smooth(method='lm', se=F) +
+        # stat_smooth(method='lm', se=F) +
         scale_colour_manual(values=rgb(colset/255)) + 
         scale_fill_manual(values=rgb(colset/255))
 
 pl2g
+
+#
+# Grids, GiniNet + Topshares
+#
+
+cairo_ps(file = here("Paper/figures/plots_ineq.eps"), width=10, height=5, family="Arial")
+grid.arrange(pl1g, pl2g, ncol=2)
+dev.off()
 
 #
 # Scatter plots of the variable with highest PIPs
@@ -127,7 +136,7 @@ pl1g <- pl1 + geom_point(size=2) +
 		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
 			  panel.border = element_rect(fill = NA, colour = 'black'),
     		  legend.position="none") +
-        stat_smooth(method='lm', se=F) +
[…]

### 2020-01-14 Scripts/191218 Figures Top10share.R

@@ -13,7 +13,8 @@ data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFac
 # View(data)
 
 # drop countries with unavailable data
-data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+# data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+data <- data[!(iso3c %in% c("BEN","COM","GIN","LBR","OMN","PSE","STP","SYC","SWZ")),] # excluding also 1 unit observations
 
 # adjust the variables
 data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
@@ -95,7 +96,7 @@ pl1g <- pl1 + geom_point(size=2) +
 		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
 			  panel.border = element_rect(fill = NA, colour = 'black'),
     		  legend.position="none") +
-        stat_smooth(method='lm', se=F) +
+        # stat_smooth(method='lm', se=F) +
         scale_colour_manual(values=rgb(colset/255)) + 
         scale_fill_manual(values=rgb(colset/255))
 
@@ -116,7 +117,7 @@ pl2g <- pl2 + geom_point(size=2) +
 		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
 			  panel.border = element_rect(fill = NA, colour = 'black'),
     		  legend.position="none") +
-        stat_smooth(method='lm', se=F) +
+        # stat_smooth(method='lm', se=F) +
         scale_colour_manual(values=rgb(colset/255)) + 
         scale_fill_manual(values=rgb(colset/255))
 
@@ -137,7 +138,7 @@ pl3g <- pl3 + geom_point(size=2) +
 		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
 			  panel.border = element_rect(fill = NA, colour = 'black'),
     		  legend.position="none") +
-        stat_smooth(method='lm', se=F) +
+        # stat_smooth(method='lm', se=F) +
         scale_colour_manual(values=rgb(colset/255)) + 
         scale_fill_manual(values=rgb(colset/255))
 
@@ -158,7 +159,7 @@ pl4g <- pl4 + geom_point(size=2) +
 		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
[…]

### 2020-01-14 Scripts/191218 Figures Top1share.R

@@ -13,7 +13,8 @@ data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFac
 # View(data)
 
 # drop countries with unavailable data
-data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+# data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+data <- data[!(iso3c %in% c("BEN","COM","GIN","LBR","OMN","PSE","STP","SYC","SWZ")),] # excluding also 1 unit observations
 
 # adjust the variables
 data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
@@ -39,7 +40,7 @@ data_f[year %in% c(2012:2014), period := 5]
 data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
 data_f[, c('year') := NULL]
 
-data_f <- data_f[!is.nan(Top1share0share),]
+data_f <- data_f[!is.nan(Top1share),]
 
 data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
 		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
@@ -95,7 +96,7 @@ pl1g <- pl1 + geom_point(size=2) +
 		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
 			  panel.border = element_rect(fill = NA, colour = 'black'),
     		  legend.position="none") +
-        stat_smooth(method='lm', se=F) +
+        # #stat_smooth(method='lm', se=F) +
         scale_colour_manual(values=rgb(colset/255)) + 
         scale_fill_manual(values=rgb(colset/255))
 
@@ -116,7 +117,7 @@ pl2g <- pl2 + geom_point(size=2) +
 		theme(plot.margin = unit(c(0.3,0.2,0.3,0.3), "cm"), 
 			  panel.border = element_rect(fill = NA, colour = 'black'),
     		  legend.position="none") +
-        stat_smooth(method='lm', se=F) +
+        # #stat_smooth(method='lm', se=F) +
         scale_colour_manual(values=rgb(colset/255)) + 
         scale_fill_manual(values=rgb(colset/255))
 
[…]

### 2020-01-14 Scripts/191219 analysis baseline 3-year.R

@@ -10,13 +10,15 @@ library(dilutBMS2)
 library(imputeTS)
 library(here)
 library(dummies)
+library(xtable)
 
 # read data file
 data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
 # View(data)
 
 # drop countries with unavailable data
-data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+# data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+data <- data[!(iso3c %in% c("BEN","COM","GIN","LBR","OMN","PSE","STP","SYC","SWZ")),] # excluding also 1 unit observations
 
 # adjust the variables
 data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
@@ -61,8 +63,57 @@ data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","E
 
 data_try <- data_try[complete.cases(data_try), ]
 
+#
+#
+#
+#
+# Summary statistics
+#
+#
+#
+#
+
+# Filter the variables for correlations
+sumstats <- data_try[, .(GiniNet,Top10share,Top1share,FIA,FIE,FID,FMD)]
+summary(sumstats)
+
+sd(sumstats$GiniNet)
+sd(sumstats$Top10share)
+sd(sumstats$Top1share)
+sd(sumstats$FIA)
+sd(sumstats$FIE)
+sd(sumstats$FID)
+sd(sumstats$FMD)
+
 # demean the data
 data_dm <- data_try[, lapply(.SD[,2:ncol(.SD)], demean, na.rm = T), by = c("iso3c","country")]
+View(data_dm)
+#
+#
+#
+#
+# Correlations
+#
+#
+#
+#
+
+# Filter the variables for correlations
+data_corr <- data_dm[, .(GiniNet,Top10share,Top1share,FIA,FIE,FID,FMD)]
+
+# compute the correlation
+corr <- cor(data_corr)
+corr <- round(corr,2) # round to 2 decimal places
+corr[upper.tri(corr)] <- "" # remove the upper triangular part of correlation matrix
+
+# print output to latex
+print(xtable(corr))
+
+#
+#
+# Estimates
+#
+#
 
 # check
 # View(data_try)
@@ -207,7 +258,7 @@ save(bma_3y_top10_dilut, file = here("Results/bma_3y_top10_dilut.Rdata"))
 #
 
[…]

### 2020-01-14 Scripts/191221 Figures robustness.R

@@ -6,6 +6,9 @@ library(data.table)
 library(dilutBMS2)
 library(here)
 library(ggplot2)
+library(gridExtra)
+
+# plot comparison, different model priors
 
 # Gini 
 
@@ -14,27 +17,47 @@ load(here("Results/bma_3y_gini_dilut.Rdata"))
 load(here("Results/bma_3y_gini_baseline.Rdata"))
 load(here("Results/bma_3y_gini_random.Rdata"))
 
+#
+#
+
 pl_comp_gini <- plotComp(bma_3y_gini_baseline, bma_3y_gini_dilut, bma_3y_gini_random,
-                         lwd=1, varNr = NULL, comp = "PIP", exact = T, include.legend = TRUE,
-                         add.grid = F, do.par = T, cex.axis=0.8, cex.xaxis=0.65,
+                         lwd=1, varNr = NULL, comp = "PIP", exact = T, include.legend = F,
+                         add.grid = F, do.par = T, cex.axis=0.8, cex.xaxis=0.8,
                          col=c(rgb(0.03, 0.15, 0.4),rgb(0.753,0,0),rgb(0, 0, 0)))
 
-cairo_ps(file = paste0(folderfig,"model_priors_comparison_gini.eps"), width=9, height=6, family="helvetica")
-pl_comp_gini
+# write in file
+cairo_ps(file = here("Paper/figures/model_priors_comparison_gini.eps"), width=9, height=6, family="Arial")
+par(mar=c(6,4,0,0))
+plotComp(bma_3y_gini_baseline, bma_3y_gini_dilut, bma_3y_gini_random,
+         lwd=1, varNr = NULL, comp = "PIP", exact = T, include.legend = F,
+         add.grid = F, do.par = T, cex.axis=0.8, cex.xaxis=0.8,
+         col=c(rgb(0.03, 0.15, 0.4),rgb(0.753,0,0),rgb(0, 0, 0)))
+
+legend("topright", legend=c("Uniform", "Dilution","Random"), pch=c(1:3), bty="n", inset=c(0.02,0.02),
+       title = expression(bold("Model priors")), title.adj = 0.1, col=c(rgb(0.03, 0.15, 0.4),rgb(0.753,0,0),rgb(0, 0, 0)))
 dev.off()
 
-# Top 10 share
+# plot comparison - Top 10 share
 load(here("Results/bma_3y_top10_dilut.Rdata"))
[…]

### 2020-01-14 Scripts/200103 analysis baseline 3-year, between estimation.R

@@ -1,0 +1,253 @@
+# Financial Development and Income Inequality
+# Baseline analysis, between estimate
+# Jan Mares, 1912019
+
+# Libraries
+library(data.table)
+library(countrycode)
+library(BMS)
+library(dilutBMS2)
+library(imputeTS)
+library(here)
+library(dummies)
+
+# read data file
+data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
+# View(data)
+
+# drop countries with unavailable data
+data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+names(data)
+# adjust the variables
+data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
+data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, EducIndex^2)]
+
+#
+#
+#  Gini index
+#
+#
+
+# Averages over years
+data_f <- data[year %in% c(2000:2014), ]
+
+# average observations by country (across time)
+data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country")]
+data_f[, c('year') := NULL]
+
+data_f <- data_f[!is.nan(GiniNet),]
+
+data_try <- data_f[, c("country","iso3c","NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
+		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
+		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
+		   "FMA","FME", # unavailable for a lot of years
+		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
+		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
+		   "ExpMetalOre","ExpFuel", # correlated with NatRes
+		   "corr","burreau","demAcc","lawOrder", # ICRG data
+		   "GFCF", # duplicit with NonequipI + EquipI ) := NULL]
+		#  "WarYears","RevCoups", # invariant in the sample
+		   "GiniMarket","RedistRel", # alternative inequality indicators
[…]

### 2020-01-14 Scripts/Output tables to LaTeX.R

@@ -8,15 +8,18 @@ library(xtable)
 
 # Load the results
 # load(file = here("Results/bma_1y_gini_baseline.Rdata"))
-# load(file = here("Results/bma_3y_gini_baseline.Rdata"))
 # load(file = here("Results/bma_5y_gini_baseline.Rdata"))
 # load(file = here("Results/bma_1y_top10_baseline.Rdata"))
-# load(file = here("Results/bma_3y_top10_baseline.Rdata"))
 # load(file = here("Results/bma_5y_top10_baseline.Rdata"))
 # load(file = here("Results/bma_1y_top1_baseline.Rdata"))
-# load(file = here("Results/bma_3y_top1_baseline.Rdata"))
-load(file = here("Results/bma_5y_top1_baseline.Rdata"))
-summary(bma_5y_top1_baseline)
+
+# load(file = here("Results/bma_5y_top1_baseline.Rdata"))
+# summary(bma_5y_top1_baseline)
+
+# 3y averages (baseline)
+load(file = here("Results/bma_3y_gini_baseline.Rdata"))
+load(file = here("Results/bma_3y_top10_baseline.Rdata"))
+load(file = here("Results/bma_3y_top1_baseline.Rdata"))
 
 #
 #
@@ -25,7 +28,7 @@ summary(bma_5y_top1_baseline)
 
 # Variable names setup
 explvarshort <- c("NatRes","GovExp","GDSavings","PopGrowth","PopTot","Infl","EducExp",
-                  "VAI","VAA","NetFDIin","RuleofLaw","Mortality","GDPpc","TradeOpen",
+                  "VAI","VAA","NetFDIin","RuleofLaw","Mortality","GDPpc","GDPgrowth","TradeOpen",
                   "LifeExp","Unemployment","EquipI","NonequipI","kaopen","FID","FIA",
                   "FIE","FMD","EFW","EducIndex","CLandPR","LeftWing","RedistAbs","globFin",
                   "globRestrict","globSoc","globPol","GDPpc#GDPpc","Infl#Infl",
@@ -37,11 +40,12 @@ explvarshort <- c("NatRes","GovExp","GDSavings","PopGrowth","PopTot","Infl","Edu
 explvarnames <- c("Natural resources rents","Government expenditures","Gross domestic savings",
                   "Population growth","Total population","Inflation","Education expenditures",
                   "Value added in industry","Value added in agriculture","Net FDI (% GDP)",
-                  "Rule of law","Mortality","GDP per capita", "Trade openness","Life expectancy",
[…]

### 2020-01-15 Scripts/191030 PWT data.R
update

@@ -44,8 +44,8 @@ investment_f <- merge(gfcf, capital_detail[,j = list(iso3c = countrycode,year = 
 					  by = c("iso3c","year"))
 
 # Construct the shares of equip and nonequip investment
-investment_f[,EquipI := mach_share*csh_i]
-investment_f[,NonequipI := rest_share*csh_i]
+investment_f[, EquipI := mach_share*csh_i]
+investment_f[, NonequipI := rest_share*csh_i]
 investment_f[, to:= mach_share + rest_share]
 
 # select required columns

### 2020-02-12 Paper/elsarticle.tex
elsarticle update

\listfiles
\documentclass[preprint, nonatbib, 10pt]{elsarticle}

% Fix for using biblatex
\makeatletter
\let\c@author\relax
\makeatother

\usepackage[utf8]{inputenc}

%-----<<<<<<<<<<<< BIBLIOGRAPHY with biblatex >>>>>>>>>>>>----- 	
\usepackage[backend=biber, style=apa]{biblatex}
\DeclareSourcemap{
  \maps[datatype=bibtex]{
    \map{ % set fields in .bib file to ignore
      \step[fieldset=issn, null]
      \step[fieldset=isbn, null]
      \step[fieldset=abstract, null]
      \step[fieldset=doi, null]
      \step[fieldset=url, null]
    }
  }
}
%
\addbibresource{literature.bib}	% bib file
%-----<<< --------------------------------- >>>-----

%% Packages
\usepackage{lineno,hyperref}
\usepackage[nolist]{acronym}
% \usepackage{natbib}
\usepackage{booktabs}
\usepackage{setspace}
\usepackage{epstopdf}
\usepackage{longtable}
\usepackage{multirow}

%
%
%% Acronyms
\acrodef{BIC}{Bayesian Information Criterion}
\acrodef{BMA}{Bayesian Model Averaging}
\acrodef{BRIC}{Bayesian Risk Inflation Criterion}
\acrodef{CBF}{Conditional Bayes Factor}
\acrodefplural{CBF}[CBFs]{Conditional Bayes Factors}
\acrodef{CSWD}{Credit Suisse Wealth Databook}
\acrodef{DINA}{Distributional National Accounts}
\acrodef{EFW}{Economic Freedom of the World}
\acrodef{FDI}{Foreign Direct Investment}
\acrodefplural{FDI}[FDIs]{Foreign Direct Investments}
\acrodef{GDP}{Gross Domestic Product}
\acrodef{GFDD}{Global Financial Development Database}
\acrodef{HBS}{Household Balance Sheet}
\acrodef{IID}{Independent Identically Distributed}
\acrodef{IVBMA}{Instrumental Variable Bayesian Model Averaging}
\acrodef{LIS}{Luxembourg Income Study}
\acrodef{MCMC}{Markov-chain Monte Carlo}
\acrodef{ML}{Marginal Likelihood}
\acrodef{OECD}{Organisation for Co-operation and Development}
\acrodef{OLS}{Ordinary Least Squares}
\acrodef{PIP}{Posterior Inclusion Probability}
\acrodefplural{PIP}[PIPs]{Posterior Inclusion Probabilities}
\acrodef{PMP}{Posterior Model Probability}
\acrodefplural{PMP}[PMPs]{Posterior Model Probabilities}
\acrodef{PWT}{Penn World Table}
\acrodef{SWIID}{Standardized World Income Inequality Database}
\acrodef{UIP}{Unit Information Prior}
\acrodef{UK}{United Kingdom}
\acrodef{US}{United States}
\acrodef{WALS}{Weighted-average Least Squares}
\acrodef{WB}{World Bank}
\acrodef{WID}{World Inequality Database}
\acrodef{WIID}{World Income Inequality Database}
%
%
\modulolinenumbers[5]

\journal{European Journal of Political Economy}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

% Numbered
% \bibliographystyle{model1-num-names}

%% Numbered without titles
% \bibliographystyle{model1a-num-names}

%% Harvard
% \bibliographystyle{model2-names}\biboptions{authoryear}

%% Vancouver numbered
% \usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
% \usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
% \bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
% \usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style, distributed in TeX Live 2019
%\bibliographystyle{elsarticle-num}
% \usepackage{numcompress}\bibliographystyle{elsarticle-num-names}
% \bibliographystyle{elsarticle-harv}\biboptions{authoryear}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Finance and Inequality - panel BMA approach\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{The author acknowledges support from Charles University Research Centre program No. UNCE/HUM/035.}

%% Group authors per affiliation:
\author[1]{Jan Mare\v{s}}
\ead{janxmares@gmail.com}
\address[1]{IES, Faculty of Social Sciences, Charles University, Opletalova 26, Prague}

%% or include affiliations in footnotes:
% \author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
% \cortext[mycorrespondingauthor]{Corresponding author}

% \address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
% \address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
We investigate the impact of financial development on income inequality differentiating between depth, efficiency and access to financial markets and institutions. We apply panel Bayesian model averaging framework to address model uncertainty to reveal that financial development has complex relationship with the income distribution within countries. The access to and efficiency of banking decrease income inequality. The size of the markets has no relevance for overall income inequality, but relates to the higher top income shares. Moreover, unemployment along with investment into non-tangible assets increase income inequality while more redistribution imply lower levels of inequality.
\end{abstract}

\begin{keyword}
Finance, Inequality, Bayesian Model Averaging
\end{keyword}

\end{frontmatter}

% \linenumbers

\section{Introduction}
\label{ch4sec:intro}
Financial development alters how much the economic opportunities depend on the individual skills, family endowments, social status or political connections as individual may depend on financial system to provide loans to start new business, attain education, or temporarily fund their consumption. The research in the area of financial development and income inequality is well established. \textcite{demirgucc2009finance}, \textcite{claessens2007finance}, and more recently \cite{de2017finance} provide extensive reviews of the topic. A similar theme emerges in all three papers. The implications from theoretical contributions provide conflicting predictions about the relationship and empirical results bring evidence for both positive and negative effect. Although majority of the papers point towards finance tightening the distribution of income this results is not universal with some papers suggesting the opposite while other stress potential non-linearities.

A key divide appears between the effect of financial development on extensive and intensive margin. The extensive margin captures the extend to which individuals, who had not been using financial services before, gain access. On the other hand, the intensive margin describes growing use of finance by the agents who had already been using it before \parencite{demirgucc2009finance}. Financial development on the extensive margin might lead to more equal opportunities and outcomes. Access to credit by previously disadvantaged groups allows human capital accumulation \parencite{galorzeira1993income, galormoav2004, braunetal2019}, formation and growth of new firms \parencite{evans1989estimated,banerjeenewman1990}, with more evenly distributed economic opportunities as a result\footnote{Having similar economic opportunities might decrease the cross-generational inequality, by diminishing the effect of e.g. parental wealth. Depending on the innate abilities and talents of the individuals, however, it may increase the inequality of income within every generation at the same time.}.

On the contrary, intensive margin of financial development might disproportionately benefit the rich who may leverage financial services for their further benefit or to protect their existing rents. \textcite{GreenwoodJovanovic1990} present a model where the finance is the key driver of inequality and the welfare gains accrued by the incumbents - primarily the rich - in the initial development stage. With time, more agents meet the fixed costs of joining the financial intermediaries and they enjoy higher returns. Consequently, the efficiency of resource allocation also increases, which enhances growth and reduces inequality. \textcite{perotti2007investor} present a framework based on political economy. They argument depends on a lobby for lower investor protection to prevent entrance of the new competitors. The politicians require higher bribe from the lobbyist the greater is their accountability for policy decisions. Thus, with increasing accountability, investor protection strengthens and spurs market entry and competition. The authors examine their prediction in a cross-section and show that better investor protection correlates with larger entry rates and higher firm density in more financially intensive sectors\footnote{In addition, they show that the most important factor of accountability is not the formal measure of democratic institutions, but newspaper readership which they interpret as broad awareness of policy choices and their outcomes.}.

Financial development may also have indirect effect on income inequality through economic growth. \textcite{townsendeueda2006} model how finance interacts with production and allocation of credit. If increased use of finance increases the demand for low- relatively to the high-skilled workers, then it may have equalizing consequences for income distribution. Empirical evidence by \textcite{beck2010big} show that bank deregulation and increased competition in loan provision in the \ac{US} primarily benefited the workers with income below the median. Similarly, \textcite{delis2014} provide evidence of bank deregulation and liberalization tightening the income distribution, although this effect is only present in countries with high-quality institutions. They attribute the effect to the changes in labour market conditions and relatively higher wages and working hours of the low-skilled workers following the reforms.

A set of distinct papers explores the relationship between inequality and growth while stressing financial markets imperfections driving the outcomes. Income inequality and growth may intersect through varying channels. Accumulation of savings, unobservable effort, and investment project size favor the prediction of growth inducing inequality. Negative impact of inequality on human capital accumulation, entrepreneurial activity provide argument for the opposing view. 
\textcite{milanovicvan2018inequality} report how income inequality in the \ac{US} has different implications for the future income growth of the rich and the poor. High inequality seems to hurt the prospects of the poor while the top of the distribution is unaffected. The rich thus disproportionately benefit from higher inequality as their subsequent income exhibit faster growth. The authors attribute this effect to the political channel the rich use to lobby in favor of the policies which support their economic interests. Preferences of the rich are ultimately more likely to determine public policy than the preferences of the majority \parencite{gilens_page_2014}. High inequality together with a credit constraint and rich driving the political process results in low government spending and lasting inequality.

The literature does not converge on the conclusions even in the empirical cross-country and panel data studies. The papers link higher levels of financial development with lower levels of inequality \parencite{beck2007finance, hamori2012, gimet2011closer, kunieda2014finance}\footnote{For an extended list, we refer to \textcite{de2017finance}.}. On the other hand, several other estimate a inequality inducing effect of finance \parencite{Jaumotte2013, jauch2016financial, de2017finance}. Finally, some authors claim there the relationship might be non-linear, conditional on a threshold value of financial development \parencite{kim2011nonlinearity,tan2012nonlinear} or institutional quality \parencite{LawSingh2014, delis2014}.

Three papers are the closest to ours, each in a different respect. First, \textcite{de2017finance} examine different dimensions of finance on income inequality. Their results suggest that financial development, financial liberalization, and banking crises all increase pre-tax income inequality within countries. Additionally, they show that the effect of financial liberalization is conditional on democratic accountability. Higher accountability mitigates the impact of liberalization on inequality. On the contrary, the financial development, proxied by the credit to GDP ratio, has inequality increasing effect irrespective of the institutional background. Second, \textcite{naceurzhang2016} take similar approach in considering multiple dimensions --- the access, efficiency, and stability of the financial sector, although not examining the indicators simultaneously.  Third, \textcite{furceri2019robust} apply \ac{WALS} to identify robust determinants of income inequality. Their approach mirrors ours in accounting for model uncertainty in the estimation. Their focus is more general rather than focused primarily on finance. We provide synthesis and extension to these papers in providing more detailed view on the link between finance in shaping income inequality and examining multiple measures of inequality while specifically identifying the determinants of top income shares along with the determinants of the overall income distribution\footnote{Captured by income Gini index.}.

% An exemption from focus solely on the size of financial sector is the study by \textcite{naceurzhang2016} which applies a similar approach to ours, taking into consideration the access, efficiency, and stability of the financial sector. However, there are severe limitation to their study. First, their coverage does not correspond to the availability of the data. Second, they do not account for the different dimensions of finance simultaneously and always use a single indicator of development at a time. Third, they do not explicitly differentiate between the banking sector and financial markets. We provide a more detailed as well as more robust picture of the financial development effect on finance in these dimensions.

% \textcite{de2017finance} examine different dimensions of finance on income inequality. Their results suggest that financial development, financial liberalization, and banking crises all increase market income inequality within countries. Additionally, they show that the effect of financial liberalization is conditional on democratic accountability. The higher the accountability, the less severe is the negative impact of liberalization on inequality. On the contrary, the financial development, proxied by the credit to GDP ratio, has inequality increasing effect irrespective of the institutional background.

% \textcite{furceri2019robust} apply WALS to identify robust determinants of income inequality. They are the closest paper to ours since they account for model uncertainty in the estimation. Their focus if more general rather than on finance specifically. Out work offers more detailed contribution in terms of the role of finance in shaping income inequality. On the top of that, we differ from their analysis by examining multiple measures of inequality and specifically identifying the determinants of top income shares along with the determinants of the overall income distribution represented by income Gini index. We provide further evidence in all these dimensions.

%
%
%
%
%

\section{Data and methodology}
The key variable in the paper is the measure of income inequality. We want to examine how financial development affects income inequality and whether the effect might by different at the top quantiles of income distribution. As the overall measure of income inequality, we rely the after-tax Gini coefficient from \ac{SWIID} by \textcite{Solt2019}, which is a standard resource in the literature\footnote{There are alternative sources of for Gini coefficient, e.g. \ac{WIID} or \ac{LIS}, but each of them brings limitations in terms of comparability or coverage.}. Its critical advantage lies in the widespread coverage across countries and time and a unified methodology which provides a reasonable level of comparability. It typically takes values in the interval between 0 and 100 where the former suggests perfect equality (everyone in the economy enjoys the same income) and the latter perfect inequality (all the income goes to only a single unit). We depart from existing papers slightly in considering the after-tax rather than the before-tax income distribution as a dependent variable.

To explore the relationship in the top part of the distribution, we choose top income share from \ac{WID}\footnote{The methodology and guidelines to database are provided by \textcite{alvaredo2016distributional}.}. The surveys suffer from well-known issues of underrepresentation of the top income earners and the distortions resulting from self-reported character of the data. This can influence not only the top income shares resulting from survey data, but also distortions in the overall measures of inequality. The data in \ac{WID} make use of income tax records in individual countries and the derived shares obtained using consistent methodology of \ac{DINA} are arguably more reliable relative to the survey-based measures which are the primary source of majority estimates of income distributions.

The data spans from 2000 to 2014. We follow the literature \parencite{dabla2015causes,de2017finance} and average both the inequality measure (dependent variable) and the potential determinants (independent variables) across three year intervals. There are important reasons for looking at the averages than observation in individual years. Annual macroeconomic data are subject to fluctuations and the data on income inequality is noisy \textcite{delis2014}. Averaging should diminish the level of noise. On the top of that, the variables at the center of our analysis, e.g. stock market capitalization or credit to \ac{GDP}, are likely to be affected by the business cycles and volatile on the yearly basis. Similar argument holds for top income shares, as they depend, among other things, on the bonuses paid out each year and capital income. We want to explore the long-term rather than the short-term relationship and that guides the choice of averaged data. Faced against the trade-off between length of the averaging periods and available observations in the time dimension, we take a compromise of three years in contrast to the literature, where generally the 5-year intervals apply. The availability of financial development indicators limits the analysis to a period from 2000 onward and we prefer to keep at least 5 unique time periods to just three under the case of 5-year average\footnote{Nevertheless, we run the estimation with 5-year averages of data as a robustness check and find no critical qualitative differences compared to the baseline.}.

Table \ref{ch4tab:sum} report the summary statistics of the income inequality variables and financial development indicators.

% \begin{table}[ht!]
%     \small
%     \caption{Summary statistics of inequality variables}
%     \label{ch4tab:ineq}
%     \centering
%        \begin{tabular}{lrrrr}
%         \toprule
%         Variable & Mean & St. Dev. & Min & Max \\
%         \midrule
%         After-tax Gini index  & & & & \\
%         Top 10\% income share & & & & \\
%         Top 1\% income share & & & & \\
%         \bottomrule
%     \end{tabular}
% \end{table}

\begin{table}[ht!]
  \small
  \caption{Summary statistics of selected variables}
  \label{ch4tab:sum}
  \centering
     \begin{tabular}{lrrrr}
      \toprule
      Variable & Mean & St. Dev. & Min & Max \\
      \midrule
      After-tax Gini index  & 36.38 & 8.12 & 22.88 & 61.16 \\
      Top 10\% income share & 0.42 & 0.12 & 0.24 & 0.71 \\
      Top 1\% income share & 0.14 & 0.06 & 0.05 & 0.38 \\
      FIA & 0.42 & 0.32 & 0.01 & 1.00 \\
      FIE & 0.60 & 0.13 & 0.11 & 0.81 \\
      FID & 0.37 & 0.30 & 0.01 & 1.00 \\      
      FMD & 0.33 & 0.32 & 0.00 & 0.99 \\
      \bottomrule
  \end{tabular}
\end{table}
%
%
%
%
%

\begin{table}[ht!]
  \small
  \caption{Correlation matrix of selected variables}
  \label{ch4tab:corr}
  \centering
  \begin{tabular}{lrrrrrrr}
    \toprule
  %  & GiniNet & Top10share & Top1share & FIA & FIE & FID & FMD \\
    % \midrule
    After-tax Gini index  & . &  &  &  &  &  &  \\
    Top 10\% income share & 0.47 & . &  &  &  &  &  \\ 
    Top 1\% income share & 0.39 & 0.84 & . &  &  &  &  \\
    FIA & -0.14 & -0.09 & -0.12 & . &  &  &  \\
    FIE & -0.14 & -0.10 & -0.06 & 0.28 & . &  &  \\ 
    FID & -0.07 & 0.12 & 0.06 & 0.60 & 0.24 & . &  \\
    FMD & 0.03 & 0.2 & 0.13 & 0.36 & 0.1 & 0.40 & . \\ 
     \bottomrule
  \end{tabular}
  \end{table}

We obtain the financial development indicators from \ac{GFDD}. The database offers detailed indicators along four dimensions of financial systems and allows to estimate the affect of changes in access, size, efficiency, and stability of financial markets. Furthermore, we can distinguish between banking sector and financial markets in all these dimensions. The data in for access and stability of stock markets remains sparse in the concerned period we must leave them out of analysis. We use the version of financial indicators from \cite{svirydzenka2016introducing}. The authors make use of principal component analysis in order to construct aggregate indicators in each characteristic of financial sector. In summary, we have indicators of financial institutions depth (FID), financial markets depth (FMD), access to financial institutions (FIA), efficiency of financial markets (FME), and institutions (FIE)\footnote{\textcite{svirydzenka2016introducing} extrapolate the indicators from top to bottom if the original variables are unavailable, we make sure that that at least one variable is available for the construction of the index and no artificial correlation introduced to the data.}. We report the composition of each indicator in Table \ref{ch4tab:finind}.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.8\textwidth, keepaspectratio]{figures/plots_ineq}
  \caption{Gini Coefficient and top shares}
  \label{ch4fig:gini_topshares}
\end{figure}

% Figure \ref{ch4fig:gini_findev} shows the scatter plots of all the available observations after averaging between Gini coefficient and selected financial indicators. Although the data is noisy, it suggest negative correlation between the variables. Figure \ref{ch4fig:gini_findev_dm} then shows the corresponding scatter plots of the data which was demeaned using the within country averages over time. %The relationship between Gini and financial indicators appears considerably weaker following the transformation, with the exception of access to financial institutions where visual inspection allows pointing to a negative correlation.

We build the choice of other explanatory variables on a reviews of income inequality drivers \parencite{roineetal2009,nolan2019drivers}, related study of finance-inequality nexus \parencite{de2017finance}, and a more general inquiry into the robust determinants of income inequality \parencite{furceri2019robust}. The potential regressors could be categorized in several groups. They control for economic and financial development, demographics, globalization, and institutional background. Table \ref{ch4app:exog} reports all the control variables and their sources.

% \textcite{nolan2019drivers} bring a survey of the literature on determinants of inequality, summarizing the complexity of the inequality dynamics. They stress that many of the determinants are interlinked which implies difficulty in assigning precise effects to individual drivers of inequality. Additionally, they encourage complementary individual country case studies to support the finding of the general cross-country estimates.
% Both theoretical and empirical studies leave out the issue of importing the financial services from abroad. We include financial globalization from KOF among our control variables.
%
%
% \section{Methodology}
% \label{ch4sec:methodology}
Methodologically, we rely on the \ac{BMA} approach which conveniently addresses the issue of model uncertainty with a large number of potential regressors. The advantages are statistical properties of the \ac{BMA} have been sketched in \textcite{Koop2003}. In application to the panel data, we make use of Frisch-Waugh-Lovell theorem and demean all the variables using time averages for individual observations. Using the time-demeaned observations in the estimation is equivalent to the estimate using the dummy variables for individual cross-sectional units. The key assumptions in \ac{BMA} are on parameter and model priors. For the parameter prior, we turn to so-called hyper-g prior. The prior provides more robust results than some other traditionally applied $g$ priors such as \ac{UIP}, \ac{BRIC} \parencite{feldkircher2012impact}. As for the model priors, the baseline estimate relies on the uniform model prior. We choose the model prior to remain agnostic about the prior probability of each examined model. While uniform prior assigns the same prior probability to each model, the distribution of the prior model space is concentrated around $k/2$, where $k$ is the number of potential covariates, and it may consequentially gravitate towards larger model sizes and higher number of covariates\footnote{See \textcite{LeySteel2009} for details.}.

\section{Results}
\label{ch4sec:results}
We examine the determinants of inequality in the panel \ac{BMA} framework and present the results in the following sections. We start with a model where we capture the overall inequality by Gini index in subsection \ref{ch4subsec:gini}. We then continue to estimations where we consider the shares of income going to the top 10\% and top 1\% of the income distribution as our dependent variable. We check the robustness of our estimates by employing alternative model and parameter priors throughout the analysis.

\subsection{Gini index of income distribution}\label{ch4subsec:gini}
We focus the analysis on the relationship between the indicators representing various aspects of financial development and income inequality. Figure \ref{ch4fig:gini_findev_dm} outlines the expected link after we have demeaned the variables using the cross-sectional averages. The relationship is not particularly strong, but we observe negative correlation between Gini index and indicators of access and efficiency of financial institutions, as suggested by a linear estimate. For size indicators of financial market and institutions, the link appears much weaker.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.8\textwidth, keepaspectratio]{figures/plots_findev_gini_dm}
  \caption{Gini Coefficient and Financial Development Indicators}
  \label{ch4fig:gini_findev_dm}
\end{figure}

% \begin{figure}
%   \caption{Gini Coefficient and Financial Development Indicators}
%   \label{ch4fig:gini_findev}
%   \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini}
% \end{figure}

Table \ref{ch4res:baseline_gini} reports the baseline results. Overall, we have 16 variables with \ac{PIP} above 0.8. The number of unique relevant regressors effectively shrinks by two if we abstract from the quadratic terms of the \ac{GDP} per capita and the education index. Most of the estimated posterior means exhibit expected signs.

The only financial indicators which occur among the top regressors are access and efficiency of financial institutions with \acp{PIP} of 1 and 0.88, respectively. The posterior mean on the coefficients in both dimension is negative so higher levels of access and efficiency are associated with lower levels of income inequality. The inequality decreasing effect of access to finance on inequality mirrors \textcite{hasan2020finance} for wealth inequality, and partially also \textcite{furceri2019robust, naceurzhang2016} who document similar effect. The observation on inequality decreasing effect of access to finance supports also \textcite{claessens2007finance} who suggest that access may equalize economic opportunities and lead to more evenly distributed income as well as theoretical predictions \parencite{braunetal2019,galormoav2004,banerjeenewman1990}. The efficiency of financial intermediation putting downward pressure on income inequality also has a precedent in \textcite{gimet2011closer}\footnote{The authors measure efficiency of banking sector by the difference between lending rate and the deposit rate (spread). They argue higher spread reflects low competition and high transaction costs. The imperfections in the credit market can skew the credit to high-income, rich households who can provide significant collateral, reinforcing the existing inequality.}. We fail to confirm that the efficiency of financial institutions is a robust determinant of inequality, however, as the \ac{PIP} markedly decreases under alternative model priors. None of the size indicators of financial institutions or markets has high probability of inclusion.

Education expenditures (\% share of \ac{GDP}) along with the education index calculated using mean and expected years of schooling show inequality decreasing effect. This is in line with prediction of \textcite{goldin2009race,deaton2013great} who claim that the skill-biased technological change should be mitigated by education. \textcite{oecd2011divided} finds similar evidence for a panel of advanced countries while \textcite{furceri2019robust} also suggest a negative relationship, although it is not fully robust to variable and sample selection. The effect of education diminishes at the higher levels of schooling as the quadratic term has also high \ac{PIP} and positive posterior mean.

\begin{table}[ht!]
  \caption{BMA, baseline results. Dependent variable after-tax Gini index, 394 observations.}\label{ch4res:baseline_gini}
  \footnotesize
  \centering
  \begin{tabular}{lrrr}
    \toprule
                           & PIP & Post Mean & Post SD \\
    \midrule
    Education expenditures & 1.00 & -0.14506 & 0.05042 \\
  GDP per capita & 1.00 & 1.62811 & 0.56870 \\
  Unemployment & 1.00 & 0.23629 & 0.05907 \\
  Non-equipment investment & 1.00 & 0.14977 & 0.05285 \\ 
  Access to financial institutions & 1.00 & -0.24629 & 0.06342 \\
  Education index (UN) & 1.00 & -0.58853 & 0.23865 \\ 
  Redistribution & 1.00 & -0.22055 & 0.05051 \\
  Education index sq. & 1.00 & 0.82252 & 0.22008 \\
  GDP per capita sq. & 1.00 & -1.44906 & 0.57565 \\
  Life expectancy & 0.99 & -0.22879 & 0.09750 \\ 
  Economic freedom & 0.99 & 0.18518 & 0.06941 \\
  Value added in agriculture & 0.97 & -0.12361 & 0.05947 \\
  Government expenditures & 0.95 & 0.12259 & 0.05743 \\ 
  Total population & 0.95 & -0.15378 & 0.08478 \\
  Financial institutions efficiency & 0.88 & -0.08374 & 0.05599 \\
  Inflation & 0.86 & 0.16879 & 0.12487 \\
  \midrule
  Inflation sq. & 0.66 & -0.11127 & 0.11201 \\ 
  Value added in industry & 0.62 & -0.05825 & 0.06407 \\
  Equipment investment & 0.44 & -0.03383 & 0.05359 \\ 
  Financial markets depth & 0.27 & 0.01554 & 0.03673 \\ 
  Gross domestic savings & 0.27 & -0.02022 & 0.04597 \\
  % Period 4 & 0.25 & 0.01282 & 0.03330 \\ 
  Social globalization & 0.23 & 0.02365 & 0.06279 \\
  Restrictions on globalization & 0.15 & 0.00874 & 0.03214 \\
  Trade openness & 0.13 & 0.00615 & 0.02589 \\
  Left-wing orientation & 0.11 & 0.00339 & 0.01720 \\ 
  Rule of law & 0.10 & -0.00333 & 0.01812 \\ 
  Net FDI (\% GDP) & 0.10 & -0.00283 & 0.01575 \\ 
  Natural resources rents & 0.08 & -0.00223 & 0.01583 \\ 
  Chinn-Ito index & 0.07 & -0.00205 & 0.01780 \\ 
  % Period 5 & 0.07 & 0.00025 & 0.02188 \\ 
  % Period 2 & 0.07 & 0.00133 & 0.01321 \\
  % Period 3 & 0.06 & -0.00027 & 0.01423 \\
  Financial globalization & 0.06 & -0.00134 & 0.01466 \\
  Civil liberties \& political rights & 0.06 & 0.00075 & 0.01131 \\ 
  Financial institutions depth & 0.05 & 0.00060 & 0.01316 \\ 
  GDP growth & 0.05 & 0.00016 & 0.01255 \\
  Population growth & 0.05 & 0.00055 & 0.01052 \\ 
  Political globalization & 0.05 & 0.00000 & 0.01432 \\  
    \bottomrule
  \end{tabular}
  \end{table}

We find evidence for the Kuznets' hypothesis about the inequality and economic development. We include three regressors which primarily capture the level of development - GDP per capita along with a square term, share of value added in agriculture, and share of value added in industry. Exploring the results of the baseline, We find evidence for an inverted-U curve for GDP per capita and baseline \acp{PIP} for linear and square component. This suggests that the inequality tends to be higher at the lower stages of economic development and only starts decreasing after reaching a threshold. The \ac{PIP} of 0.97 for value added in agriculture along with a negative posterior mean for its coefficient supports this idea further. The economies at lower stages of development generally report higher shares of agricultural sector which is also more labour intensive, i.e. not exacerbating the inequalities that stem from unequal distribution of capital. The level of unemployment is associated with elevated income inequality. The mechanism is direct through the lower share of income going to labour in case of high unemployment rates. The effect of unemployment is also documented on cross-section by \textcite{furceri2019robust}. We also explore the effect of non-equipment investment\footnote{We construct the indicator using the detail on capital investment prom \ac{PWT} and split the overall investment into non-equipment (structures, transportation, and other assets - software / intellectual property products) and equipment investment (machinery, computers + communication equipment).}, which we believe could proxy for the technological progress and skill-biased technological change. The \ac{PIP} suggests positive correlation with income inequality as \textcite{goldin2009race,dabla2015causes}.

We rely on redistribution to capture the effect of taxes and social expenditure on income inequality. Our measure is the the difference between before-tax and after-tax Gini coefficient from the \ac{SWIID} database. Therefore, we concentrate on the direct effects of policies on the household income. Given the global nature of our data and its limitations, we cannot estimate the potential indirect effects on the pre-tax distribution of income using, for example, top marginal tax rates \parencite{alvaredoetal2013}, corporate tax rates \parencite{fuest2018higher}. Nevertheless, the \ac{PIP} of our indicator is very high and it it negatively relates to inequality as expected. The government expenditures, an often used regressors in the literature also has a perfect inclusion probability, but with positive posterior mean. The intuitive first-order effect of government expenditure should be through reduction of inequality through general social spending on transfers, education, and health. We account for these explicitly with redistribution and education expenditures of the government. \textcite{anderson2017does} introduce a meta-analysis where they show how the examined relationship depends on the type of spending considered and how it depends on the measure of income inequality. Additionally, they suggest that the redistributive impact has not extended over the entire distribution, but has mostly centered towards middle-income groups. In line with their conclusions, our results suggest the effect of government spending might run in the positive direction when the key equalizing function have been accounted for or perhaps point towards reverse causality.

Life expectancy should proxy for changing demographics. We find a negative link between life expectancy and income inequality. \textcite{goldsteinlee2014} examine the channels between population ageing and inequality. Stretching the economic life-cycle are associated with larger within-group variance as a cohort ages. They argue this should pronounce inequality in older populations. On the other hand, if we consider retirement age, pension structure in many countries equalize income flows and therefore the older populations may report lower levels of inequality. The posterior mean of our estimated coefficient indicates the latter scenario.

Inflation is the only variable which could proxy for monetary policy and its influence on income inequality. It is among the relevant regressors with \ac{PIP} of 0.86 and shows inequality inducing association. The interest in the effects of monetary policy, and macroprudential tools in particular, on inequality is relatively recent, following the introduction of unconventional measures following the Great Recession \parencite{frost2018macroprudential}. Theoretically, the effect is ambiguous as high inflation of asset prices might benefit high-income households than the low income as they receive higher shares of financial income. Also, the assets held by low-income households tend to be much more liquid, thus inflation hurts them more relatively to the high-income ones. On the contrary, cut in rates usually benefits the borrowers and increased economic activity benefits the bottom-part of income distribution \parencite{furceri2019robust}. Our evidence supports the view of inflation enhancing income inequality, at least at moderate levels, as its square term is borderline on the inclusion, but negative in terms of posterior mean of its coefficient. This would suggest that the above mentioned theoretical effects interact and manifest differently at different inflation levels.

The index of the economic freedom of the world describes the overall business conditions by considering regulatory and legal environment. More economic freedom to trade and run daily business makes it easier to exploit economic opportunities and the high \ac{PIP} with positive posterior mean of the coefficient point to economic freedom making the distribution of income less equal. We consider many other potential explanatory variables, but in case of Gini index, they show low \acp{PIP}. We do not find any measure of globalization, institutions, or trade openness relevant for the overall distribution of income. In the next section, we concentrate on the top income share and whether the top part of the distribution is perhaps associated with different regressors.

\subsection{Top income shares}
Financial development may influence various part of the income distribution differently. We therefore follow with baseline estimates with the top income shares as dependent variables. Table \ref{ch4res:baseline_top10} reports the results for top 10\% share and Table \ref{ch4res:baseline_top1} for the top 1\% share. There is large overlap of the most important determinants for the top 10\% income share and Gini coefficient. GDP per capita, life expectancy, government expenditures, education index (but not expenditures), and inflation remain among the relevant regressors and their posterior means are consistent with the estimation for the overall income distribution. However, \acp{PIP} of some of the variables decline significantly and some other become relevant. Figure \ref{ch4fig:comp_ineq_sel} provides a visual comparison of the inclusion probabilities.

\begin{table}[ht!]
  \caption{BMA, baseline results. Dependent variable Top 10\% share, 394 observations.}\label{ch4res:baseline_top10}
    \footnotesize
    \centering
  \begin{tabular}{lrrr}
      \toprule
                              & PIP & Post Mean & Post SD \\
      \midrule
  Natural resources rents & 1.00 & -0.15595 & 0.04773 \\
  GDP per capita & 1.00 & 2.13915 & 0.54791 \\
  Life expectancy & 1.00 & -0.53327 & 0.09806 \\
  Financial institutions depth & 1.00 & 0.20288 & 0.05638 \\
  Access to financial institutions & 1.00 & -0.25218 & 0.06712 \\
  Financial markets depth & 1.00 & 0.19376 & 0.04976 \\
  GDP per capita sq. & 1.00 & -1.80042 & 0.55408 \\ 
  Government expenditures & 1.00 & 0.15638 & 0.05565 \\ 
  Rule of law & 1.00 & -0.12607 & 0.04665 \\
  Education index (UN) & 1.00 & -0.49573 & 0.22152 \\ 
  Education index sq. & 1.00 & 0.58180 & 0.20423 \\
  Equipment investment & 1.00 & -0.13953 & 0.05578 \\ 
  Trade openness & 0.98 & 0.12632 & 0.05604 \\
  Inflation & 0.80 & 0.06758 & 0.05407 \\ 
  \midrule
  Gross domestic savings & 0.65 & 0.06566 & 0.06764 \\
  Left-wing orientation & 0.63 & 0.04567 & 0.04858 \\ 
  Total population & 0.57 & 0.07224 & 0.08621 \\ 
  Financial institutions efficiency & 0.55 & -0.04309 & 0.05278 \\
  Non-equipment investment & 0.49 & 0.03864 & 0.05320 \\ 
  Value added in industry & 0.31 & -0.02652 & 0.05253 \\
  Financial globalization & 0.22 & -0.01390 & 0.03673 \\ 
  Value added in agriculture & 0.18 & -0.01163 & 0.03405 \\ 
  Unemployment & 0.18 & -0.01019 & 0.03113 \\
  Restrictions on globalization & 0.13 & 0.00637 & 0.02398 \\ 
  Civil liberties \& political rights & 0.12 & 0.00529 & 0.02137 \\
  GDP growth & 0.12 & 0.00608 & 0.02464 \\ 
  Net FDI (\% GDP) & 0.07 & -0.00217 & 0.01377 \\ 
  Redistribution & 0.07 & -0.00243 & 0.01571 \\
  Chinn-Ito index & 0.06 & 0.00170 & 0.01373 \\ 
  % Period 2 & 0.06 & -0.00160 & 0.01267 \\
  % Period 5 & 0.05 & 0.00158 & 0.01431 \\ 
  Population growth & 0.05 & 0.00097 & 0.01043 \\
  Economic freedom & 0.04 & 0.00055 & 0.01416 \\ 
  Education expenditures & 0.04 & -0.00062 & 0.01030 \\
  Social globalization & 0.04 & -0.00004 & 0.01918 \\ 
  Political globalization & 0.04 & 0.00060 & 0.01279 \\
  % Period 3 & 0.03 & -0.00007 & 0.00873 \\ 
  % Period 4 & 0.03 & -0.00017 & 0.00827 \\
  Inflation sq. & 0.03 & -0.00045 & 0.01581 \\
  \bottomrule
  \end{tabular}
\end{table}

In the case of the Top 10\% share, education expenditures, non-equipment investment, unemployment, redistribution, and an index of economic freedom drop out. With the exception of the economic freedom, there are good reasons to believe these factors mainly drive lower and middle-part of the income distribution, rather than the very top. While education expenditures may support public education system and allow for human capital accumulation in the low-income households, such effect might not be as relevant for the concentration at the top. We have discussed previously how redistribution policies mostly affect the middle quantiles of the distribution and high unemployment rates traditionally do not concern the well educated high-income households and individuals. Also share of value added in agriculture now has a low \ac{PIP}. The inclusion of non-equipment investment also decreases, while it retains the positive posterior mean. Most importantly, depth of financial markets and financial institutions now exhibit perfect \acp{PIP} and they seem to be associated with income distributions more concentrated at the top. The access to financial institutions remains relevant and negatively correlated with inequality. Among other variables with higher \acp{PIP}, we have natural resources rents, rule of law, equipment investment, and trade openness. For the natural resources rents, we get a negative posterior mean. The evidence is in line with \textcite{goderismalone2011}, who describe a mechanism of income equalizing natural resource booms through the benefits for unskilled workers in labour intensive sector. A positive link between inequality and trade openness has been suggested by \textcite{Jaumotte2013,dabla2015causes} for the cross-sectional datasets and by \textcite{milanovicvan2018inequality} in the case of \ac{US}. The negative posterior coefficient on the rule of law is in line with the prediction by \textcite{perotti2007investor} and is sole variable indicating the importance of institutions posited in \textcite{acemoglu2003cross,acemoglu2015rise}. Equipment investment (machinery) may channel the resources to the bottom part of the income distribution through increase long-term growth rates and upward pressure on wages.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.8\textwidth, keepaspectratio]{figures/comp_baseline_pip_sel}
  \caption{\acp{PIP} with different inequality measures}
  \label{ch4fig:comp_ineq_sel}
  \begin{minipage}{0.8\textwidth}
    \footnotesize
    \emph{Note: The comparison only shows variables which show \ac{PIP} $> 0.9$ in at least one of the models.}
    \end{minipage}
\end{figure}

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\textwidth, keepaspectratio]{figures/topsh_comparison_all}
  \caption{\acp{PIP} for different top income shares with baseline priors}
  \label{ch4fig:comp_topshares}
\end{figure}

\begin{table}[ht!]
  \caption{BMA, baseline results. Dependent variable Top 1\% share, 394 observations.}\label{ch4res:baseline_top1}
    \footnotesize
    \centering
  \begin{tabular}{lrrr}
    \toprule
                                 & PIP & Post Mean & Post SD \\
    \midrule
  Rule of law & 1.00 & -0.17039 & 0.04688 \\
  GDP per capita & 1.00 & 0.48163 & 0.29090 \\
  Life expectancy & 1.00 & -0.45605 & 0.06980 \\
  Equipment investment & 1.00 & -0.19727 & 0.05526 \\
  Financial institutions depth & 1.00 & 0.16380 & 0.05689 \\ 
  Access to financial institutions & 1.00 & -0.25945 & 0.06419 \\
  Financial markets depth & 1.00 & 0.15454 & 0.04989 \\ 
  Government expenditures & 0.99 & 0.11633 & 0.04718 \\ 
  Inflation & 0.99 & 0.11773 & 0.04922 \\
  Trade openness & 0.93 & 0.10374 & 0.05711 \\ 
  \midrule
  Restrictions on globalization & 0.47 & 0.03804 & 0.05418 \\ 
  Non-equipment investment & 0.42 & 0.03197 & 0.05001 \\ 
  Natural resources rents & 0.34 & -0.02182 & 0.04037 \\ 
  Financial globalization & 0.28 & -0.02001 & 0.04269 \\
  Value added in industry & 0.21 & -0.01389 & 0.03634 \\
  Value added in agriculture & 0.17 & -0.00969 & 0.03053 \\ 
  Net FDI (\% GDP) & 0.16 & -0.00730 & 0.02401 \\
  Chinn-Ito index & 0.15 & 0.00828 & 0.02797 \\
  GDP per capita sq. & 0.15 & -0.08401 & 0.28484 \\
  Economic freedom & 0.12 & -0.00945 & 0.03762 \\ 
  Gross domestic savings & 0.11 & 0.00634 & 0.02688 \\
  Redistribution & 0.11 & 0.00495 & 0.02085 \\
  GDP growth & 0.11 & -0.00504 & 0.02208 \\
  Social globalization & 0.09 & -0.00626 & 0.03233 \\
  Education index (UN) & 0.09 & -0.01003 & 0.07309 \\ 
  % Period 3 & 0.07 & -0.00214 & 0.01452 \\
  % Period 2 & 0.06 & -0.00146 & 0.01249 \\
  Unemployment & 0.05 & 0.00131 & 0.01287 \\ 
  % Period 5 & 0.05 & 0.00140 & 0.01362 \\
  Political globalization & 0.05 & 0.00137 & 0.01473 \\ 
  Financial institutions efficiency & 0.04 & -0.00070 & 0.01024 \\ 
  Education index sq. & 0.04 & 0.01380 & 0.07857 \\ 
  Total population & 0.04 & -0.00060 & 0.01484 \\ 
  Inflation sq. & 0.04 & 0.00053 & 0.01828 \\ 
  Population growth & 0.04 & 0.00020 & 0.00844 \\
  % Period 4 & 0.04 & 0.00019 & 0.00892 \\ 
  Education expenditures & 0.04 & -0.00002 & 0.00935 \\ 
  Left-wing orientation & 0.04 & -0.00028 & 0.00843 \\ 
  Civil liberties \& political rights & 0.04 & 0.00009 & 0.00846 \\
    \bottomrule
  \end{tabular}
\end{table}

The results using the Top 1\% share for top regressors differs in the drop of \acp{PIP} for natural resource rents, and education attainment. We hypothesize that natural resource rents might redistribute the share of income from the top decile to the rest of the distribution without actually affecting the share of the income going to the very top. The top 1\% income earners might well be among the ones enjoying capital rents from the country resources. While education appears to have equalizing income effect for the lower part of the distribution, it is reasonable that it wears out for the very top income earners in the distribution. Figure \ref{ch4fig:comp_topshares} depicts comparison of inclusion probabilities of all variables considered in the estimation for the Top 10\% share and Top 1\% share.

\subsection{Robustness checks}
We perform robustness checks for all three specifications employing alternative model priors\footnote{We also perform robustness checks using alternative hyperparameter $g$ and Markov-chain samplers, but these do not affect our results.}. We choose random and dilution priors to address prior model size concentrated around the mean number of potential regressors and correlation among them by each of the model priors, respectively.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\textwidth, keepaspectratio]{figures/model_priors_comparison_gini}
  \caption{Robustness checks with alternative model priors, Gini coefficient}
  \label{ch4fig:gini_comp}
\end{figure}

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\textwidth, keepaspectratio]{figures/model_priors_comparison_top10}
  \caption{Robustness checks with alternative model priors, Top 10\% share}
  \label{ch4fig:top10_comp}
\end{figure}

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\textwidth, keepaspectratio]{figures/model_priors_comparison_top1}
  \caption{Robustness checks with alternative model priors, Top 1\% share}
  \label{ch4fig:top1_comp}
\end{figure}

The alternative priors influence the results similarly in all three specifications. The random mode prior decreases \acp{PIP} for the regressors as it generally prefers smaller models. Nevertheless, the results are only marginally effected for Gini coefficient as well as the top income shares with a few exception of education index in the case of Top 10\% share and trade openness for both top income shares. The use of dilution model prior which penalizes the models with highly correlated regressors, we see more significant differences. Above all, the quadratic terms show very low inclusion probabilities. That is not surprising since they are correlated with their original values. This irrelevance of quadratic terms is universal across inequality measures and we may argue that it is due to the construction of the concerned variables. While the \acp{PIP} for other regressors remain similar to the baseline in the case of top income shares, for the Gini index of overall income inequality, we observe some regressors which are now penalized by the dilution prior. Taking the set of the important regressors in the baseline specification, we observe drop in the inclusion probability for education index, life expectancy, and economic freedom. Given high correlations among the set of regressors, this is not surprising and allows to narrow down the number of regressors robustly associated with income inequality.

%
%
%
%
%

\section{Conclusion}
\label{ch4sec:conclusion}
In the paper, we explore the robust determinants of income inequality with a special attention given to indicators of financial development. We choose financial indicators that reflect the access, efficiency, and size of financial markets and institutions. We believe that the detailed indicators provide better proxies for the functions of finance - screen investment opportunities, monitor the debtors who were provided funding, as well as pooling and management of risk. We allow for model uncertainty by employing \ac{BMA} and examine a number of other potential determinants of income inequality.

We show that financial development has complex relationship on income distribution. While access to finance has a universally inequality decreasing effect, the larger size of financial markets and financial institutions associates with higher top income shares, irrelevant for the Gini coefficient of the income distribution. This finding suggests that the size of financial markets also equalizes the income among the first 9 deciles.

We find a few other important covariates for income inequality. Education, redistribution, and changing demographic structure seem to be linked with lower income inequality. On the contrary, unemployment, investment other than machinery, economic freedom - ease of pursuing economic opportunities, and inflation all positively relate to income inequality. As in the case of financial indicators, we find that the link could be more complex. When looking at the top income shares, some of the covariates (education, unemployment, redistribution, and economic freedom) cease to be relevant, while some other (trade openness, rule of law, and machinery investment) seem to matter.

The results we present warrant caution in drawing quick conclusions on the matter of income inequality determinants. While finance, technology, and trade likely affect the distributional outcomes, it can have varying influence on different parts of the income distribution.

% Finance captures the capacity of financial intermediaries and markets to screen investment opportunities, monitor the debtors who were provided funding, as well as pooling and management of risk. With inequality, the focus of the paper is on the inequality in the distribution of income. Arguably, the literature discusses other concepts of inequality, e.g. intergenerational persistence of relative income differences or equality of opportunity \parencite{demirgucc2009finance}.

% \textcite{kroszneretal2007} show that financial crises have relative more severe impact on the sectors which depend more on external financing. The consequences of crises on firms relate to institutional environment and materialize through lower production capacity and competition.

% Overall, these results suggest that policies seeking to improve the governance and robustness of local banks should be prioritized over size-enhancing reforms from a normative point of view \parencite{gimet2011closer}.

% The benefits of capital markets liberalization seem to be concentrated to the top of the income distribution. Top quintile of the distribution accrues nearly all of the income growth following the liberalization while the share of middle three quantiles decreases and the bottom remains unaffected \parencite{das2003income}.

\clearpage
%
%
%
%
%
%
%
%
%
%
\newpage
\printbibliography
\clearpage
%
%
%
%
%
\appendix
\renewcommand{\thesection}{A\arabic{section}}%
\renewcommand{\thetable}{A\arabic{table}}%
\renewcommand{\thefigure}{A\arabic{figure}}%
\renewcommand{\theequation}{A\arabic{equation}}%
\setcounter{equation}{0}%
\setcounter{table}{0}%
\setcounter{figure}{0}%

\section{Appendix}
\label{ch4sec:appch2}

\subsection*{The composition of financial indicators}
\label{ch4subsec:finind_comp}
\begin{table}[ht!]
    \small
    \caption{Underlying Components of Financial Development Indicators}
    \label{ch4tab:finind}
    \centering
    \begin{tabular}{ll}
      \toprule
      \textsc{Indicator} & \textsc{Measure} \\
      \midrule
      \multicolumn{2}{l}{\textbf{Financial institutions}} \\
      \midrule
      \multirow{2}{*}{Access} 	& Bank branches per 100,000 adults \\
                                  & ATMs per 100,000 adults \\
      \midrule
      \multirow{6}{*}{Efficiency}		& Net interest margin \\
                                  & Lending-deposits spread \\ 
                                  & Noninterest income to total income \\
                                  & Overhead costs to total assets \\
                                  & Return on assets \\
                                  & Return on equity \\

      \midrule
      \multirow{4}{*}{Depth}	& Domestic private credit to the real sector to the GDP \\
                                  & Pension fund assets/GDP \\
                                  & Mutual fund assets/GDP \\
                                  & Insurance premiums life and nonlife/GDP \\
      \midrule
      \multicolumn{2}{l}{\textbf{Financial markets}} \\
      \midrule
      \multirow{6}{*}{Depth} 	& Stock market capitalization/GDP \\
                                  & Stocks traded/GDP \\
                                  & International debt securities of government/GDP \\
                                  & Total debt securities of financial corporations/GDP \\
                                  & Total debt securities of nonfinancial corporations/GDP \\
      \bottomrule
    \end{tabular}
\end{table}

\clearpage
\subsection*{Dataset description}
\begin{center}
\footnotesize
\begin{longtable}{l p{0.50\linewidth} p{0.3\linewidth}}
    \caption{List of variables}
    \label{ch4app:exog} \\
    \toprule 
  Variable & Definition (+ optional comments) & Source \\
    \midrule 
  GiniNet & After\-tax Gini index based on distribution of income (The Standardized World Income Inequality Database). & \href{http://fsolt.org/swiid/}{Solt (2019)} \\
  GiniMarket & Before-tax Gini index based on distribution of income  (The Standardized World Income Inequality Database). & \href{http://fsolt.org/swiid/}{Solt (2019)} \\	
  Top10share & Share of income going top decile of the distribution. & \href{https://wid.world/}{WID} \\	
  Top1share & Share of income going top percentile of the distribution. & \href{https://wid.world/}{WID} \\
  FIA & Access to financial institutions & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\textcite{svirydzenka2016introducing}} \\
  FID & Financial institutions depth & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\textcite{svirydzenka2016introducing}} \\
  FIE & Financial institutions efficiency & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\textcite{svirydzenka2016introducing}} \\
  FMD & Financial markets depth & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\textcite{svirydzenka2016introducing}} \\
  FME & Financial markets efficiency & \href{http://data.imf.org/?sk=F8032E80-B36C-43B1-AC26-493C5B1CD33B}{\textcite{svirydzenka2016introducing}} \\
  GDPpc & Level of GDP per capita & \href{http://data.worldbank.org/indicator/NY.GDP.PCAP.PP.KD}{WB} \\
  NatRes & Total natural resource rents are the sum of oil rents, natural gas rents, coal rents (hard and soft), mineral rents, and forest rents. & \href{http://data.worldbank.org/indicator/NY.GDP.TOTL.RT.ZS}{WB} \\
  PopGrowth & Annual population growth 1980-2009 & \href{http://data.worldbank.org/indicator/SP.POP.GROW}{WB} \\
  GovExp & General government final consumption expenditure (formerly general government consumption). & \href{http://data.worldbank.org/indicator/NE.CON.GOVT.ZS}{WB}\\
  NNSavings & Net national savings (gross national savings less the value of consumption of fixed capital, \% GNI). & \href{http://data.worldbank.org/indicator/NY.ADJ.NNAT.GN.ZS}{WB} \\
  EducExp & Education expenditure refers to the current operating expenditures in education, including wages and salaries and excluding capital investments in buildings and equipment.. & \href{http://data.worldbank.org/indicator/NY.ADJ.AEDU.GN.ZS}{WB} \\
  Infl & Inflation as measured by the consumer price index. & \href{http://data.worldbank.org/indicator/FP.CPI.TOTL.ZG}{WB} \\
  VAA & Agriculture, forestry, and fishing value added (\% GDP). & \href{http://data.worldbank.org/indicator/NV.AGR.TOTL.ZS}{WB} \\
  VAI & Industry value added (\% GDP). & \href{http://data.worldbank.org/indicator/NV.IND.TOTL.ZS}{WB} \\
  GFCF & Gross fixed capital formation (\% of GDP). & \href{http://data.worldbank.org/indicator/NE.GDI.FTOT.ZS}{WB} \\
  NetFDI &  Foreign direct investment, net inflows (\% of GDP). & \href{http://data.worldbank.org/indicator/BX.KLT.DINV.WD.GD.ZS}{WB} \\
  GDPgrowth & Annual growth of GDP. & \href{http://data.worldbank.org/indicator/NY.GDP.MKTP.KD.ZG}{WB} \\
  LifeExp & Life expectancy at birth.  & \href{http://data.worldbank.org/indicator/SP.DYN.LE00.IN}{WB}\\
  LabForce & Total labor force comprises people ages 15 and older who meet the International Labor Organization definition of the economically active population: all people who supply labor for the production of goods and services during a specified period. Labor force total. & \href{http://data.worldbank.org/indicator/SL.TLF.CACT.ZS}{WB} \\
%   PopDens & Population density (people per sq. km of land area). & \href{http://data.worldbank.org/indicator/EN.POP.DNST}{WB} \\
  RuleOfLaw & Rule of law estimate & \href{http://data.worldbank.org/indicator/RL.EST}{WB} \\
  CLandPR & Average of index for civil liberties and political rights & \href{https://freedomhouse.org/report/freedom-world-2016/methodology}{Freedom House} \\
%   OutwardO & Trade (\% of GDP)  & \href{http://data.worldbank.org/indicator/NE.TRD.GNFS.ZS}{WB} \\
  ChinnIto & Chinn-Ito index of financial openness. & \href{http://web.pdx.edu/~ito/Chinn-Ito_website.htm}{Chinn-Ito} \\ 
  LeftWing & Dummy equal to 1 when left oriented party lead the country. & \href{http://www.nsd.uib.no/macrodataguide/set.html?id=11&sub=1}{DPI} \\
  ActivRestrict & Activity restrictions. Regulatory restrictions on bank activities and the mixing of banking and commerce. & \href{http://faculty.haas.berkeley.edu/ross_levine/regulation.htm}{Barth et al. (2013)} \\
  CapitalReg & Capital Regulatory index. & \href{http://faculty.haas.berkeley.edu/ross_levine/regulation.htm}{Barth et al. (2013)} \\
  DiversIndex & Whether there are explicit, verifiable, quantifiable guidelines for asset diversification and banks are allowed to make loans abroad. & \href{http://faculty.haas.berkeley.edu/ross_levine/regulation.htm}{Barth et al. (2013)} \\
  EducIndex & Calculated using mean years of schooling and expected years of schooling & \href{http://hdr.undp.org/en/content/education-index}{UN} \\
  NetInterestMargin & Accounting value of banks' net interest revenue as a share of average interest-bearing assets; a measure of the efficiency of the banking sector. & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
  BankZScore & return on banks' assets plus the ratio of banks' equity and assets, divided by the standard deviation of the return on assets (ROA+equity/assets)/sd(ROA); a measure of stability of the banking sector & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
  Privatecredit & Domestic private credit to the real sector to GDP; a measure of the depth of the banking sector & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
  MarketCap & Value of listed shares to GDP; a measure of the depth of stock markets.& \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
  MarketTurn & Stock market value traded to total market capitalization; a measure of the efficiency of stock markets. & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
  BankBranches & Number of bank branches per 100,000 adults & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
  Loan2Deposits & Loan-to-deposit ratio. & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
  Redist & Difference between market (pre-tax) and net (after-tax0 Gini index based on distribution of income  (The Standardized World Income Inequality Database). & \href{http://fsolt.org/swiid/}{Solt (2016)} \\	
  FinLib & Averaged components of Economic Freedom of the World index 3D (freedom to own foreign currency accounts), 4C (black-market exchange rates), 4D (controls of the movement of capital and people), and 5A (credit market regulations). & \href{https://www.fraserinstitute.org/economic-freedom/dataset}{\textcite{gwartney2017}} \\
  \bottomrule
\end{longtable}
\end{center}

\end{document}

### 2020-02-13 Paper/elsarticle.tex
update

@@
+%!TEX program = pdflatex
+%!BIB program = biber
+
 \listfiles
@@
 \makeatother
-
-\usepackage[utf8]{inputenc}
 
@@
 The size of the markets has no relevance for overall income inequality, but relates to the higher top income shares.
-Moreover, unemployment along with investment into non-tangible assets increase income inequality while more redistribution imply lower levels of inequality.
+Moreover, unemployment along with investment into non-tangible assets increase income inequality while more redistribution implies lower levels of inequality.
 \end{abstract}

### 2020-12-16 Notes/notes.tex
201216 update

%!TEX program = xelatex
%!BIB program = bibtex

\documentclass[a4paper,11pt]{article}
\usepackage{graphicx}
\usepackage{fontspec}
\usepackage[round]{natbib} %standard package for bibliography

\begin{document}

\section{Related literature}
% \label{sec:literature}
The research in the area of financial development and income inequality is well established. \citet{demirgucc2009finance}, \citet{claessens2007finance}, and more recently \cite{de2017finance} provide extensive reviews of the topic. The similar theme emerges in all these papers, with the conclusions from both theoretical and empirical contributions provide conflicting prediction about the relationship and well as not unambiguously conclusive empirical results. Tentative conclusions point towards finance tightening the distribution of income. Among other channels, accessing credit allows human capital accumulation and new firm formation and growth equalize economic opportunities. Furthermore, the indirect effect of finance on labour markets through faster economic growth and increased competition diproportionately benefits the low-income individuals.

Conflicting prediction of finance on inequality - intensive vs. extensive margin. Intergenerational persistence might fall with increased economic oportunities while widening the income distribution in every generation.

DIRECT vs. INDIRECT mechanism - access to finance, expanding opportunities of the previsously disadvantaged groups or intensifying the use of financial services by the incumbent individuals and firms. The financial system having second-order effects on inequality by influencing aggregate production and allocation of credit and consequantly the demand for high- vs. low- skilled workers. \citet{braunetal2019} link growth, inequality, and finance in their theoretical model (access to markets inhibits growth).

\citet{delis2014} bank regulations and inequality - nice intro

% Financial development alters how much are economic opportunities depend on the individual skills, family endowments, social status or political connections. Individual depend on financial system to provide loans to start new business, attain education

Both theoretical and empirical studies leave out the issue of importing the financial services from abroad. We include financial globalization from KOF among our control variables.

\citet{pikettyandzucman2014}
\citet{van2018inequality} provide evidence that income inequality in the US has different implications for the future income growth of the rich and the poor. High inequality seems to hurt the prospects of the poor while the top of the distribution is unaffected. The rich thus disproportianately benefit from higher inequality as their subsequent income exhibit faster growth. The authors attribute this effect to the political channel the rich use to lobby in favour of the policies which support their economic interests. Preferences of the rich are ultimately more likely to determine public policy than the preferences of the majority \citep{gilens_page_2014}. High inequality together with a credit constraint and rich driving the political process results in low government spending and lasting inequality.

\citet{marrero2013inequality} introduce two types of inequality - inequality of opportunity and inequality of effort/luck. Applying this to the EU countries and US states, they show that inequality of opportunity (driven by race and parental education ...) is negatively related to growth while the residual "good" inequality is growth inducing.

Income inequality and growth may intersect through varying channels. Accumulation of savings, unobservable effort, and investment project size favour the prediction on growth inducing inequality. Negative impact of inequality on human capital accumulation, entrepreneurial activity, and hence growth provides argument for the opposing view.

The average income inequality rose across OECD by 1.4 percentage points \citep{oecd2013crisis}.

\citet{LawSingh2014} argue that financial development decrease income inequality by allowing the poor to invest in human and physical capital.

\citet{nolan2019drivers} bring a survey of the literature on determinants of inequality, summarizing the complexity of the inequality dynamics. They stress that many of the determinants are interlinked which implies difficulty in assigning precise effects to individual drivers of inequality. Additionally, they encourage complementary individual country case studies to support the finding of the general cross-country estimates.

\subsection{theory}
\textsc{human capital investment} - with imperfect markets restrict the capital accumulation, differences between human (optimal is equal distribution) and physical capital (diminishin returns to physical capital do not materialize at the level of individual ownership of capital because it is not embodies in individuals \cite{demirgucc2009finance}. This implies that equal distribution is not necessarily optimal). \textsc{political consequences} - higher inequality will increase the proportion of population favouring greater redistribution.  \textsc{wage discrimination by specific characteristics, e. g. race}
%
\textsc{investment opportunities, savings}. Minimum investment requirements of fixed costs associated with high-return investments, including entrepreneurship. => higher return to wealthier individuals (e.g. \citet{AtjeJovanovich1993,demel2008qje}).
\textsc{savings behaviour} - if inheritance is a function on income, then rich parents leave larger bequests than the poor and preserve relative differences in incomes. Research indicates that financial development does not alter aggregate savings rates. Is this also valid for more detailed look on poor and rich households? If the effect differs then the distribution of saving is altered by financial development influence the relative income levels across households while keeping the aggregate savings rate unchanged.

\textsc{Summary}

\citet{galormoav2004} provide a theoretical framework of thinking about the link between income inequality and financial development. In their model, the income inequality falls due to the decrease of financial friction which allows for physical and human capital accumulation of the poor. That is the extensive margin of financial development at work.

\citet{banerjeenewman1990} also offer a theoretical setup in favour of the capital accumulation hypothesis due to finance (liberalization as claimed by \citet{de2017finance}) and inequality decreasing effect.

\citet{GreenwoodJovanovic1990} on the contrary present a model where the intensive margin of the financial development is the key force and the benefits of more finance are accrued by the incumbents - primarily the rich - which implies an inequality increasing effect in the early stages of the development. With tim more agents meet the fixed costs of joining the financial intermediaries and they enjoy higher returns. The efficiency of resource allocation increases, which enhances growth and reduces inequality.

An exemption from focus solely on the size of financial sector is the study by \citet{naceurzhang2016} which applies a similar approach to ours, taking into consideration the access, efficiency, and stability of the financial sector. However, there are severe limitation to their study. First, their coverarage does not correspond to the availability of the data. Second, they do not account for the different dimensions of finance simultaneously and always use a single indicator of development at a time. Third, they do not explicitly differentiate between the banking sector and financial markets. We provide a more detailed as well as more robust picture of the financial development effect on finance in these dimensions.

\citet{gwartney2017} provide an economic freedom index of the Fraser Institute.

\citet{de2017finance} examine different dimensions of finance on income inequality. Their results suggest that financial development, financial liberalization, and banking crises all increase market income inequality within countries. Additionally, they show that the effect of financial liberalization is conditional on democratic accountability. The higher the accountability, the less severe is the negative impact of liberalization on inequality. On the contrary, the financial development, proxied by the credit to GDP ratio, has inequality increasing effect irrespective of the institutional background.

\citet{claessens2007finance} explore the connectedness of growth, finance, and inequality, stressing the importance of institutional background in the effects.

\citet{perotti2007investor} present a theoretical framework how established interests lobby for lower investor protection to prevent entrance of the new competitors. Greater accountability of the politicians leads to a higher bribe required from the lobbyists'. Investor protection strengthens and increases market entry and competition. They examine their prediction in a cross-section and show that better investor protection correlates with larger entry rates and higher firm density in more financially intensive sectors. Furthermore, they show that the most important factor of accountability is not the formal measure of democratic institutions, but \emph{newspaper readership} which they interpret as broad awareness of policy choices and their outcomes. \textsc{size does not matter for entry when investor protection is accounted for}.

Summary of the theoretical preditions in \citet{demirgucc2009finance} on pages 20-23 for the review.

\citet{furceri2019robust} apply WALS to identify robust determinants of income inequality. They are the closest paper to ours since they account for model uncertainty in the estimation. Their focus if more general rathar than on finance specifically. Out work offers more detailed contribution in terms of the role of finance in shaping income inequality. On the top of that, we differ from their analysis by examining multiple measures of inequality and specifically identifying the determinants of top income shares along with the determinants of the overall income distribution represented by income Gini index. We provide further evidence in all these dimensions.

\subsection{empirics}
\citet{beck2007finance} show that the development of financial intermediaries is accompanied by faster reductions in income inquality and poverty rates.
\citet{beck2010big} document how bank deregulation in the US tightened income distribution by benefiting primarily the bottom 50\% of earners. They attribute the effect to the changes in labour market conditions and relatively higher wages and working hours of the low-skilled workers following the reforms.

\citet{claessens2007finance} argue that althugh the deeper financial systems generally provide better opportunities of access to finance, the relationship is not universal.

The benefits of capital markets liberalization seem to be concetrated to the top of the income distribution. Top quintile of the distribution accrues nearly all of the income growth following the liberalization while the share of middle three quantiles decreases and the bottom remains unaffected \citep{das2003income}.

Diwan (1999) show how labour shares fall follwing a crises only partially recovering in the subsequent years.

\citet{kroszneretal2007} show that financial crises have relative more severe impact on the sectors which depend more on external financing. The consequences of crises on firms relate to institutional environment and materialize through lower production capacity and competition.

\citet{glaeseretal2003} model intitutional subversion by the rich classes and their ability to prevent institutional reforms. Such environment often leads to massive redistribution from the rich to the poor, which is assumed to be less efficient than carefully implemented institutional reforms.

\clearpage
%
\bibliographystyle{chicagoa}
\bibliography{literature}
%

\end{document}

### 2020-12-16 Paper/financeinequality.tex
201216 update

--- 2020-02-13 Paper/elsarticle.tex
@@
   We investigate the impact of financial development on income inequality differentiating between depth, efficiency and access to financial markets and institutions.
-We apply panel Bayesian model averaging framework to address model uncertainty to reveal that financial development has complex relationship   with the income distribution within countries.
+We apply panel Bayesian model averaging framework to address model uncertainty to reveal that financial development has a complex relationship with the income distribution within countries.
 The access to and efficiency of banking decrease income inequality.
-The size of the markets has no relevance for overall income inequality, but relates to the higher top income shares.
-Moreover, unemployment along with investment into non-tangible assets increase income inequality while more redistribution implies lower levels of inequality.
+The size of the markets has no relevance for overall income inequality but relates to the higher top income shares.
+Moreover, unemployment, along with investment into non-tangible assets increase income inequality while more redistribution implies lower levels of inequality.
 \end{abstract}
@@
 \section{Introduction}
-\label{ch4sec:intro}
+\label{sec:intro}
 Financial development alters how much the economic opportunities depend on the individual skills, family endowments, social status or political connections as individual may depend on financial system to provide loans to start new business, attain education, or temporarily fund their consumption.
 The research in the area of financial development and income inequality is well established.
-\textcite{demirgucc2009finance}, \textcite{claessens2007finance}, and more recently \cite{de2017finance} provide extensive reviews of the topic.
-A similar theme emerges in all three papers.
-The implications from theoretical contributions provide conflicting predictions about the relationship and empirical results bring evidence for both positive and negative effect.
-Although majority of the papers point towards finance tightening the distribution of income this results is not universal with some papers suggesting the opposite while other stress potential non-linearities.
+\textcite{demirgucc2009finance}, \textcite{claessens2007finance}, and more recently \textcite{de2017finance} provide broad reviews of the topic.
+A similar theme emerges, noting that the implications from theoretical contributions provide conflicting predictions about the relationship.
+The empirical results then bring evidence for both positive and negative effect.
+Although the majority of the research points towards finance tightening the distribution of income, this result is not universal with some papers suggesting the opposite while other stress potential non-linearities.
 
-A key divide appears between the effect of financial development on extensive and intensive margin.
-The extensive margin captures the extend to which individuals, who had not been using financial services before, gain access.
-On the other hand, the intensive margin describes growing use of finance by the agents who had already been using it before \parencite{demirgucc2009finance}.
+A fundamental divide appears between the effect of financial development on the extensive and intensive margin.
+The extensive margin captures the extent to which individuals, who had not been using financial services before, gain access.
+On the other hand, the intensive margin describes the growing use of finance by the agents who had already been using it before \parencite{demirgucc2009finance}.
 Financial development on the extensive margin might lead to more equal opportunities and outcomes.
-Access to credit by previously disadvantaged groups allows human capital accumulation \parencite{galorzeira1993income, galormoav2004, braunetal2019}, formation and growth of new firms \parencite{evans1989estimated,banerjeenewman1990}, with more evenly distributed economic opportunities as a result\footnote{Having similar economic opportunities might decrease the cross-generational inequality, by diminishing the effect of e.g.
+Access to credit by previously disadvantaged groups allows human capital accumulation \parencite{galorzeira1993income, galormoav2004, braunetal2019} as well as formation and growth of new firms \parencite{evans1989estimated, banerjeenewman1990}, with more evenly distributed economic opportunities as a result\footnote{Having similar economic opportunities might decrease the cross-generational inequality, by diminishing the effect of, e.g.
 parental wealth.
@@
 
-On the contrary, intensive margin of financial development might disproportionately benefit the rich who may leverage financial services for their further benefit or to protect their existing rents.
+On the contrary, the intensive margin of financial development might disproportionately benefit the rich who may leverage financial services for their further benefit or to protect their existing rents.
 \textcite{GreenwoodJovanovic1990} present a model where the finance is the key driver of inequality and the welfare gains accrued by the incumbents - primarily the rich - in the initial development stage.
-With time, more agents meet the fixed costs of joining the financial intermediaries and they enjoy higher returns.
+With time, more agents meet the fixed costs of joining the financial intermediaries, and they enjoy higher returns.
 Consequently, the efficiency of resource allocation also increases, which enhances growth and reduces inequality.
 \textcite{perotti2007investor} present a framework based on political economy.
-They argument depends on a lobby for lower investor protection to prevent entrance of the new competitors.
+Their argument depends on a lobby for lower investor protection to prevent entrance of the new competitors.
 The politicians require higher bribe from the lobbyist the greater is their accountability for policy decisions.
 Thus, with increasing accountability, investor protection strengthens and spurs market entry and competition.
-The authors examine their prediction in a cross-section and show that better investor protection correlates with larger entry rates and higher firm density in more financially intensive sectors\footnote{In addition, they show that the most important factor of accountability is not the formal measure of democratic institutions, but newspaper readership which they interpret as broad awareness of policy choices and their outcomes.}.
+The authors examine their prediction in a cross-section and show that better investor protection correlates with larger entry rates and higher firm density in more financially intensive sectors\footnote{In addition, they demonstrate that the most important factor of accountability is not the formal measure of democratic institutions, but newspaper readership which they interpret as broad awareness of policy choices and their outcomes.}.
 
-Financial development may also have indirect effect on income inequality through economic growth.
+Financial development may also have an indirect effect on income inequality through economic growth.
 \textcite{townsendeueda2006} model how finance interacts with production and allocation of credit.
 If increased use of finance increases the demand for low- relatively to the high-skilled workers, then it may have equalizing consequences for income distribution.
-Empirical evidence by \textcite{beck2010big} show that bank deregulation and increased competition in loan provision in the \ac{US} primarily benefited the workers with income below the median.
+Empirical evidence by \textcite{beck2010big} shows that bank deregulation and increased competition in loan provision in the US primarily benefited the workers with income below the median.
 Similarly, \textcite{delis2014} provide evidence of bank deregulation and liberalization tightening the income distribution, although this effect is only present in countries with high-quality institutions.
@@
 
-A set of distinct papers explores the relationship between inequality and growth while stressing financial markets imperfections driving the outcomes.
+A set of distinct papers explores the relationship between inequality and growth while stressing financial market imperfections in driving the outcomes.
 Income inequality and growth may intersect through varying channels.
-Accumulation of savings, unobservable effort, and investment project size favor the prediction of growth inducing inequality.
-Negative impact of inequality on human capital accumulation, entrepreneurial activity provide argument for the opposing view.
-\textcite{milanovicvan2018inequality} report how income inequality in the \ac{US} has different implications for the future income growth of the rich and the poor.
+Accumulation of savings, unobservable effort, and investment project size favour the prediction of growth-inducing inequality.
+The negative impact of inequality on human capital accumulation and entrepreneurial activity provides an argument for the opposing view.
+\textcite{milanovicvan2018inequality} report how income inequality in the US has different implications for the future income growth of the rich and the poor.
 High inequality seems to hurt the prospects of the poor while the top of the distribution is unaffected.
-The rich thus disproportionately benefit from higher inequality as their subsequent income exhibit faster growth.
-The authors attribute this effect to the political channel the rich use to lobby in favor of the policies which support their economic interests.
+The rich thus disproportionately benefit from higher inequality as their subsequent income exhibits faster growth.
+The authors attribute this effect to the political channel where the rich use to lobby in favor of the policies which support their economic interests.
 Preferences of the rich are ultimately more likely to determine public policy than the preferences of the majority \parencite{gilens_page_2014}.
@@
 The papers link higher levels of financial development with lower levels of inequality \parencite{beck2007finance, hamori2012, gimet2011closer, kunieda2014finance}\footnote{For an extended list, we refer to \textcite{de2017finance}.}.
-On the other hand, several other estimate a inequality inducing effect of finance \parencite{Jaumotte2013, jauch2016financial, de2017finance}.
-Finally, some authors claim there the relationship might be non-linear, conditional on a threshold value of financial development \parencite{kim2011nonlinearity,tan2012nonlinear} or institutional quality \parencite{LawSingh2014, delis2014}.
+On the other hand, several other estimate an inequality inducing effect of finance \parencite{Jaumotte2013, jauch2016financial, de2017finance}.
+Finally, some authors claim that the relationship might be non-linear, conditional on a threshold value of financial development \parencite{kim2011nonlinearity,tan2012nonlinear} or institutional quality \parencite{LawSingh2014, delis2014}.
 
-Three papers are the closest to ours, each in a different respect.
+The contribution of the paper is fourfold.
+First, we efficiently account for model uncertainty relying on the panel \ac{BMA} framework.
+Second, we use the \ac{WID} data on the top income shares, collected based on the tax collection data.
+Arguably, the data is superior to the income surveys as it amends issues of underrepresentation of high-income individuals and underreporting income.
+Third, we simultaneously consider different proxies of financial development to identify the most relevant channels through which finance affects inequality and, fourth, we examine multiple measures of income inequality to distinguish the diverse effects of finance across the income distribution.
+
+Three papers are the closest to ours, each in a different respects.
 First, \textcite{de2017finance} examine different dimensions of finance on income inequality.
@@
 On the contrary, the financial development, proxied by the credit to GDP ratio, has inequality increasing effect irrespective of the institutional background.
-Second, \textcite{naceurzhang2016} take similar approach in considering multiple dimensions --- the access, efficiency, and stability of the financial sector, although not examining the indicators simultaneously.
+Second, \textcite{naceurzhang2016} take a similar approach in considering multiple dimensions --- the access, efficiency, and stability --- of the financial sector when studying determinants of inequality and poverty, but they do not attempt to consider the indicators simultaneously.
 Third, \textcite{furceri2019robust} apply \ac{WALS} to identify robust determinants of income inequality.
-Their approach mirrors ours in accounting for model uncertainty in the estimation.
-Their focus is more general rather than focused primarily on finance.
-We provide synthesis and extension to these papers in providing more detailed view on the link between finance in shaping income inequality and examining multiple measures of inequality while specifically identifying the determinants of top income shares along with the determinants of the overall income distribution\footnote{Captured by income Gini index.}.
+Their approach mirrors ours in accounting for model uncertainty in the estimation, but their focus is more general rather than focused primarily on finance.
+We provide a synthesis and a natural extension to these papers by offering more detail on how finance in shapes income inequality.
+On top of that, we examine multiple measures of inequality while specifically identifying the determinants of top income shares along with the determinants of the overall income distribution.
 
-% An exemption from focus solely on the size of financial sector is the study by \textcite{naceurzhang2016} which applies a similar approach to ours, taking into consideration the access, efficiency, and stability of the financial sector.
-However, there are severe limitation to their study.
-First, their coverage does not correspond to the availability of the data.
-Second, they do not account for the different dimensions of finance simultaneously and always use a single indicator of development at a time.
-Third, they do not explicitly differentiate between the banking sector and financial markets.
-We provide a more detailed as well as more robust picture of the financial development effect on finance in these dimensions.
+The chapter continues with a description of the data and methodology in Section \ref{sec:datameth}.
+We then provide two sets of results in Section \ref{sec:results}.
+We employ the Gini coefficient as a measure of inequality in one and income shares of top decile and percentile in the other.
+Section \ref{sec:robustness} provides robustness checks of the results and Section \ref{sec:conclusion} concludes.
 
-% \textcite{de2017finance} examine different dimensions of finance on income inequality.
-Their results suggest that financial development, financial liberalization, and banking crises all increase market income inequality within countries.
-Additionally, they show that the effect of financial liberalization is conditional on democratic accountability.
-The higher the accountability, the less severe is the negative impact of liberalization on inequality.
-On the contrary, the financial development, proxied by the credit to GDP ratio, has inequality increasing effect irrespective of the institutional background.
-
-% \textcite{furceri2019robust} apply WALS to identify robust determinants of income inequality.
-They are the closest paper to ours since they account for model uncertainty in the estimation.
-Their focus if more general rather than on finance specifically.
-Out work offers more detailed contribution in terms of the role of finance in shaping income inequality.
-On the top of that, we differ from their analysis by examining multiple measures of inequality and specifically identifying the determinants of top income shares along with the determinants of the overall income distribution represented by income Gini index.
-We provide further evidence in all these dimensions.
-
-%
-%
-%
 %
@@
 \section{Data and methodology}
+\label{sec:datameth}
 The key variable in the paper is the measure of income inequality.
-We want to examine how financial development affects income inequality and whether the effect might by different at the top quantiles of income distribution.
-As the overall measure of income inequality, we rely the after-tax Gini coefficient from \ac{SWIID} by \textcite{Solt2019}, which is a standard resource in the literature\footnote{There are alternative sources of for Gini coefficient, e.g.
+We want to examine how financial development affects income inequality and whether the effect might be different at the top quantiles of the income distribution.
+As the overall measure of income inequality, we rely upon the after-tax Gini coefficient from \ac{SWIID} by \textcite{Solt2019}, which is a standard resource in the literature\footnote{There are alternative sources of for Gini coefficient, e.g.
 \ac{WIID} or \ac{LIS}, but each of them brings limitations in terms of comparability or coverage.}.
@@
 It typically takes values in the interval between 0 and 100 where the former suggests perfect equality (everyone in the economy enjoys the same income) and the latter perfect inequality (all the income goes to only a single unit).
-We depart from existing papers slightly in considering the after-tax rather than the before-tax income distribution as a dependent variable.
+We depart from existing papers slightly in considering the after-tax rather than the before-tax income distribution as the dependent variable.
+We choose the after-tax income Gini coefficient as we also include the proxy of redistribution among the regressors to account for taxation and transfers indirectly.
+Since we define redistribution as a difference between before-tax and after-tax Gini coefficients, the estimate is not substantially influenced by using either of the two as the dependent variable\footnote{\textcite{furceri2019robust} make a similar argument that unless redistribution is systematically correlated with other regressors, their effect on the net and gross inequality should be same.
+As a robustness check, we also ran the estimation with before-tax Gini coefficient, and the single qualitative difference is only the sign of the posterior mean of redistribution coefficient.}.
+In our case, using after-tax allows for convenient interpretation.
 
-To explore the relationship in the top part of the distribution, we choose top income share from \ac{WID}\footnote{The methodology and guidelines to database are provided by \textcite{alvaredo2016distributional}.}.
-The surveys suffer from well-known issues of underrepresentation of the top income earners and the distortions resulting from self-reported character of the data.
-This can influence not only the top income shares resulting from survey data, but also distortions in the overall measures of inequality.
-The data in \ac{WID} make use of income tax records in individual countries and the derived shares obtained using consistent methodology of \ac{DINA} are arguably more reliable relative to the survey-based measures which are the primary source of majority estimates of income distributions.
+To explore the relationship in the top part of the distribution, we choose top income shares from the \ac{WID}\footnote{The methodology and guidelines to database are provided by \textcite{alvaredo2016distributional}.}.
+The surveys suffer from well-known issues of underrepresentation of the top income earners and the distortions resulting from the self-reported character of the data.
+This can influence not only the top income shares originating from the survey data, but it may also distort the overall measures of inequality.
+The data in \ac{WID} make use of income tax records in individual countries and the derived shares obtained using the consistent methodology of \ac{DINA} are arguably more reliable relative to the survey-based measures which are the primary source of majority estimates of income distributions.
 
 The data spans from 2000 to 2014.
-We follow the literature \parencite{dabla2015causes,de2017finance} and average both the inequality measure (dependent variable) and the potential determinants (independent variables) across three year intervals.
+We follow the literature \parencite{dabla2015causes,de2017finance} and average both the inequality measure (dependent variable) and the potential determinants (independent variables) across 3-year intervals.
 There are important reasons for looking at the averages than observation in individual years.
-Annual macroeconomic data are subject to fluctuations and the data on income inequality is noisy \textcite{delis2014}.
+Annual macroeconomic data are subject to fluctuations, and the data on income inequality is noisy \parencite{delis2014}.
 Averaging should diminish the level of noise.
 On the top of that, the variables at the center of our analysis, e.g.
-stock market capitalization or credit to \ac{GDP}, are likely to be affected by the business cycles and volatile on the yearly basis.
-Similar argument holds for top income shares, as they depend, among other things, on the bonuses paid out each year and capital income.
+stock market capitalization or credit to \ac{GDP}, are likely to be affected by the business cycles and volatile on a yearly basis.
+A similar argument holds for top income shares, as they depend, among other things, on the bonuses paid out each year as well as capital income.
 We want to explore the long-term rather than the short-term relationship and that guides the choice of averaged data.
-Faced against the trade-off between length of the averaging periods and available observations in the time dimension, we take a compromise of three years in contrast to the literature, where generally the 5-year intervals apply.
-The availability of financial development indicators limits the analysis to a period from 2000 onward and we prefer to keep at least 5 unique time periods to just three under the case of 5-year average\footnote{Nevertheless, we run the estimation with 5-year averages of data as a robustness check and find no critical qualitative differences compared to the baseline.}.
+Faced against the trade-off between the length of the averaging periods and available observations in the time dimension, we take a compromise of three years in contrast to the literature, where the 5-year intervals generally apply.
+The availability of financial development indicators limits the analysis to a period from 2000 onward, and we prefer to keep at least five unique time periods to just three under the case of 5-year average\footnote{Nevertheless, we run the estimation with 5-year averages of data as a robustness check and find no critical qualitative differences compared to the baseline.
+The results are available in the Appendix of the paper}.
+Two reasons further support this decision.
+First, the 3-year models display much better convergence which we grant to more available observations and higher variation in the 3-year averaged data.
+Second, more periods allow for a more robust estimate when we lag the explanatory variables in order to address endogeneity in one of the robustness checks.
 
-Table \ref{ch4tab:sum} report the summary statistics of the income inequality variables and financial development indicators.
-
-% \begin{table}[ht!]
-%     \small
-%     \caption{Summary statistics of inequality variables}
-%     \label{ch4tab:ineq}
-%     \centering
-%        \begin{tabular}{lrrrr}
-%         \toprule
-%         Variable & Mean & St. Dev. & Min & Max \\
-%         \midrule
-%         After-tax Gini index  & & & & \\
-%         Top 10\% income share & & & & \\
-%         Top 1\% income share & & & & \\
-%         \bottomrule
-%     \end{tabular}
-% \end{table}
+Table \ref{tab:sum} reports the summary statistics of the income inequality variables and financial development indicators.
 
@@
   \caption{Summary statistics of selected variables}
-  \label{ch4tab:sum}
+  \label{tab:sum}
   \centering
@@
   \caption{Correlation matrix of selected variables}
-  \label{ch4tab:corr}
+  \label{tab:corr}
   \centering
@@
   We obtain the financial development indicators from \ac{GFDD}.
-The database offers detailed indicators along four dimensions of financial systems and allows to estimate the affect of changes in access, size, efficiency, and stability of financial markets.
-Furthermore, we can distinguish between banking sector and financial markets in all these dimensions.
-The data in for access and stability of stock markets remains sparse in the concerned period we must leave them out of analysis.
-We use the version of financial indicators from \cite{svirydzenka2016introducing}.
-The authors make use of principal component analysis in order to construct aggregate indicators in each characteristic of financial sector.
-In summary, we have indicators of financial institutions depth (FID), financial markets depth (FMD), access to financial institutions (FIA), efficiency of financial markets (FME), and institutions (FIE)\footnote{\textcite{svirydzenka2016introducing} extrapolate the indicators from top to bottom if the original variables are unavailable, we make sure that that at least one variable is available for the construction of the index and no artificial correlation introduced to the data.}.
-We report the composition of each indicator in Table \ref{ch4tab:finind}.
+The database offers detailed indicators along four dimensions of financial systems and allows to estimate the effect of changes in access, size, efficiency, and stability of financial markets.
+Furthermore, we can distinguish between the banking sector and financial markets in all these dimensions.
+The data for access and stability of stock markets remains sparse in the concerned period we must leave them out of the analysis.
+We use the version of financial indicators from \textcite{svirydzenka2016introducing}.
+The authors make use of principal component analysis in order to construct aggregate indicators in each characteristic of the financial sector.
+In summary, we have indicators of financial institutions depth (FID), financial markets depth (FMD), access to financial institutions (FIA), the efficiency of financial markets (FME), and institutions (FIE)\footnote{\textcite{svirydzenka2016introducing} extrapolate the indicators from top to bottom if the original variables are unavailable, we make sure that that at least one variable is available for the construction of the index and no artificial correlation introduced to the data.}.
+We report the composition of each indicator in Table \ref{tab:finind}.
 
 \begin{figure}[ht!]
+  \caption{Gini Coefficient and top shares}
+  \label{fig:gini_topshares}
   \centering
-  \includegraphics[width=0.8\textwidth, keepaspectratio]{figures/plots_ineq}
-  \caption{Gini Coefficient and top shares}
-  \label{ch4fig:gini_topshares}
+  \includegraphics[width=0.8\textwidth, keepaspectratio]{figures//plots_ineq}
 \end{figure}
 
-% Figure \ref{ch4fig:gini_findev} shows the scatter plots of all the available observations after averaging between Gini coefficient and selected financial indicators.
+% Figure \ref{fig:gini_findev} shows the scatter plots of all the available observations after averaging between Gini coefficient and selected financial indicators.
 Although the data is noisy, it suggest negative correlation between the variables.
-Figure \ref{ch4fig:gini_findev_dm} then shows the corresponding scatter plots of the data which was demeaned using the within country averages over time.
+Figure \ref{fig:gini_findev_dm} then shows the corresponding scatter plots of the data which was demeaned using the within country averages over time.
 %The relationship between Gini and financial indicators appears considerably weaker following the transformation, with the exception of access to financial institutions where visual inspection allows pointing to a negative correlation.
 
-We build the choice of other explanatory variables on a reviews of income inequality drivers \parencite{roineetal2009,nolan2019drivers}, related study of finance-inequality nexus \parencite{de2017finance}, and a more general inquiry into the robust determinants of income inequality \parencite{furceri2019robust}.
-The potential regressors could be categorized in several groups.
+We base the choice of other explanatory variables on the reviews of income inequality drivers \parencite{roineetal2009,nolan2019drivers}, a related study of finance-inequality nexus \parencite{de2017finance}, and a more general inquiry into the robust determinants of income inequality \parencite{furceri2019robust}.
+The potential regressors could be categorized into several groups.
 They control for economic and financial development, demographics, globalization, and institutional background.
-Table \ref{ch4app:exog} reports all the control variables and their sources.
-
-% \textcite{nolan2019drivers} bring a survey of the literature on determinants of inequality, summarizing the complexity of the inequality dynamics.
-They stress that many of the determinants are interlinked which implies difficulty in assigning precise effects to individual drivers of inequality.
-Additionally, they encourage complementary individual country case studies to support the finding of the general cross-country estimates.
-% Both theoretical and empirical studies leave out the issue of importing the financial services from abroad. We include financial globalization from KOF among our control variables.
+Table \ref{app:exog} reports all the control variables and their sources.
 %
 %
-% \section{Methodology}
-% \label{ch4sec:methodology}
-Methodologically, we rely on the \ac{BMA} approach which conveniently addresses the issue of model uncertainty with a large number of potential regressors.
-The advantages are statistical properties of the \ac{BMA} have been sketched in \textcite{Koop2003}.
+% \section{Methodology}, could be extended if necessary
+% \label{sec:methodology}
+Methodologically, we rely on the \ac{BMA} approach, which conveniently addresses the issue of model uncertainty with a large number of potential regressors.
+The advantages and statistical properties of the \ac{BMA} have been described in \textcite{Koop2003}.
 In application to the panel data, we make use of Frisch-Waugh-Lovell theorem and demean all the variables using time averages for individual observations.
@@
 As for the model priors, the baseline estimate relies on the uniform model prior.
-We choose the model prior to remain agnostic about the prior probability of each examined model.
-While uniform prior assigns the same prior probability to each model, the distribution of the prior model space is concentrated around $k/2$, where $k$ is the number of potential covariates, and it may consequentially gravitate towards larger model sizes and higher number of covariates\footnote{See \textcite{LeySteel2009} for details.}.
+We choose the model before to remain agnostic about the prior probability of each examined model.
+While uniform prior assigns the same prior probability to each model, the distribution of the prior model space is concentrated around $k/2$, where $k$ is the number of potential covariates, and it may consequentially gravitate towards larger model sizes and a higher number of covariates\footnote{See \textcite{LeySteel2009} for details.}.
 
 \section{Results}
-\label{ch4sec:results}
+\label{sec:results}
 We examine the determinants of inequality in the panel \ac{BMA} framework and present the results in the following sections.
-We start with a model where we capture the overall inequality by Gini index in subsection \ref{ch4subsec:gini}.
+We start with a model where we capture the overall inequality by Gini index in Subsection \ref{subsec:gini}.
 We then continue to estimations where we consider the shares of income going to the top 10\% and top 1\% of the income distribution as our dependent variable.
@@
 
-\subsection{Gini index of income distribution}\label{ch4subsec:gini}
+\subsection{Gini index of income distribution}\label{subsec:gini}
 We focus the analysis on the relationship between the indicators representing various aspects of financial development and income inequality.
-Figure \ref{ch4fig:gini_findev_dm} outlines the expected link after we have demeaned the variables using the cross-sectional averages.
-The relationship is not particularly strong, but we observe negative correlation between Gini index and indicators of access and efficiency of financial institutions, as suggested by a linear estimate.
+Figure \ref{fig:gini_findev_dm} outlines the expected link after we have demeaned the variables using the cross-sectional averages.
+The relationship is not particularly strong, but we observe a negative correlation between Gini index and indicators of access and efficiency of financial institutions, as suggested by a linear estimate.
 For size indicators of financial market and institutions, the link appears much weaker.
@@
 \begin{figure}[ht!]
+  \caption{Gini Coefficient and Financial Development Indicators}
+  \label{fig:gini_findev_dm}
   \centering
-  \includegraphics[width=0.8\textwidth, keepaspectratio]{figures/plots_findev_gini_dm}
-  \caption{Gini Coefficient and Financial Development Indicators}
-  \label{ch4fig:gini_findev_dm}
+  \includegraphics[width=0.8\textwidth, keepaspectratio]{figures//plots_findev_gini_dm}
 \end{figure}
@@
 %   \caption{Gini Coefficient and Financial Development Indicators}
-%   \label{ch4fig:gini_findev}
-%   \includegraphics[width=\textwidth, keepaspectratio]{figures/plots_findev_gini}
+%   \label{fig:gini_findev}
+%   \includegraphics[width=\textwidth, keepaspectratio]{figures//plots_findev_gini}
 % \end{figure}
 
-Table \ref{ch4res:baseline_gini} reports the baseline results.
+Table \ref{res:baseline_gini} reports the baseline results.
 Overall, we have 16 variables with \ac{PIP} above 0.8.
@@
 The only financial indicators which occur among the top regressors are access and efficiency of financial institutions with \acp{PIP} of 1 and 0.88, respectively.
-The posterior mean on the coefficients in both dimension is negative so higher levels of access and efficiency are associated with lower levels of income inequality.
+The posterior mean on the coefficients in both dimensions is negative, so higher levels of access and efficiency are associated with lower levels of income inequality.
 The inequality decreasing effect of access to finance on inequality mirrors \textcite{hasan2020finance} for wealth inequality, and partially also \textcite{furceri2019robust, naceurzhang2016} who document similar effect.
-The observation on inequality decreasing effect of access to finance supports also \textcite{claessens2007finance} who suggest that access may equalize economic opportunities and lead to more evenly distributed income as well as theoretical predictions \parencite{braunetal2019,galormoav2004,banerjeenewman1990}.
-The efficiency of financial intermediation putting downward pressure on income inequality also has a precedent in \textcite{gimet2011closer}\footnote{The authors measure efficiency of banking sector by the difference between lending rate and the deposit rate (spread).
-They argue higher spread reflects low competition and high transaction costs.
+The observation on inequality decreasing effect of access to finance also supports \textcite{claessens2007finance} who suggest that access may equalize economic opportunities and lead to a more evenly distributed income as well as with theoretical predictions \parencite{braunetal2019,galormoav2004,banerjeenewman1990}.
+The efficiency of financial intermediation putting downward pressure on income inequality also has a precedent in \textcite{gimet2011closer}\footnote{The authors measure the efficiency of the banking sector by the difference between the lending rate and the deposit rate (spread).
+They argue that higher spread reflects low competition and high transaction costs.
 The imperfections in the credit market can skew the credit to high-income, rich households who can provide significant collateral, reinforcing the existing inequality.}.
-We fail to confirm that the efficiency of financial institutions is a robust determinant of inequality, however, as the \ac{PIP} markedly decreases under alternative model priors.
-None of the size indicators of financial institutions or markets has high probability of inclusion.
+We fail to confirm that the efficiency of financial institutions is a robust determinant of inequality; however, as the \ac{PIP} markedly decreases under alternative model priors.
+None of the size indicators of financial institutions or markets has a high probability of inclusion.
 
 Education expenditures (\% share of \ac{GDP}) along with the education index calculated using mean and expected years of schooling show inequality decreasing effect.
-This is in line with prediction of \textcite{goldin2009race,deaton2013great} who claim that the skill-biased technological change should be mitigated by education.
-\textcite{oecd2011divided} finds similar evidence for a panel of advanced countries while \textcite{furceri2019robust} also suggest a negative relationship, although it is not fully robust to variable and sample selection.
+It is in line with the prediction of \textcite{goldin2009race,deaton2013great} who claim that the skill-biased technological change should be mitigated by education.
+\textcite{oecd2011divided} finds similar evidence for a panel of advanced countries while \textcite{furceri2019robust} also suggest a negative relationship, although it is not entirely robust to variable and sample selection.
 The effect of education diminishes at the higher levels of schooling as the quadratic term has also high \ac{PIP} and positive posterior mean.
@@
 \begin{table}[ht!]
-  \caption{BMA, baseline results. Dependent variable after-tax Gini index, 394 observations.}\label{ch4res:baseline_gini}
+  \caption{BMA, baseline results. Dependent variable after-tax Gini index, 394 observations.}\label{res:baseline_gini}
   \footnotesize
@@
 
-We find evidence for the Kuznets' hypothesis about the inequality and economic development.
-We include three regressors which primarily capture the level of development - GDP per capita along with a square term, share of value added in agriculture, and share of value added in industry.
-Exploring the results of the baseline, We find evidence for an inverted-U curve for GDP per capita and baseline \acp{PIP} for linear and square component.
-This suggests that the inequality tends to be higher at the lower stages of economic development and only starts decreasing after reaching a threshold.
-The \ac{PIP} of 0.97 for value added in agriculture along with a negative posterior mean for its coefficient supports this idea further.
-The economies at lower stages of development generally report higher shares of agricultural sector which is also more labour intensive, i.e.
-not exacerbating the inequalities that stem from unequal distribution of capital.
+  We find evidence for the Kuznets' hypothesis about inequality and economic development.
+We include three regressors which primarily capture the level of development - GDP per capita along with its square term, the share of value-added in agriculture, and share of value-added in industry.
+Exploring the results of the baseline, We find evidence for an inverted-U curve for GDP per capita and baseline \acp{PIP} for the linear and square component.
+This suggests that inequality tends to be lower at the initial stages of economic development, then increases as the economic output grows, and only starts decreasing after reaching a threshold.
+The \ac{PIP} of 0.97 for value-added in agriculture along with a negative posterior mean for its coefficient supports this idea further.
+The economies at lower stages of development generally report higher shares of the agricultural sector which is also more labour intensive, i.e.
+not exacerbating the inequalities that stem from an unequal distribution of capital.
 The level of unemployment is associated with elevated income inequality.
@@
   We rely on redistribution to capture the effect of taxes and social expenditure on income inequality.
-Our measure is the the difference between before-tax and after-tax Gini coefficient from the \ac{SWIID} database.
-Therefore, we concentrate on the direct effects of policies on the household income.
-Given the global nature of our data and its limitations, we cannot estimate the potential indirect effects on the pre-tax distribution of income using, for example, top marginal tax rates \parencite{alvaredoetal2013}, corporate tax rates \parencite{fuest2018higher}.
-Nevertheless, the \ac{PIP} of our indicator is very high and it it negatively relates to inequality as expected.
-The government expenditures, an often used regressors in the literature also has a perfect inclusion probability, but with positive posterior mean.
+Our measure is the difference between before-tax and after-tax Gini coefficient from the \ac{SWIID} database.
+Therefore, we concentrate on the direct effects of policies on household income.
+Given the global nature of our data and its limitations, we cannot estimate the potential indirect effects on the pre-tax distribution of income using, for example, top marginal tax rates \parencite{alvaredoetal2013} nor corporate tax rates \parencite{fuest2018higher}.
+Nevertheless, the \ac{PIP} of our indicator is very high and it negatively relates to income inequality as expected.
+The government expenditures, an often used regressor in the literature, also has a perfect inclusion probability, but with positive posterior mean.
 The intuitive first-order effect of government expenditure should be through reduction of inequality through general social spending on transfers, education, and health.
-We account for these explicitly with redistribution and education expenditures of the government.
-\textcite{anderson2017does} introduce a meta-analysis where they show how the examined relationship depends on the type of spending considered and how it depends on the measure of income inequality.
-Additionally, they suggest that the redistributive impact has not extended over the entire distribution, but has mostly centered towards middle-income groups.
-In line with their conclusions, our results suggest the effect of government spending might run in the positive direction when the key equalizing function have been accounted for or perhaps point towards reverse causality.
+We account for these explicitly, however, with redistribution and education expenditures of the government.
+\textcite{anderson2017does} introduce a meta-analysis where they show how the examined relationship depends on the type of spending considered and the measure of income inequality.
+Additionally, they suggest that the redistributive impact has not extended over the entire distribution, but has mostly centred towards middle-income groups.
+In line with their conclusions, our results suggest the effect of government spending might run in a positive direction when the key equalizing function have been accounted for or perhaps point towards reverse causality.
 
@@
 We find a negative link between life expectancy and income inequality.
-\textcite{goldsteinlee2014} examine the channels between population ageing and inequality.
-Stretching the economic life-cycle are associated with larger within-group variance as a cohort ages.
-They argue this should pronounce inequality in older populations.
+\textcite{goldsteinlee2014} examine the channels between population ageing and inequality and argue that stretching the economic life-cycle is associated with larger within-group variance as cohort ages.
+They claim this should pronounce inequality in older populations.
 On the other hand, if we consider retirement age, pension structure in many countries equalize income flows and therefore the older populations may report lower levels of inequality.
@@
 The interest in the effects of monetary policy, and macroprudential tools in particular, on inequality is relatively recent, following the introduction of unconventional measures following the Great Recession \parencite{frost2018macroprudential}.
-Theoretically, the effect is ambiguous as high inflation of asset prices might benefit high-income households than the low income as they receive higher shares of financial income.
-Also, the assets held by low-income households tend to be much more liquid, thus inflation hurts them more relatively to the high-income ones.
+Theoretically, the effect is ambiguous as high inflation of asset prices might benefit high-income households more as they receive higher shares of financial income.
+Also, the assets held by low-income households tend to be much more liquid; thus inflation hurts them more relatively to the high-income ones.
 On the contrary, cut in rates usually benefits the borrowers and increased economic activity benefits the bottom-part of income distribution \parencite{furceri2019robust}.
-Our evidence supports the view of inflation enhancing income inequality, at least at moderate levels, as its square term is borderline on the inclusion, but negative in terms of posterior mean of its coefficient.
-This would suggest that the above mentioned theoretical effects interact and manifest differently at different inflation levels.
+Our evidence supports the view of inflation enhancing income inequality, at least up to a threshold, as its square term is borderline on the inclusion, but negative in terms of posterior mean of its coefficient.
+This would suggest that the above-mentioned theoretical effects interact and manifest with different strength at varying inflation levels.
 
-The index of the economic freedom of the world describes the overall business conditions by considering regulatory and legal environment.
+  The index of the economic freedom of the world describes the overall business conditions by considering the regulatory and legal environment.
 More economic freedom to trade and run daily business makes it easier to exploit economic opportunities and the high \ac{PIP} with positive posterior mean of the coefficient point to economic freedom making the distribution of income less equal.
-We consider many other potential explanatory variables, but in case of Gini index, they show low \acp{PIP}.
-We do not find any measure of globalization, institutions, or trade openness relevant for the overall distribution of income.
-In the next section, we concentrate on the top income share and whether the top part of the distribution is perhaps associated with different regressors.
+We consider many other potential explanatory variables, but in the case of the Gini index, they show low \acp{PIP}.
+We do not find any measure of globalization, institutions, or trade openness relevant to the overall distribution of income.
+In the next section, we concentrate on the top income shares and whether the top part of the distribution where we observe partially distinct results.
 
   \subsection{Top income shares}
-Financial development may influence various part of the income distribution differently.
-We therefore follow with baseline estimates with the top income shares as dependent variables.
-Table \ref{ch4res:baseline_top10} reports the results for top 10\% share and Table \ref{ch4res:baseline_top1} for the top 1\% share.
+  Financial development may influence various parts of the income distribution differently.
+We therefore follow by baseline estimates with the top income shares as dependent variables.
+Table \ref{res:baseline_top10} reports the results for top 10\% share and Table \ref{res:baseline_top1} for the top 1\% share.
 There is large overlap of the most important determinants for the top 10\% income share and Gini coefficient.
@@
 However, \acp{PIP} of some of the variables decline significantly and some other become relevant.
-Figure \ref{ch4fig:comp_ineq_sel} provides a visual comparison of the inclusion probabilities.
+Figure \ref{fig:comp_ineq_sel} provides a visual comparison of the inclusion probabilities.
 
 \begin{table}[ht!]
-  \caption{BMA, baseline results. Dependent variable Top 10\% share, 394 observations.}\label{ch4res:baseline_top10}
+  \caption{BMA, baseline results. Dependent variable Top 10\% share, 394 observations.}\label{res:baseline_top10}
     \footnotesize
@@
 In the case of the Top 10\% share, education expenditures, non-equipment investment, unemployment, redistribution, and an index of economic freedom drop out.
-With the exception of the economic freedom, there are good reasons to believe these factors mainly drive lower and middle-part of the income distribution, rather than the very top.
-While education expenditures may support public education system and allow for human capital accumulation in the low-income households, such effect might not be as relevant for the concentration at the top.
-We have discussed previously how redistribution policies mostly affect the middle quantiles of the distribution and high unemployment rates traditionally do not concern the well educated high-income households and individuals.
-Also share of value added in agriculture now has a low \ac{PIP}.
+With the exception of economic freedom, there are good reasons to believe these factors mainly drive lower and middle part of the income distribution, rather than the very top.
+While education expenditures may support the public education system and allow for human capital accumulation in low-income households, such an effect might not be as relevant for the concentration at the top.
+We have discussed previously how redistribution policies mostly affect the middle quantiles of the distribution and high unemployment rates traditionally do not concern the well educated high-income households.
+Also, the share of value-added in agriculture now has a low \ac{PIP}.
 The inclusion of non-equipment investment also decreases, while it retains the positive posterior mean.
 Most importantly, depth of financial markets and financial institutions now exhibit perfect \acp{PIP} and they seem to be associated with income distributions more concentrated at the top.
-The access to financial institutions remains relevant and negatively correlated with inequality.
-Among other variables with higher \acp{PIP}, we have natural resources rents, rule of law, equipment investment, and trade openness.
+Access to financial institutions remains relevant and negatively correlated with inequality.
+Among other variables with higher \acp{PIP}, we have natural resources rents, the rule of law, equipment investment, and trade openness.
 For the natural resources rents, we get a negative posterior mean.
-The evidence is in line with \textcite{goderismalone2011}, who describe a mechanism of income equalizing natural resource booms through the benefits for unskilled workers in labour intensive sector.
-A positive link between inequality and trade openness has been suggested by \textcite{Jaumotte2013,dabla2015causes} for the cross-sectional datasets and by \textcite{milanovicvan2018inequality} in the case of \ac{US}.
-The negative posterior coefficient on the rule of law is in line with the prediction by \textcite{perotti2007investor} and is sole variable indicating the importance of institutions posited in \textcite{acemoglu2003cross,acemoglu2015rise}.
+The evidence is in line with \textcite{goderismalone2011}, who describe a mechanism of income equalizing natural resource booms through the benefits for unskilled workers in labour-intensive sector.
+A positive link between inequality and trade openness has been suggested by \textcite{Jaumotte2013,dabla2015causes} for the cross-sectional datasets and by \textcite{milanovicvan2018inequality} in the case of the US.
+The negative posterior coefficient on the rule of law is in line with the prediction by \textcite{perotti2007investor} and is a sole variable indicating the importance of institutions posited in \textcite{acemoglu2003cross,acemoglu2015rise}.
 Equipment investment (machinery) may channel the resources to the bottom part of the income distribution through increase long-term growth rates and upward pressure on wages.
@@
 \begin{figure}[ht!]
+  \caption{\acp{PIP} with different inequality measures}
+  \label{fig:comp_ineq_sel}
   \centering
   \includegraphics[width=0.8\textwidth, keepaspectratio]{figures/comp_baseline_pip_sel}
-  \caption{\acp{PIP} with different inequality measures}
-  \label{ch4fig:comp_ineq_sel}
   \begin{minipage}{0.8\textwidth}
@@
 \begin{figure}[ht!]
+  \caption{\acp{PIP} for different top income shares with baseline priors}
+  \label{fig:comp_topshares}
   \centering
-  \includegraphics[width=\textwidth, keepaspectratio]{figures/topsh_comparison_all}
-  \caption{\acp{PIP} for different top income shares with baseline priors}
-  \label{ch4fig:comp_topshares}
+  \includegraphics[width=\textwidth, keepaspectratio]{figures//topsh_comparison_all}
 \end{figure}
@@
 \begin{table}[ht!]
-  \caption{BMA, baseline results. Dependent variable Top 1\% share, 394 observations.}\label{ch4res:baseline_top1}
+  \caption{BMA, baseline results. Dependent variable Top 1\% share, 394 observations.}\label{res:baseline_top1}
     \footnotesize
@@
 The top 1\% income earners might well be among the ones enjoying capital rents from the country resources.
-While education appears to have equalizing income effect for the lower part of the distribution, it is reasonable that it wears out for the very top income earners in the distribution.
-Figure \ref{ch4fig:comp_topshares} depicts comparison of inclusion probabilities of all variables considered in the estimation for the Top 10\% share and Top 1\% share.
+While education appears to have an equalizing income effect for the lower part of the distribution, it is reasonable that it wears out for the very top income earners in the distribution.
+Figure \ref{fig:comp_topshares} depicts comparison of inclusion probabilities of all variables considered in the estimation for the Top 10\% share and Top 1\% share.
 
-\subsection{Robustness checks}
+\section{Robustness checks}
+\label{sec:robustness}
 We perform robustness checks for all three specifications employing alternative model priors\footnote{We also perform robustness checks using alternative hyperparameter $g$ and Markov-chain samplers, but these do not affect our results.}.
@@
 \begin{figure}[ht!]
+  \caption{Robustness checks with alternative model priors, Gini coefficient}
+  \label{fig:gini_comp}
   \centering
   \includegraphics[width=\textwidth, keepaspectratio]{figures/model_priors_comparison_gini}
-  \caption{Robustness checks with alternative model priors, Gini coefficient}
-  \label{ch4fig:gini_comp}
 \end{figure}
@@
 \begin{figure}[ht!]
+  \caption{Robustness checks with alternative model priors, Top 10\% share}
+  \label{fig:top10_comp}
   \centering
   \includegraphics[width=\textwidth, keepaspectratio]{figures/model_priors_comparison_top10}
-  \caption{Robustness checks with alternative model priors, Top 10\% share}
-  \label{ch4fig:top10_comp}
 \end{figure}
@@
 \begin{figure}[ht!]
+  \caption{Robustness checks with alternative model priors, Top 1\% share}
+  \label{fig:top1_comp}
   \centering
   \includegraphics[width=\textwidth, keepaspectratio]{figures/model_priors_comparison_top1}
-  \caption{Robustness checks with alternative model priors, Top 1\% share}
-  \label{ch4fig:top1_comp}
 \end{figure}
@@
 The alternative priors influence the results similarly in all three specifications.
-The random mode prior decreases \acp{PIP} for the regressors as it generally prefers smaller models.
-Nevertheless, the results are only marginally effected for Gini coefficient as well as the top income shares with a few exception of education index in the case of Top 10\% share and trade openness for both top income shares.
-The use of dilution model prior which penalizes the models with highly correlated regressors, we see more significant differences.
+The random model prior decreases the \acp{PIP} for the regressors as it generally prefers smaller models.
+Nevertheless, the results are only marginally effected for Gini coefficient as well as the top income shares with a few exceptions of education index in the case of Top 10\% share and trade openness for both top income shares.
+When we apply a dilution prior which penalizes the models with highly correlated regressors, we see more significant differences.
 Above all, the quadratic terms show very low inclusion probabilities.
@@
 Taking the set of the important regressors in the baseline specification, we observe drop in the inclusion probability for education index, life expectancy, and economic freedom.
-Given high correlations among the set of regressors, this is not surprising and allows to narrow down the number of regressors robustly associated with income inequality.
+Given high correlations among the set of regressors, this is not surprising and allows us to narrow down further the number of regressors robustly associated with income inequality.
 
@@
 \section{Conclusion}
-\label{ch4sec:conclusion}
-In the paper, we explore the robust determinants of income inequality with a special attention given to indicators of financial development.
+\label{sec:conclusion}
+In the paper, we explore the robust determinants of income inequality with special attention given to indicators of financial development.
 We choose financial indicators that reflect the access, efficiency, and size of financial markets and institutions.
@@
 
-We show that financial development has complex relationship on income distribution.
-While access to finance has a universally inequality decreasing effect, the larger size of financial markets and financial institutions associates with higher top income shares, irrelevant for the Gini coefficient of the income distribution.
-This finding suggests that the size of financial markets also equalizes the income among the first 9 deciles.
+We show that financial development has a complex relationship with income distribution.
+While access to finance has a universal inequality decreasing effect, the larger size of financial markets and financial institutions associates with higher top income shares.
+The depth measures are, however, irrelevant for the Gini coefficient of the income distribution.
+This finding suggests that the size of financial markets may likely equalize the income among the first nine deciles.
 
@@
 Education, redistribution, and changing demographic structure seem to be linked with lower income inequality.
-On the contrary, unemployment, investment other than machinery, economic freedom - ease of pursuing economic opportunities, and inflation all positively relate to income inequality.
-As in the case of financial indicators, we find that the link could be more complex.
-When looking at the top income shares, some of the covariates (education, unemployment, redistribution, and economic freedom) cease to be relevant, while some other (trade openness, rule of law, and machinery investment) seem to matter.
+On the contrary, unemployment, investment other than machinery, economic freedom - ease of pursuing economic opportunities measured by an index of economic freedom, and inflation are all positively related to income inequality.
+As in the case of financial indicators, we find that the link could be more complicated.
+When looking at the top income shares, some of the covariates (education, unemployment, redistribution, and economic freedom) cease to be relevant, while some other (trade openness, the rule of law, and machinery investment) seem to matter.
 
@@
 While finance, technology, and trade likely affect the distributional outcomes, it can have varying influence on different parts of the income distribution.
-
-% Finance captures the capacity of financial intermediaries and markets to screen investment opportunities, monitor the debtors who were provided funding, as well as pooling and management of risk.
-With inequality, the focus of the paper is on the inequality in the distribution of income.
-Arguably, the literature discusses other concepts of inequality, e.g.
-intergenerational persistence of relative income differences or equality of opportunity \parencite{demirgucc2009finance}.
-
-% \textcite{kroszneretal2007} show that financial crises have relative more severe impact on the sectors which depend more on external financing.
-The consequences of crises on firms relate to institutional environment and materialize through lower production capacity and competition.
-
-% Overall, these results suggest that policies seeking to improve the governance and robustness of local banks should be prioritized over size-enhancing reforms from a normative point of view \parencite{gimet2011closer}.
-
-% The benefits of capital markets liberalization seem to be concentrated to the top of the income distribution.
-Top quintile of the distribution accrues nearly all of the income growth following the liberalization while the share of middle three quantiles decreases and the bottom remains unaffected \parencite{das2003income}.
 
@@
 \section{Appendix}
-\label{ch4sec:appch2}
+\label{appch2}
+
+\subsection*{Robustness checks}
+\begin{table}[ht!]
+  \centering
+  \small
+  \caption{BMA, results using 5-year averages. Dependent variable after-tax Gini index, 237 observations.}\label{res:5y_gini}
+  \begin{tabular}{lrrr}
+    \toprule
+   & PIP & Post Mean & Post SD \\
+    \midrule
+    Education expenditures & 1.00 & -0.20717 & 0.06349 \\
+    Value added in agriculture & 1.00 & -0.21937 & 0.07488 \\ 
+    GDP per capita & 1.00 & 1.82050 & 0.67271 \\ 
+    Life expectancy & 1.00 & -0.40392 & 0.12755 \\ 
+    Unemployment & 1.00 & 0.20341 & 0.06811 \\
+    Non-equipment investment & 1.00 & 0.20972 & 0.06696 \\ 
+    Access to financial institutions & 1.00 & -0.24817 & 0.08502 \\ 
+    Redistribution & 1.00 & -0.24501 & 0.06560 \\ 
+    Equipment investment & 1.00 & -0.17577 & 0.07223 \\ 
+    Education index (UN) & 1.00 & 0.23121 & 0.22046 \\ 
+    GDP per capita sq. & 1.00 & -1.62965 & 0.69937 \\
+    Value added in industry & 0.98 & -0.16154 & 0.07777 \\ 
+    Total population & 0.91 & -0.23792 & 0.13311 \\ 
+    Period 2 & 0.85 & 0.22049 & 0.14295 \\ 
+    Period 3 & 0.83 & 0.33874 & 0.23232 \\
+    Financial institutions efficiency & 0.72 & -0.07259 & 0.07062 \\ 
+    Financial markets depth & 0.53 & 0.04762 & 0.06528 \\
+    Government expenditures & 0.45 & 0.03445 & 0.05634 \\ 
+    Trade openness & 0.32 & 0.02846 & 0.05972 \\
+    Net FDI (\% GDP) & 0.30 & -0.01797 & 0.04012 \\ 
+    Education index sq. & 0.23 & 0.07123 & 0.20006 \\ 
+    Median age & 0.22 & -0.03112 & 0.09112 \\ 
+    Economic freedom & 0.18 & 0.01209 & 0.04296 \\ 
+    Financial institutions depth & 0.14 & 0.00671 & 0.03208 \\ 
+    Gross domestic savings & 0.11 & 0.00457 & 0.02938 \\ 
+    Political globalization & 0.10 & 0.00380 & 0.02894 \\
+    Natural resources rents & 0.10 & -0.00290 & 0.02291 \\ 
+    Restrictions on globalization & 0.09 & 0.00222 & 0.02066 \\ 
+    Financial globalization & 0.09 & 0.00226 & 0.02209 \\
+    Rule of law & 0.08 & -0.00214 & 0.01901 \\ 
+    Chinn-Ito index & 0.08 & 0.00164 & 0.01868 \\
+    Inflation & 0.08 & -0.00003 & 0.02409 \\ 
+    Social globalization & 0.08 & 0.00238 & 0.03812 \\
+    Population growth & 0.08 & 0.00144 & 0.01775 \\
+    GDP growth & 0.08 & 0.00144 & 0.02068 \\
+    Left-wing orientation & 0.08 & 0.00099 & 0.01578 \\ 
+    Civil liberties \& political rights & 0.07 & 0.00046 & 0.01448 \\
+    Inflation sq. & 0.00 & 0.00044 & 0.01696 \\ 
+    \bottomrule
+  \end{tabular}
+  \end{table}
+
+\begin{table}[ht!]
+    \centering
+    \small
+    \caption{BMA, results using 5-year averages. Dependent variable Top 10\% share, 248 observations.}\label{res:5y_top10}
+    \begin{tabular}{lrrr}
+      \toprule
+     & PIP & Post Mean & Post SD \\
+      \midrule
+      GDP per capita & 1.00 & 2.43240 & 0.72823 \\ 
+      Life expectancy & 1.00 & -0.71435 & 0.12493 \\
+      Financial institutions depth & 1.00 & 0.22910 & 0.06928 \\
+      Financial markets depth & 1.00 & 0.20245 & 0.05987 \\ 
+      Natural resources rents & 1.00 & -0.16995 & 0.06525 \\ 
+      GDP per capita sq. & 1.00 & -2.00889 & 0.74540 \\ 
+      Equipment investment & 1.00 & -0.19131 & 0.07268 \\ 
+      Government expenditures & 1.00 & 0.16863 & 0.06649 \\
+      Trade openness & 1.00 & 0.17805 & 0.07117 \\ 
+      Access to financial institutions & 0.99 & -0.21177 & 0.08756 \\ 
+      Left-wing orientation & 0.93 & 0.10136 & 0.05834 \\ 
+      Total population & 0.91 & 0.19759 & 0.11773 \\ 
+      Rule of law & 0.85 & -0.09518 & 0.06703 \\ 
+      Education index (UN) & 0.80 & -0.46592 & 0.34614 \\
+      Education index sq. & 0.79 & 0.49719 & 0.34504 \\ 
+      Gross domestic savings & 0.66 & 0.08137 & 0.08271 \\ 
+      Value added in agriculture & 0.56 & -0.07408 & 0.08899 \\ 
+      Financial institutions efficiency & 0.52 & -0.04852 & 0.06384 \\
+      Value added in industry & 0.43 & -0.05450 & 0.08111 \\ 
+      Political globalization & 0.21 & 0.01674 & 0.04917 \\ 
+      Unemployment & 0.17 & -0.01028 & 0.03389 \\ 
+      Inflation & 0.14 & 0.02605 & 0.11365 \\
+      Financial globalization & 0.13 & -0.00795 & 0.03274 \\ 
+      Population growth & 0.13 & -0.00590 & 0.02584 \\
+      Period 3 & 0.12 & 0.00818 & 0.03851 \\ 
+      Net FDI (\% GDP) & 0.12 & -0.00492 & 0.02292 \\
+      Social globalization & 0.10 & -0.00759 & 0.04529 \\ 
+      Non-equipment investment & 0.10 & 0.00374 & 0.02229 \\ 
+      Period 2 & 0.09 & -0.00305 & 0.02088 \\
+      Chinn-Ito index & 0.08 & 0.00262 & 0.01973 \\
+      Economic freedom & 0.08 & -0.00195 & 0.02455 \\ 
+      Education expenditures & 0.08 & -0.00162 & 0.01938 \\
+      Inflation sq. & 0.07 & -0.02586 & 0.11169 \\
+      Civil liberties \& political rights & 0.07 & 0.00149 & 0.01628 \\
+      GDP growth & 0.07 & 0.00119 & 0.01756 \\
+      Restrictions on globalization & 0.06 & 0.00071 & 0.01643 \\ 
+      Redistribution & 0.06 & -0.00049 & 0.01509 \\
+      \bottomrule
+    \end{tabular}
+    \end{table}
+
+    \begin{table}[ht!]
+      \centering
+      \small
+      \caption{BMA, results using 5-year averages. Dependent variable Top 1\% share, 248 observations.}\label{res:5y_top1}
+      \begin{tabular}{lrrr}
+        \toprule
+       & PIP & Post Mean & Post SD \\
+        \midrule
+        GDP per capita & 1.00 & 0.62260 & 0.35454 \\
+  Life expectancy & 1.00 & -0.57281 & 0.08981 \\ 
+  Equipment investment & 1.00 & -0.24601 & 0.07297 \\ 
+  Access to financial institutions & 1.00 & -0.29347 & 0.08101 \\ 
+  Financial institutions depth & 1.00 & 0.18949 & 0.07033 \\
+  Financial markets depth & 1.00 & 0.18562 & 0.06481 \\ 
+  Rule of law & 0.98 & -0.14120 & 0.06108 \\ 
+  Government expenditures & 0.82 & 0.08860 & 0.06512 \\ 
+  Trade openness & 0.77 & 0.09529 & 0.07711 \\ 
+  Value added in agriculture & 0.54 & -0.05397 & 0.06806 \\
+  Net FDI (\% GDP) & 0.34 & -0.02701 & 0.04919 \\ 
+  GDP growth & 0.33 & -0.02864 & 0.05416 \\ 
+  Inflation & 0.27 & 0.12719 & 0.26014 \\ 
+  Financial globalization & 0.24 & -0.02019 & 0.04855 \\
+  Inflation sq. & 0.24 & -0.12342 & 0.25456 \\
+  Chinn-Ito index & 0.23 & 0.02064 & 0.04977 \\
+  GDP per capita sq. & 0.19 & -0.12032 & 0.35046 \\ 
+  Political globalization & 0.14 & 0.01158 & 0.04063 \\
+  Period 2 & 0.13 & -0.00828 & 0.03020 \\
+  Economic freedom & 0.12 & -0.01092 & 0.04185 \\ 
+  Financial institutions efficiency & 0.11 & -0.00656 & 0.02705 \\
+  Period 3 & 0.10 & 0.00749 & 0.03410 \\
+  Redistribution & 0.10 & 0.00532 & 0.02408 \\
+  Social globalization & 0.09 & -0.00936 & 0.04582 \\
+  Total population & 0.08 & 0.00639 & 0.03494 \\ 
+  Non-equipment investment & 0.07 & 0.00317 & 0.02029 \\ 
+  Left-wing orientation & 0.06 & 0.00221 & 0.01583 \\ 
+  Restrictions on globalization & 0.06 & 0.00240 & 0.01962 \\
+  Value added in industry & 0.05 & -0.00204 & 0.01842 \\ 
+  Gross domestic savings & 0.05 & -0.00117 & 0.01651 \\
+  Unemployment & 0.04 & -0.00091 & 0.01415 \\ 
+  Education index (UN) & 0.04 & -0.00015 & 0.03208 \\
+  Civil liberties \& political rights & 0.04 & -0.00073 & 0.01128 \\ 
+  Natural resources rents & 0.03 & 0.00008 & 0.01104 \\ 
+  Population growth & 0.03 & 0.00047 & 0.00973 \\
+  Education expenditures & 0.03 & -0.00006 & 0.01103 \\ 
+  Education index sq. & 0.00 & 0.00141 & 0.02697 \\
+       \bottomrule
+   \end{tabular}
+\end{table}
+\clearpage
 
 \subsection*{The composition of financial indicators}
-\label{ch4subsec:finind_comp}
+\label{subsec:finind_comp}
 \begin{table}[ht!]
@@
     \caption{Underlying Components of Financial Development Indicators}
-    \label{ch4tab:finind}
+    \label{tab:finind}
     \centering
@@
     \caption{List of variables}
-    \label{ch4app:exog} \\
+    \label{app:exog} \\
     \toprule 
@@
   CLandPR & Average of index for civil liberties and political rights & \href{https://freedomhouse.org/report/freedom-world-2016/methodology}{Freedom House} \\
-%   OutwardO & Trade (\% of GDP)  & \href{http://data.worldbank.org/indicator/NE.TRD.GNFS.ZS}{WB} \\
+  OutwardO & Trade (\% of GDP)  & \href{http://data.worldbank.org/indicator/NE.TRD.GNFS.ZS}{WB} \\
   ChinnIto & Chinn-Ito index of financial openness. & \href{http://web.pdx.edu/~ito/Chinn-Ito_website.htm}{Chinn-Ito} \\ 
@@
   Loan2Deposits & Loan-to-deposit ratio. & \href{http://data.worldbank.org/data-catalog/global-financial-development}{GFDD} \\
-  Redist & Difference between market (pre-tax) and net (after-tax0 Gini index based on distribution of income  (The Standardized World Income Inequality Database).
+  Redist & Difference between market (pre-tax) and net (after-tax) Gini index based on distribution of income  (The Standardized World Income Inequality Database).
 & \href{http://fsolt.org/swiid/}{Solt (2016)} \\

### 2020-12-16 Scripts/191219 analysis alternative with pre-tax Gini index.R

@@ -39,10 +39,10 @@ data_f[year %in% c(2009:2011), period := 4]
 data_f[year %in% c(2012:2014), period := 5]
 
 # average observations by period and country
-data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
+data_f <- data_f[, lapply(.SD, function(x) {mean(as.numeric(x), na.rm = T)}), by = c("iso3c","country","period")]
 data_f[, c('year') := NULL]
 
-data_f <- data_f[!is.nan(GiniNet),]
+data_f <- data_f[!is.nan(GiniMarket),]
 
 data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
 		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
@@ -83,12 +83,12 @@ data_dm[, c("period","period_1","WarYears","RevCoups","Top1share","Top10share") 
 dummies <- colnames(dum)[2:(ncol(dum))]
 
 # run BMA
-bma_3y_gini_baseline <- bms(data_dm, iter=5000000, burn=1000000, mprior = "uniform", g = "hyper",
+bma_3y_pretaxgini_baseline <- bms(data_dm, iter=5000000, burn=1000000, mprior = "uniform", g = "hyper",
                     nmodel=5000, mcmc="bd.int",
                     fixed.reg = dummies, user.int = F)
 
-coef(bma_3y_gini_baseline, exact = T, std.coefs = T)
-summary(bma_3y_gini_baseline)
+coef(bma_3y_pretaxgini_baseline, exact = T, std.coefs = T)
+summary(bma_3y_pretaxgini_baseline)
 
 # save the results into file
-save(bma_3y_gini_baseline, file = here("Results/bma_3y_ginimarket_baseline.Rdata"))
+save(bma_3y_pretaxgini_baseline, file = here("Results/bma_3y_ginimarket_baseline.Rdata"))

### 2020-12-16 Scripts/191219 analysis baseline 3-year.R

@@ -11,6 +11,21 @@ library(imputeTS)
 library(here)
 library(dummies)
 library(xtable)
+library(rlang)
+library(dplyr)
+
+# SD function
+
+xtsum <- function(data, varname, unit) {
+  varname <- enquo(varname)
+  loc.unit <- enquo(unit)
+ores <- data %>% summarise(ovr.mean=mean(!! varname, na.rm=TRUE), ovr.sd=sd(!! varname, na.rm=TRUE), ovr.min = min(!! varname, na.rm=TRUE), ovr.max=max(!! varname, na.rm=TRUE), ovr.N=sum(as.numeric((!is.na(!! varname)))))
+bmeans <- data %>% group_by(!! loc.unit) %>% summarise(meanx=mean(!! varname, na.rm=T), t.count=sum(as.numeric(!is.na(!! varname))))
+bres <- bmeans %>% ungroup() %>% summarise(between.sd = sd(meanx, na.rm=TRUE), between.min = min(meanx, na.rm=TRUE), between.max=max(meanx, na.rm=TRUE), Units=sum(as.numeric(!is.na(t.count))), t.bar=mean(t.count, na.rm=TRUE))
+wdat <- data %>% group_by(!! loc.unit) %>% mutate(W.x = scale(!! varname, scale=FALSE))
+wres <- wdat %>% ungroup() %>% summarise(within.sd=sd(W.x, na.rm=TRUE), within.min=min(W.x, na.rm=TRUE), within.max=max(W.x, na.rm=TRUE))
+return(list(ores=ores,bres=bres,wres=wres))
+}
 
 # read data file
 data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
@@ -41,7 +56,7 @@ data_f[year %in% c(2009:2011), period := 4]
 data_f[year %in% c(2012:2014), period := 5]
 
 # average observations by period and country
-data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
+data_f <- data_f[, lapply(.SD, function(x){mean(as.numeric(x), na.rm = T)}), by = c("iso3c","country","period")]
 data_f[, c('year') := NULL]
 
 data_f <- data_f[!is.nan(GiniNet),]
[…]

### 2020-12-16 Scripts/191221 Figures robustness.R

@@ -37,6 +37,18 @@ legend("topright", legend=c("Uniform", "Dilution","Random"), pch=c(1:3), bty="n"
        title = expression(bold("Model priors")), title.adj = 0.1, col=c(rgb(0.03, 0.15, 0.4),rgb(0.753,0,0),rgb(0, 0, 0)))
 dev.off()
 
+# write in file, presentation
+cairo_ps(file = here("Paper/figures/model_priors_comparison_gini_pres.eps"), width=9, height=6, family="Arial")
+par(mar=c(6,4,0,0), bg='#f9f9f9')
+plotComp(bma_3y_gini_baseline, bma_3y_gini_dilut, bma_3y_gini_random,
+         lwd=1, varNr = NULL, comp = "PIP", exact = T, include.legend = F,
+         add.grid = F, do.par = T, cex.axis=1, cex.xaxis=1,
+         col=c(rgb(0.03, 0.15, 0.4),rgb(0.753,0,0),rgb(0, 0, 0)))
+
+legend("topright", legend=c("Uniform", "Dilution","Random"), pch=c(1:3), bty="n", inset=c(0.02,0.02),
+       title = expression(bold("Model priors")), title.adj = 0.1, col=c(rgb(0.03, 0.15, 0.4),rgb(0.753,0,0),rgb(0, 0, 0)))
+dev.off()
+
 # plot comparison - Top 10 share
 load(here("Results/bma_3y_top10_dilut.Rdata"))
 load(here("Results/bma_3y_top10_baseline.Rdata"))
@@ -60,6 +72,21 @@ legend("topright", legend=c("Uniform", "Dilution","Random"), pch=c(1:3), bty="n"
 
 dev.off()
 
+# write in file, presentation
+cairo_ps(file = here("Paper/figures/model_priors_comparison_top10_pres.eps"), width=9, height=6, family="Arial")
+par(mar=c(6,4,0,0), bg='#f9f9f9')
+plotComp(bma_3y_top10_baseline, bma_3y_top10_dilut, bma_3y_top10_random,
+         lwd=1, varNr = NULL, comp = "PIP", exact = T, include.legend = F,
+         add.grid = F, do.par = T, cex.axis=1, cex.xaxis=1,
+         col=c(rgb(0.03, 0.15, 0.4),rgb(0.753,0,0),rgb(0, 0, 0)))
+
[…]

### 2020-12-16 Scripts/200103 Theoretical Gini indices.r

@@ -4,6 +4,8 @@
 # Libraries
 library(data.table)
 library(ineq)
+library(ggplot2)
+library(here)
 
 # define the inputs
 p = seq(0,1,0.01)
@@ -17,9 +19,30 @@ gini = 0.56
 s = sqrt(2)*qnorm((gini+1)/2) # for log-normal distribution
 alpha = 1/(2*gini)+1/2 # for pareto distribution
 
-# plot
-plot(p,Lc.lognorm(p, parameter=s), type="l", col="brown")
+# put data into data table
+data <- data.table(p,gini, s=sqrt(2)*qnorm((gini+1)/2),alpha = 1/(2*gini)+1/2,
+                   lognorm = Lc.lognorm(p, parameter=s), pareto = Lc.pareto(p, parameter=alpha))
+
+# ggplot
+pl_1 <- ggplot(data, aes(x=p,y=lognorm, colour='brown'))
+pl_1 + geom_line() + geom_line(aes(x=p, y=pareto, color='blue')) + theme_bw() +
+       scale_colour_manual(name='', values=c('brown','blue')) + xlab('Cumulative share of population') + ylab('Cumulative share of income') +
+       theme(legend.position = 'none',
+             legend.background=element_rect(color="#f9f9f9",fill="#f9f9f9"),
+             plot.background=element_rect(color="#f9f9f9",fill="#f9f9f9"))
+
+# write into file
+cairo_ps(file = here("Paper/figures/theoretical_ginis_pres.eps"), width=9, height=6, family="Arial")
+pl_1 + geom_line() + geom_line(aes(x=p, y=pareto, color='blue')) + theme_bw() +
+       scale_colour_manual(name='', values=c('brown','blue')) + xlab('Cumulative share of population') + ylab('Cumulative share of income') +
+       theme(legend.position = 'none',
+             legend.background=element_rect(color="#f9f9f9",fill="#f9f9f9"),
+             plot.background=element_rect(color="#f9f9f9",fill="#f9f9f9"))
+dev.off()
+
+# plot, base
+plot(p, Lc.lognorm(p, parameter=s), type="l", col="brown")
 lines(p,Lc.pareto(p, parameter=alpha), col="blue")
 
 # compute the gini with specific alpha coefficient in Pareto distribution
-gini = 1/((2*aplha+0.45)-1)
+# gini = 1/((2*alpha+0.45)-1)

### 2020-12-16 Scripts/200103 analysis baseline 3-year, between estimation.R

@@ -63,8 +63,31 @@ setcolorder(data_try, c("GiniNet", names(data_try[,names(data_try)!="GiniNet"]))
 # drop redundant columns, consistency with the panel estimation
 data_try[, c("WarYears","RevCoups","Top1share","Top10share") := NULL]
 
+#
+#
+
+# Initial analysis
+
+#
+#
+
+# Within and between standard deviation
+sd(data_try$FIA)
+sd(data_try$FIE)
+sd(data_try$FID)
+sd(data_try$FMD)
+
+#
+#
+
+# BMA estimate
+
+#
+#
+
+
 # run BMA
-bma_gini_BE <- bms(data_try, iter=500000, burn=100000, mprior = "uniform", g = "hyper",
+bma_gini_BE <- bms(data_try, iter=3000000, burn=1000000, mprior = "uniform", g = "hyper",
                     nmodel=5000, mcmc="bd.int", user.int = F)
 
 coef(bma_gini_BE, exact = T, std.coefs = T)
@@ -77,7 +100,7 @@ save(bma_gini_BE, file = here("Results/bma_gini_BE.Rdata"))
 #
 
 # dilution model prior (accounting for correlation between regressors)
-bma_gini_BE_dilut <- dilutBMS2::bms(data_try, iter=5000000, burn=1000000, mprior = "dilut",
+bma_gini_BE_dilut <- dilutBMS2::bms(data_try, iter=3000000, burn=1000000, mprior = "dilut",
 						  g = "hyper", nmodel=5000, mcmc="bd.int", user.int = F)
 
 coef(bma_gini_BE_dilut, exact = T, std.coefs = T)

### 2020-12-16 Scripts/200326 analysis 5-year averages.r

@@ -10,8 +10,7 @@ library(here)
 library(dummies)
 
 # # Set the working directory
-# wd <- c("c:/Users/JM/Documents/GitHub/finineq/")
-# setwd(wd)
+library(here)
 
 # read data file
 data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
@@ -41,7 +40,7 @@ data_f[year %in% c(2005:2009), period := 2]
 data_f[year %in% c(2010:2014), period := 3] 
 
 # average observations by period and country
-data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
+data_f <- data_f[, lapply(.SD, function(x) {mean(as.numeric(x), na.rm = T)}), by = c("iso3c","country","period")]
 data_f[, c('year') := NULL]
 
 data_f <- data_f[!is.nan(GiniNet),]
@@ -93,10 +92,11 @@ bma_5y_gini_baseline <- bms(data_dm, iter=5000000, burn=1000000, mprior = "unifo
                     fixed.reg = dummies, user.int = F)
 
 coef(bma_5y_gini_baseline, exact = T, std.coefs = T)
-summary(bma_5y_gini_baseline)
+summary(bma_5ygini_baseline)
 
 # save the results into file
 save(bma_5y_gini_baseline, file = here("Results/bma_5y_gini_baseline.Rdata"))
+load('Results/bma_5y_gini_baseline.Rdata')
 
 #
 #
@@ -142,7 +142,7 @@ data_f[year %in% c(2005:2009), period := 2]
 data_f[year %in% c(2010:2014), period := 3] 
 
 # average observations by period and country
-data_f <- data_f[, lapply(.SD, mean(as.numeric(x)), na.rm = T), by = c("iso3c","country","period")]
+data_f <- data_f[, lapply(.SD, function(x) {mean(as.numeric(x), na.rm = T)}), by = c("iso3c","country","period")]
 data_f[, c('year') := NULL]
 
 data_f <- data_f[!is.nan(Top10share),]
@@ -236,7 +236,7 @@ data_f[year %in% c(2005:2009), period := 2]
 data_f[year %in% c(2010:2014), period := 3] 
 
 # average observations by period and country
[…]

### 2020-12-16 Scripts/200326 splitting the sample to pre- and after-crisis 2007.r

@@ -8,10 +8,10 @@
 # Libraries
 library(data.table)
 library(countrycode)
-library(dilutBMS2)
+# library(dilutBMS2)
 library(imputeTS)
 library(dummies)
-
+library(BMS)
 # Set the working directory
 library(here)
 
@@ -31,7 +31,60 @@ data[, c("GDPpc#GDPpc","Infl#Infl","EducIndex#EducIndex") := .(GDPpc^2, Infl^2, 
 #
 #
 
-# analysis - splitting into 2 cross-sections
+# analysis - full sample averaged
+
+#
+#
+
+# Estimate using full sample
+data_full <- data
+
+# Average observations by country
+data_f <- data_full[, lapply(.SD, function(x) {mean(as.numeric(x), na.rm = T)}), by = c("iso3c","country")]
+data_f[, c('year') := NULL] # drop year column
+
+data_f <- data_f[!is.nan(GiniNet),] # any NA in Gini?
+
+data_try <- data_f[, c("NetFDIout","TaxR","NNSavings","EmplRate","PubEducExp","EnrolPri", # unavailable for a lot of years
+		   "EnrolSec","EnrolTer","wage.mintomean.OECD","wage.mintomedian.OECD",  # unavailable for a lot of years
+		   "rminwage.OECD","meanwage.OECD","minwage.ILO","meanwage.ILO", "ud", # unavailable for a lot of years
+		   "FMA","FME", # unavailable for a lot of years
+		   "BankAcc","BankBranches","StockOutTop10","PrivateCredit","MarketCap", # original financial indicators
+		   "MarketTurn","ZScore","StockVol","NIM", # original financial indicators
+		   "ExpMetalOre","ExpFuel", # correlated with NatRes
+		   "corr","burreau","demAcc","lawOrder", # ICRG data
+		   "GFCF", # duplicit with NonequipI + EquipI ) := NULL]
+		   # "WarYears","RevCoups", # invariant in the sample
+		   "GiniMarket","RedistRel", # alternative inequality indicators
+		   "GovSize","LegalSystem","SoundMoney","TradeFreedom","Regulation", # subcomponents of EFW
+		   "PrEdu","SecEdu","TerEdu") := NULL]
+
+data_try <- data_try[complete.cases(data_try), ]
+
+# drop redundant columns
+data_full <- data_try[, c("iso3c","country","WarYears","RevCoups","Top1share","Top10share") := NULL]
+
[…]

### 2020-12-16 Scripts/200701 analysis 3-year lagged explanatory variables.R

@@ -1,0 +1,435 @@
+# Financial Development and Income Inequality
+# Baseline analysis, 3-y averages
+# Jan Mares, 1912019
+
+# Libraries
+library(data.table)
+library(countrycode)
+library(BMS)
+library(dilutBMS2)
+library(imputeTS)
+library(here)
+library(dummies)
+library(xtable)
+library(rlang)
+library(dplyr)
+
+# SD function
+
+xtsum <- function(data, varname, unit) {
+  varname <- enquo(varname)
+  loc.unit <- enquo(unit)
+ores <- data %>% summarise(ovr.mean=mean(!! varname, na.rm=TRUE), ovr.sd=sd(!! varname, na.rm=TRUE), ovr.min = min(!! varname, na.rm=TRUE), ovr.max=max(!! varname, na.rm=TRUE), ovr.N=sum(as.numeric((!is.na(!! varname)))))
+bmeans <- data %>% group_by(!! loc.unit) %>% summarise(meanx=mean(!! varname, na.rm=T), t.count=sum(as.numeric(!is.na(!! varname))))
+bres <- bmeans %>% ungroup() %>% summarise(between.sd = sd(meanx, na.rm=TRUE), between.min = min(meanx, na.rm=TRUE), between.max=max(meanx, na.rm=TRUE), Units=sum(as.numeric(!is.na(t.count))), t.bar=mean(t.count, na.rm=TRUE))
+wdat <- data %>% group_by(!! loc.unit) %>% mutate(W.x = scale(!! varname, scale=FALSE))
+wres <- wdat %>% ungroup() %>% summarise(within.sd=sd(W.x, na.rm=TRUE), within.min=min(W.x, na.rm=TRUE), within.max=max(W.x, na.rm=TRUE))
+return(list(ores=ores,bres=bres,wres=wres))
+}
+
+# read data file
+data <- data.table(read.csv(file = here('data_hm.csv'), header = T, stringsAsFactors = F))
+#  View(data)
+
+# drop countries with unavailable data
+# data <- data[!(iso3c %in% c("BEN","COM","PSE","STP","SYC","SWZ")),]
+data <- data[!(iso3c %in% c("BEN","COM","GIN","LBR","OMN","PSE","STP","SYC","SWZ")),] # excluding also 1 unit observations
+
+# adjust the variables
+data[, c("PopTot","GDPpc") := .(log(PopTot), log(GDPpc))]
+
+# Add square of GDPpc and Inflation
[…]

### 2020-12-16 Scripts/Output tables to LaTeX.R

@@ -21,6 +21,19 @@ load(file = here("Results/bma_3y_gini_baseline.Rdata"))
 load(file = here("Results/bma_3y_top10_baseline.Rdata"))
 load(file = here("Results/bma_3y_top1_baseline.Rdata"))
 
+# pre-tax gini coefficient 
+load(file = here("Results/bma_3y_ginimarket_baseline.Rdata"))
+
+# 5y averages
+load(file = here("Results/bma_5y_gini_baseline.Rdata"))
+load('Results/bma_5y_top10_baseline.Rdata')
+load('Results/bma_5y_top1_baseline.Rdata')
+
+# Lags
+load(file = here("Results/bma_3y_gini_baseline_lag.Rdata"))
+load('Results/bma_3y_top10_baseline_lag.Rdata')
+load('Results/bma_3y_top1_baseline_lag.Rdata')
+
 #
 #
 
@@ -53,7 +66,7 @@ explvarnames <- c("Natural resources rents","Government expenditures","Gross dom
 
 # Bind short and long variables names together
 varnames <- cbind(explvarshort, explvarnames)  
-
+length(explvarnames)
 # Results hyper prior
 # results_all <- coef(bma_1y_gini_baseline, exact=T, std.coefs = T)
 # results_all <- coef(bma_5y_gini_baseline, exact=T, std.coefs = T)
@@ -66,7 +79,10 @@ varnames <- cbind(explvarshort, explvarnames)
 # 3y averages (baseline) 
 # results_all <- coef(bma_3y_gini_baseline, exact=T, std.coefs = T)
 # results_all <- coef(bma_3y_top10_baseline, exact=T, std.coefs = T)
-results_all <- coef(bma_3y_top1_baseline, exact=T, std.coefs = T)
+# results_all <- coef(bma_3y_top1_baseline, exact=T, std.coefs = T)
+
+# 3y averages, pre-tax Gini
+results_all <- coef(bma_3y_top1_baseline_lag, exact=T, std.coefs = T)
 
 # List of variables sorted by PIP 
 varlist <- data.frame(explvarshort=row.names(results_all))
